### 2025-03-05 paper/paper.tex
Universal commits presents...

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{a4wide}

\title{Neural Networks Loss Landscape Convergence in Different Low-Dimensional Spaces}
\author{Tem Nikitin, Nikita Kiselev, Vladislav Meshkov, Andrey Grabovoy}
\date{2025}
\begin{document}
\maketitle

\begin{abstract}
    Understanding how the loss landscape of neural networks evolves as the training set size increases is crucial for
    optimizing performance and ensuring reliable generalization. While it is well known that larger datasets can alter
    the shape of this high-dimensional landscape, the exact point at which additional data no longer brings substantial
    changes remains underexplored.

    In this paper, we examine neural network models and show that their loss landscapes begin to stabilize once the
    training set grows beyond a certain threshold, revealing a connection between dataset size and the geometry of the
    loss surface. To elucidate this phenomenon, we propose a method that projects the full parameter space onto a
    low-dimensional subspace derived from top eigenvectors (e.g., from the Hessian). Focusing on these principal directions
    preserves critical curvature information while providing a more interpretable view of how the loss surface in the
    vicinity of local minima behaves as more data become available. We further leverage targeted sampling strategies,
    applying Monte-Carlo estimation to capture the structure of this reduced loss landscape more precisely.

    We validate our insights through comprehensive experiments on image classification tasks, demonstrating that this
    low-dimensional analysis can reveal when the landscape effectively settles, and thus helps determine a minimum viable
    dataset size. Our findings shed light on the relationship between dataset scale and optimization geometry, and suggest
    practical strategies for balancing computational costs with the benefits of additional training data.
\end{abstract}
\paragraph{Keywords:} 
Neural networks, Loss landscape, Low-dimensional subspace, Hessian eigenvectors, Monte Carlo estimation, Dataset size threshold.

\section{Introduction}

\bibliographystyle{unsrt}
\bibliography{bibliography}
\end{document}

### 2025-03-13 paper/main.tex
fixed abstract, added introduction, related work, preliminaries and a lot of typo mistakes

\documentclass{article}
\input{new_commands}

% if you need to pass options to natbib, use, e.g.:
\PassOptionsToPackage{numbers, sort, compress}{natbib}
% before loading neurips_2024

% ready for submission
%\usepackage{neurips_2024}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
\usepackage[preprint]{neurips_2025}

% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2024}

% to avoid loading the natbib package, add option nonatbib:
%    \usepackage[nonatbib]{neurips_2024}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors

%%%

\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{xcolor}
\usepackage{textcomp}
\usepackage{manyfoot}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{listings}

\newtheorem{theorem}{Theorem} % continuous numbers
%%\newtheorem{theorem}{Theorem}[section] % sectionwise numbers
%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
\newtheorem{proposition}[theorem]{Proposition}% 
\newtheorem{lemma}{Lemma}% 
%%\newtheorem{proposition}{Proposition} % to get separate numbers for theorem and proposition etc.

\newtheorem{example}{Example}
\newtheorem{remark}{Remark}

\newtheorem{definition}{Definition}
\newtheorem{assumption}{Assumption}

%%%

\title{Neural Networks Loss Landscape Convergence in Different Low-Dimensional Spaces}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{
  Tem Nikitin\\
  Moscow Institute of Physics and Technology\\
  Moscow, Russia\\
  \texttt{nikitin.artem.a@phystech.su}\\
  \And
  Nikita Kiselev\\
  Moscow Institute of Physics and Technology\\
  Moscow, Russia\\
  \texttt{kiselev.ns@phystech.su}\\
  \And
  Andrey Grabovoy\\
  Moscow Institute of Physics and Technology\\
  Moscow, Russia\\
  \texttt{grabovoy.av@phystech.su}\\
  \And
  Vladislav Meshkov\\
  Moscow Institute of Physics and Technology\\
  Moscow, Russia\\
  \texttt{meshkov.ns@phystech.su}\\
}

\begin{document}

\maketitle

\begin{abstract}
  Understanding how the loss landscape of neural networks evolves as the training set size increases is crucial for
  optimizing performance. It is well known that larger datasets can alter the shape of this high-dimensional landscape.
  But the exact point at which additional data no longer brings substantial changes remains underexplored.

  In this paper, we examine neural network models and show that their loss landscapes begin to stabilize once the
  training set grows beyond a certain threshold, revealing a connection between dataset size and the geometry of the
  loss surface. To understand this phenomenon, we propose a method that projects the full parameter space onto a
  low-dimensional subspace derived from top eigenvectors of the Hessian. That provides a more interpretable view of how
  the loss surface in the vicinity of local minima behaves as more data become available. We use different sampling
  strategies, applying Monte-Carlo estimation to capture the structure of this reduced loss landscape more precisely.

  We validate our insights through experiments on image classification tasks, demonstrating that this low-dimensional
  analysis can reveal when the landscape effectively settles, and thus helps determine a minimum viable dataset size.
  Our findings shed light on the relationship between dataset scale and optimization geometry, and suggest
  practical strategies for balancing computational costs with the benefits of additional training data.
\end{abstract}

\textbf{Keywords:}
Neural networks, Loss landscape, Low-dimensional subspace, Hessian eigenvectors, Monte Carlo estimation, Dataset size threshold.

\section{Introduction}\label{sec:intro}

Neural networks have become irreplaceable in various aspects of life and have a lot of applications now
(e.g. image classification, language models, recommender systems, etc.). However, as datasets and models grow larger and
larger, we get higher accuracy and breakthrough results. However some problems connected with computation resources have
appeared. Sometimes models have so many parameters so we need time we cannot provide for their training. Some attempts to
explore and compare optimization methods have already been made \cite{soydaner2020comparison}. The core problem of all these
solutions hides in their locality, as the size of the neural network and training dataset remains unchanged. Our research goal
is to explore when adding more data no longer brings significant improvements in network performance and to explain
obtained results in terms of the curvature and shape of the neural network’s loss landscape.

The object of research centers on the loss landscape of neural networks. Specifically, how it changes while adding new samples
to the training dataset. There remains a problem: processing of large datasets requires a lot of time.
We need to determine when increasing dataset size stops reshaping this loss landscape for general neural networks.
But the question of a minimum viable dataset size --- a threshold beyond which further data bring insignificant changes ---
remains underexplored. Moreover, connecting such a threshold to generalization capabilities \cite{wu2017towards} will
reduce computation requirements as we show that there is an upper bound of "active" dataset samples.

We suggest using the Hessian of the loss function to calculate its close approximation. However, the computation of all
Hessian requires a bunch of time, so we consider a projection of neural network parameters space to a certain subspace.
We place the following tasks to the front:
\begin{enumerate}
  \item Constructing a Hessian-based projection of the loss landscape which finds main curvature directions
        (top eigenvectors) in parameter space.
  \item Using different sampling methods, identifying at which dataset size the shape of the loss function stabilizes.
  \item We apply visualization methods to characterize the loss landscape next to the local minima.
  \item Providing applicable theoretical criteria to determine when adding more samples has minimal effect.
        We refer to The Matrix Cookbook \cite{petersen2012matrix} as a hint for matrix operations.
\end{enumerate}

The proposed solution is novel as it links Hessian estimation to the idea of a minimal dataset threshold. This enables
advantages such as more cost-effective dataset collection and a clearer understanding of how dataset size interacts with the
geometry of the loss function. And applying recent methods for low-rank Hessian estimation
can speed up and solve the problem of high computation cost of top eigenvalues for very large networks.

We validate our analysis results on well-known image classification tasks like MNIST \cite{deng2012mnist} and Fashion-MNIST
\cite{xiao2017fashion}. By connecting practical results to theoretical bounds, we offer a method for identifying
a computation-efficiency method that balances performance with accuracy gains.

\textbf{Contributions.} Our contributions can be summarized as follows:
\begin{itemize}
  \item We present a Hessian-based approach that uses a projection into a low-dimensional subspace of top eigenvectors to
        find the critical sufficient dataset size.
  \item We demonstrate the validity of our theoretical framework through empirical studies on MNIST and Fashion-MNIST,
        incrementally increasing dataset size until additional data makes an impact to the loss function curvature.
  \item We highlight the implications of our findings for practical data collection strategies, showing how the detection of
        a sufficient dataset size can reduce computation time.
\end{itemize}

\textbf{Outline.} The rest of the paper is organized as follows...but I don't know how yet.

\section{Related Work}\label{sec:rw}

\textbf{Hessian-based analysis.}
The landscape of loss functions has been explored from various perspectives in the literature. Methods connected with Hessian
are central to understanding the convergence process and loss function landscape \cite{kiselev2024unraveling}. Recent works
observed that the Hessian of some models tends to have a low effective rank, with only a small quantity of eigenvalue
differences from zero. Empirical explorations from \cite{sagun2018empirical} show that the Hessian rank is often very low in
the vicinity of local minima, highlighting a significantly smaller “active subspace”.
Prior work investigating how large eigenvalues emerge during training \cite{ghorbani2019investigation}.

Some results have already been reached for fully connected and convolutional architectures \cite{meshkov2024convnets},
but they need generalization.

\textbf{Loss Landscape Geometry.}
Other recent researches aimed at how dataset size affects both performance and the geometry of loss function landscapes.
In the assumption of enough training samples, neural networks often reach smoother and flatter minima with better
generalization \cite{wu2017towards}. However, collecting large datasets is expensive, as well as the computation
requirements for them, raising questions about the most effective ratio between dataset and model sizes
\cite{hoffmann2022training}. As understanding the landscape of loss functions is key to the solution, prior work has looked
at visualizing loss surfaces \cite{li2018visualizing}.

\section{Preliminaries}\label{sec:prelim}

\subsection{General notation}

In this section, we introduce the general notation used in the rest of the paper and the basic assumptions.

We consider a $K$-label classification problem.
So, let's consider $p(y|x)$ a conditional probability, which maps unobserved variable $x \in \mathcal{X}$ to the corresponding
output $y \in \mathcal{Y}$. We consider $\mathcal{Y}$ is a subspace (or same space) as $\mathbb{R}^K$. Let $f_{\omega}(\cdot)$
be a neural network with a list of parameters $\omega$. Let $\Omega$ be a space of parameters ($\omega \in \Omega$).

Let
$$\mathcal{D} = \left\{ (x_i, y_i) \, | \, i = 1, \dots, m \right\}$$
be a given dataset of size $m$ consists of i.i.d.
samples. Let $l(\cdot, \cdot)$ be a given loss function (e.g. cross-entrophy) where first argument refers to neural network's
result and the second argument refers to the true answer. To simplify, define: $$l_i(\omega) := l(f_{\omega}(x_i), y_i).$$

\begin{definition}
  The empirical loss function for the first $k$ elements is:
  $$\mathcal{L}_k(\omega) = \frac1k \sum\limits_1^k l_i(\omega), \,\, \mathcal{L}(\omega) := \mathcal{L}_m(\omega).$$
\end{definition}

Thus, the difference between losses for neighbouring samplem sizes is:
$$\mathcal{L}_{k}(\omega) - \mathcal{L}_{k-1}(\omega) = \frac{l_{k}(\omega) - \mathcal{L}_{k-1}(\omega)}{k}.$$

\begin{definition}
  Let the Hessian of $\mathcal{L}_k(\omega)$ be:
  $$\mathcal{H}_k(\omega) = \nabla^2_{\omega}\mathcal{L}_k(\omega) = \frac1k \sum\limits_1^k \nabla^2_{\omega} l_i(\omega).$$
\end{definition}

\begin{definition}
  To calculate the overall loss landscape changing, one has to integrate the absolute difference for the entire parameter space.
  We define \textbf{weighted difference} as:
  $$\Delta_{k} = \int \left( \mathcal{L}_{k}(\omega) - \mathcal{L}_{k-1}(\omega) \right)^2 p(\omega) d\omega,$$
  where $p(\omega)$ describes the priority of the particular parameter points so we can make $p(\omega)$ have higher values
  next to the local minima.
\end{definition}

We firther investigate this difference and aimed at exploration of how adding a new object to the dataset changes the value.
We interested in convergence of this value and properties of loss function when the training dataset size limits to $\infty$.

\begin{definition}
  Let $\Delta$ be a positive hyperparameter that indicates the stop-difference for $\Delta_k$. If
  $$\Delta_{k^*} < \Delta \wedge \left( \forall k < k^*: \, \Delta_k \geq \Delta \right),$$
  we can say that $k^*$ samples in the dataset are enough to describe the distribution of data from the general population.
  We call $k^*$ as \textbf{sufficient}.
\end{definition}

\subsection{Assumptions}

\begin{assumption}
  Let $\omega^*$ be the local minimum of both $\mathcal{L}_{k-1}(\omega)$ and $\mathcal{L}_{k}(\omega)$.
  Thus, $$\nabla \mathcal{L}_{k-1}(\omega^*) = \nabla \mathcal{L}_{k}(\omega^*) = 0.$$
\end{assumption}

This assumption allows us to explore the behavior and the geometry of the loss function landscape at only one point.

Furthermore, using second-order Taylor's approximation for $\mathcal{L}_{k}(\omega)$ at $\omega^*$ we get:
$$\mathcal{L}_{k}(\omega) \approx \mathcal{L}_{k}(\omega^*) + \frac12 (\omega - \omega^*)^T \mathcal{H}_k(\omega^*)
  (\omega - \omega^*)$$

\begin{assumption}
  We can assume parameters w to be random, which will lead to quite natural condition: p(w) can be even a prior
  distribution of w, so: $$\Delta_{k} = \mathbb{E}_{p(\omega)} \left( \mathcal{L}_{k}(\omega) - \mathcal{L}_{k-1}(\omega)
    \right)^2 = \mathbb{D} \left( \mathcal{L}_{k}(\omega) - \mathcal{L}_{k-1}(\omega) \right) + \left( \mathbb{E} \left(
      \mathcal{L}_{k}(\omega) - \mathcal{L}_{k-1}(\omega) \right) \right)^2$$
\end{assumption}

\section{Method}\label{sec:method}

As the work with the full Hessian is computation complex, let's define a projection of neural network parameters space
$\Omega$ to a certain subspace $\Theta$. Let $\mathcal{P}$ be a projection matrix from $\Omega$ to $\Theta$. Let $\theta$ be
a projection: $\theta = \mathcal{P}\omega$.

\section{Experiments}\label{sec:exp}

To verify the theoretical estimates obtained, we conducted a detailed empirical study...

\section{Discussion}\label{sec:disc}

TADAAA!

\section{Conclusion}\label{sec:concl}

TADAAA!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{unsrtnat}
\bibliography{references}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\appendix
\section{Appendix / supplemental material}\label{app}

\subsection{Additional experiments / Proofs of Theorems}\label{app:exp}

TADAAA!

\end{document}

### 2025-03-13 paper/main.tex
Universal commits presents...

@@
 
-\textbf{Outline.} The rest of the paper is organized as follows...but I don't know how yet.
+\textbf{Outline.} The rest of the paper is organized as follows. Section 2 divides existing results into some
+topics, highlighting their key contributions and findings. Section 3 considers general notation and
+some preliminary calculations. In Section 4, we provide theoretical bounds on the hessian and losses
+difference norms. Empirical study of the obtained results is given in Section 5. We summarize and
+present the results in Sections 6 and 7. Additional experiments and proofs of theorems are included
+in the Appendix A.
 

### 2025-03-13 paper/main.tex
Universal commits presents...

@@
 explore and compare optimization methods have already been made \cite{soydaner2020comparison}. The core problem of all these
-solutions hides in their locality, as the size of the neural network and training dataset remains unchanged. Our research goal
-is to explore when adding more data no longer brings significant improvements in network performance and to explain
+solutions hides in their locality, as the size of the neural network and training dataset remains unchanged. In this paper we
+explore when adding more data no longer brings significant improvements in network performance and to explain
 obtained results in terms of the curvature and shape of the neural network’s loss landscape.
 
-The object of research centers on the loss landscape of neural networks. Specifically, how it changes while adding new samples
+Our research on the loss landscape of neural networks. Specifically, how it changes while adding new samples
 to the training dataset. There remains a problem: processing of large datasets requires a lot of time.
@@
 We consider a $K$-label classification problem.
-So, let's consider $p(y|x)$ a conditional probability, which maps unobserved variable $x \in \mathcal{X}$ to the corresponding
-output $y \in \mathcal{Y}$. We consider $\mathcal{Y}$ is a subspace (or same space) as $\mathbb{R}^K$. Let $f_{\omega}(\cdot)$
-be a neural network with a list of parameters $\omega$. Let $\Omega$ be a space of parameters ($\omega \in \Omega$).
+So, let's consider $p(y|\bx)$ a conditional probability, which maps unobserved variable $\bx \in \bX$ to the corresponding
+output $y \in \by$. We consider $\by$ is a subspace (or same space) as $\mathbb{R}^K$. Let $f_{\bw}(\cdot)$
+be a neural network with a list of parameters $\bw$. Let $\Omega$ be a space of parameters ($\bw \in \Omega$).
 
 Let
-$$\mathcal{D} = \left\{ (x_i, y_i) \, | \, i = 1, \dots, m \right\}$$
+$$\mathcal{D} = \left\{ (\bx_i, y_i) \, | \, i = 1, \dots, m \right\}$$
 be a given dataset of size $m$ consists of i.i.d.
-samples. Let $l(\cdot, \cdot)$ be a given loss function (e.g. cross-entrophy) where first argument refers to neural network's
-result and the second argument refers to the true answer. To simplify, define: $$l_i(\omega) := l(f_{\omega}(x_i), y_i).$$
+samples. Let $\ell(\cdot, \cdot)$ be a given loss function (e.g. cross-entrophy) where first argument refers to neural network's
+result and the second argument refers to the true answer. To simplify, define:
+$$\ell_i(\bw) := \ell(f_{\bw}(\bx_i), y_i).$$
 
@@
   The empirical loss function for the first $k$ elements is:
-  $$\mathcal{L}_k(\omega) = \frac1k \sum\limits_1^k l_i(\omega), \,\, \mathcal{L}(\omega) := \mathcal{L}_m(\omega).$$
+  $$\mathcal{L}_k(\bw) = \frac1k \sum\limits_1^k \ell_i(\bw), \,\, \mathcal{L}(\bw) := \mathcal{L}_m(\bw).$$
 \end{definition}
@@
 Thus, the difference between losses for neighbouring samplem sizes is:
-$$\mathcal{L}_{k}(\omega) - \mathcal{L}_{k-1}(\omega) = \frac{l_{k}(\omega) - \mathcal{L}_{k-1}(\omega)}{k}.$$
+$$\mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw) = \frac{\ell_{k}(\bw) - \mathcal{L}_{k-1}(\bw)}{k}.$$
 
 \begin{definition}
-  Let the Hessian of $\mathcal{L}_k(\omega)$ be:
-  $$\mathcal{H}_k(\omega) = \nabla^2_{\omega}\mathcal{L}_k(\omega) = \frac1k \sum\limits_1^k \nabla^2_{\omega} l_i(\omega).$$
+  Let the Hessian of $\mathcal{L}_k(\bw)$ be:
+  $$\bH_k(\bw) = \nabla^2_{\bw}\mathcal{L}_k(\bw) = \frac1k \sum\limits_1^k \nabla^2_{\bw} \ell_i(\bw).$$
 \end{definition}
@@
   We define \textbf{weighted difference} as:
-  $$\Delta_{k} = \int \left( \mathcal{L}_{k}(\omega) - \mathcal{L}_{k-1}(\omega) \right)^2 p(\omega) d\omega,$$
-  where $p(\omega)$ describes the priority of the particular parameter points so we can make $p(\omega)$ have higher values
+  $$\Delta_{k} = \int \left( \mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw) \right)^2 p(\bw) dw,$$
+  where $p(\bw)$ describes the priority of the particular parameter points so we can make $p(\bw)$ have higher values
   next to the local minima.
@@
   Let $\Delta$ be a positive hyperparameter that indicates the stop-difference for $\Delta_k$. If
-  $$\Delta_{k^*} < \Delta \wedge \left( \forall k < k^*: \, \Delta_k \geq \Delta \right),$$
+  $$k^* = \inf\limits_k\{\Delta_k < \Delta\}$$
   we can say that $k^*$ samples in the dataset are enough to describe the distribution of data from the general population.
@@
 \begin{assumption}
-  Let $\omega^*$ be the local minimum of both $\mathcal{L}_{k-1}(\omega)$ and $\mathcal{L}_{k}(\omega)$.
-  Thus, $$\nabla \mathcal{L}_{k-1}(\omega^*) = \nabla \mathcal{L}_{k}(\omega^*) = 0.$$
+  Let $\bw^*$ be the local minimum of both $\mathcal{L}_{k-1}(\bw)$ and $\mathcal{L}_{k}(\bw)$.
+  Thus, $$\nabla \mathcal{L}_{k-1}(\bw^*) = \nabla \mathcal{L}_{k}(\bw^*) = 0.$$
 \end{assumption}
@@
 
-Furthermore, using second-order Taylor's approximation for $\mathcal{L}_{k}(\omega)$ at $\omega^*$ we get:
-$$\mathcal{L}_{k}(\omega) \approx \mathcal{L}_{k}(\omega^*) + \frac12 (\omega - \omega^*)^T \mathcal{H}_k(\omega^*)
-  (\omega - \omega^*)$$
+Furthermore, using second-order Taylor's approximation for $\mathcal{L}_{k}(\bw)$ at $\bw^*$ we get:
+$$\mathcal{L}_{k}(\bw) \approx \mathcal{L}_{k}(\bw^*) + \frac12 (\bw - \bw^*)^{\T} \bH_k(\bw^*)
+  (\bw - \bw^*)$$
 
 \begin{assumption}
-  We can assume parameters w to be random, which will lead to quite natural condition: p(w) can be even a prior
-  distribution of w, so: $$\Delta_{k} = \mathbb{E}_{p(\omega)} \left( \mathcal{L}_{k}(\omega) - \mathcal{L}_{k-1}(\omega)
-    \right)^2 = \mathbb{D} \left( \mathcal{L}_{k}(\omega) - \mathcal{L}_{k-1}(\omega) \right) + \left( \mathbb{E} \left(
-      \mathcal{L}_{k}(\omega) - \mathcal{L}_{k-1}(\omega) \right) \right)^2$$
+  We can assume parameters $\bw$ to be random, which will lead to quite natural condition: $p(\bw)$ can be even a prior
+  distribution of $\bw$, so:
+  $$\Delta_{k} = \mathbb{E}_{p(\bw)} \left( \mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw)
+    \right)^2 = \mathbb{D} \left( \mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw) \right) + \left( \mathbb{E} \left(
+      \mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw) \right) \right)^2$$
 \end{assumption}
@@
 As the work with the full Hessian is computation complex, let's define a projection of neural network parameters space
-$\Omega$ to a certain subspace $\Theta$. Let $\mathcal{P}$ be a projection matrix from $\Omega$ to $\Theta$. Let $\theta$ be
-a projection: $\theta = \mathcal{P}\omega$.
+$\Omega$ to a certain subspace $\Theta$. Let $\bP$ be a projection matrix from $\Omega$ to $\Theta$. Let $\btheta$ be
+a projection: $\btheta = \bP\bw$.
 

### 2025-03-20 paper/main.tex
Universal commits presents...

@@
   Let $\Delta$ be a positive hyperparameter that indicates the stop-difference for $\Delta_k$. If
-  $$k^* = \inf\limits_k\{\Delta_k < \Delta\}$$
+  $$k^* = \inf\limits_k\{\forall m \geq k: \, \Delta_m < \Delta\}$$
   we can say that $k^*$ samples in the dataset are enough to describe the distribution of data from the general population.

### 2025-03-20 code/src/calc.py

@@ -1,0 +1,100 @@
+import copy
+import scipy.stats as sps
+import numpy as np
+
+from src.params_directions import create_random_direction
+from src.params_directions import init_from_params
+from src.params_directions import inplace_sum_models
+from src.utils import create_losses_func
+
+from tqdm.auto import tqdm
+from src.params_directions import calc_sum_models
+
+
+class LossCalculator:
+    def __init__(self, model, loader, criterion):
+        self.model = model
+        self.calc_losses_func = create_losses_func(loader, criterion)
+
+    def calc_losses(self, coef_grid, direction_norm):
+        """
+            grid: 2d from [-1, 1]*[-1, 1]
+        """
+
+        result = {}
+        direction1 = create_random_direction(self.model, external_factor=direction_norm)
+        direction2 = create_random_direction(self.model, external_factor=direction_norm)
+
+        for coef1, coef2 in tqdm(coef_grid):
+            target_add = [p1 * coef1 + p2 * coef2 for p1, p2 in zip(direction1, direction2)]
+            target_add_model = copy.deepcopy(self.model)
+            init_from_params(target_add_model, target_add)
+
+            target_model = calc_sum_models(self.model, target_add_model, 1, 1)
+            losses = self.calc_losses_func(target_model)
+            result[(coef1, coef2)] = losses
+        return result
+
+
+class DeltaCalculator:
+    r"""
+        Base class for calc delta
+        delta_k = (L_{k+1} - L_k)
+    """
+
+    def __init__(self, model, loader, criterion):
+        self.model = model
+        self.calc_losses_func = create_losses_func(loader, criterion)
+        self.directions = None
+
+    def calc_shifted_losses(self, mode, mode_params):
+        target_model = None
+        if mode == 'random-subspace-proj':
+            if self.directions is None:
+                self.directions = [create_random_direction(self.model, external_factor=1.0) for _ in
+                                   range(mode_params['dim'])]
+            coefs = list(sps.norm(np.zeros(mode_params['dim']), mode_params['sigma']).rvs())
+            target_add_params = [coef * d[i] for coef, d in zip(coefs, self.directions) for i in
+                                 range(len(self.directions[0]))]
+
+            target_model = copy.deepcopy(self.model)
+            init_from_params(target_model, target_add_params)
+
+            target_model = inplace_sum_models(target_model, self.model, 1.0, 1.0)
+
[…]

### 2025-03-20 code/src/models.py

@@ -1,0 +1,52 @@
+# --========================================-- #
+#   * Author  : NTheme - All rights reserved
+#   * Created : 20 March 2025, 1:40 AM
+#   * File    : models.py
+#   * Project : untitled
+# --========================================-- #
+
+import torch.nn as nn
+import torch.nn.functional as F
+
+
+class MLP(nn.Module):
+    def __init__(self, layers_num, hidden, input_channels, input_sizes, classes):
+        """
+        layers_num: количество скрытых слоёв
+        hidden: число нейронов в каждом скрытом слое
+        input_channels: число каналов во входном изображении
+        input_sizes: (height, width)
+        classes: число выходных классов
+        """
+        super(MLP, self).__init__()
+
+        self.layers_num = layers_num
+        self.hidden = hidden
+        self.input_channels = input_channels
+        self.input_sizes = input_sizes
+        self.classes = classes
+
+        input_dim = input_channels * input_sizes[0] * input_sizes[1]
+
+        layers = [nn.Linear(input_dim, hidden)]
+        for _ in range(layers_num - 1):
+            layers.append(nn.Linear(hidden, hidden))
+        layers.append(nn.Linear(hidden, classes))
+
+        self.linears = nn.ModuleList(layers)
+
+    def forward(self, x):
+        # Разворачиваем входные изображения в вектор
+        x = x.view(-1, self.input_channels * self.input_sizes[0] * self.input_sizes[1])
+
+        # Проходим по всем слоям, добавляя ReLU после каждого скрытого слоя
+        for i, layer in enumerate(self.linears):
+            x = layer(x)
+            # ReLU для всех слоёв, кроме выходного
+            if i < self.layers_num:
+                x = F.relu(x)
+        return x
+
+    @property
+    def device(self):
+        return next(iter(self.parameters())).device

### 2025-03-20 code/src/params_directions.py

@@ -1,0 +1,130 @@
+import torch
+import copy
+
+
+def get_weights(net):
+    """ Extract parameters from net, and return a list of tensors"""
+    return [p.data for p in net.parameters()]
+
+
+def get_random_weights(weights):
+    """
+        Produce a random direction that is a list of random Gaussian tensors
+        with the same shape as the network's weights, so one direction entry per weight.
+    """
+    return [torch.randn(w.size(), device=w.device) for w in weights]
+
+
+def normalize_direction(direction, weights, norm='filter'):
+    """
+        Rescale the direction so that it has similar norm as their corresponding
+        model in different levels.
+
+        Args:
+          direction: a variables of the random direction for one layer
+          weights: a variable of the original model for one layer
+          norm: normalization method, 'filter' | 'layer' | 'weight'
+    """
+    if norm == 'filter':
+        # Rescale the filters (weights in group) in 'direction' so that each
+        # filter has the same norm as its corresponding filter in 'weights'.
+        for d, w in zip(direction, weights):
+            d.mul_(w.norm() / (d.norm() + 1e-10))
+    elif norm == 'layer':
+        # Rescale the layer variables in the direction so that each layer has
+        # the same norm as the layer variables in weights.
+        direction.mul_(weights.norm() / direction.norm())
+    elif norm == 'weight':
+        # Rescale the entries in the direction so that each entry has the same
+        # scale as the corresponding weight.
+        direction.mul_(weights)
+    elif norm == 'dfilter':
+        # Rescale the entries in the direction so that each filter direction
+        # has the unit norm.
+        for d in direction:
+            d.div_(d.norm() + 1e-10)
+    elif norm == 'dlayer':
+        # Rescale the entries in the direction so that each layer direction has
+        # the unit norm.
+        direction.div_(direction.norm())
+
+
+def normalize_directions_for_weights(direction, weights, norm='filter', ignore='biasbn'):
+    """
+        The normalization scales the direction entries according to the entries of weights.
+    """
+    assert (len(direction) == len(weights))
+    for d, w in zip(direction, weights):
+        if d.dim() <= 1:
+            if ignore == 'biasbn':
+                d.fill_(0)  # ignore directions for weights with 1 dimension
+            else:
+                d.copy_(w)  # keep directions for weights/bias that are only 1 per node
+        else:
+            normalize_direction(d, w, norm)
+
+
+def inplace_sum_models(model1, model2, coef1, coef2):
[…]

### 2025-03-20 code/src/utils.py

@@ -1,0 +1,106 @@
+# --========================================-- #
+#   * Author  : NTheme - All rights reserved
+#   * Created : 20 March 2025, 1:33 AM
+#   * File    : utils.py
+#   * Project : untitled
+# --========================================-- #
+
+import torch
+from torch.utils.data import DataLoader
+from torchvision import datasets
+import numpy as np
+from tqdm.auto import tqdm
+
+
+def init_dataloader(dataset_name,
+                    transform,
+                    batch_size=64,
+                    dataset_load_path='data/',
+                    train_mode=True,
+                    size=None):
+    if dataset_load_path[-1] != '/':
+        dataset_load_path += '/'
+    dataset_name = dataset_name
+    dataset_load_path = dataset_load_path + dataset_name + '_dataset'
+
+    name2dataset = {
+        'MNIST': datasets.MNIST,
+        'FashionMNIST': datasets.FashionMNIST,
+        'CIFAR10': datasets.CIFAR10,
+        'CIFAR100': datasets.CIFAR100,
+    }
+
+    dataset = name2dataset[dataset_name](dataset_load_path,
+                                         download=True,
+                                         train=train_mode,
+                                         transform=transform)
+    if size is not None and size != -1:
+        dataset = torch.utils.data.Subset(dataset, np.arange(size))
+    loader = DataLoader(dataset, batch_size=batch_size, shuffle=train_mode, num_workers=2)
+    return loader
+
+
+def train(model,
+          optimizer,
+          criterion,
+          dataloader,
+          num_epochs,
+          device, log=False):
+    model.train()
+
+    losses = []
+    for epoch in tqdm(np.arange(1, num_epochs + 1)):
+        running_loss = 0.0
+        batches_count = 0
+
+        for x, y in dataloader:
+            x, y = x.to(device), y.to(device)
+
+            optimizer.zero_grad()
+            output = model(x)
+            loss = criterion(output, y)
+            loss.backward()
+            optimizer.step()
+
+            losses.append(loss.detach().cpu().item())
+
+            batches_count += 1
+            running_loss += losses[-1]
+
+        avg_loss = running_loss / batches_count
+        if log:
+            print(f"Epoch [{epoch + 1}/{num_epochs}]: loss = {avg_loss:.4f}")
+
+    return losses
+
+
+def valid(model,
+          criterion,
+          dataloader,
+          device):
+    model.eval()
+
+    acc = []
+    losses = []
[…]

### 2025-03-20 code/src/visualize.py

@@ -1,0 +1,297 @@
+import itertools
+from matplotlib import cm
+import numpy as np
+import matplotlib.pyplot as plt
+from mpl_toolkits.mplot3d import Axes3D
+import copy
+import scipy.stats as sps
+from tqdm.auto import tqdm
+
+from src.calc import DeltaCalculator
+from src.calc import LossCalculator
+
+
+class LossVisualizer:
+    def __init__(self, model, loader, criterion, grid_step=0.1, direction_norm=1):
+        """
+            grid_step in [0, 1]
+        """
+        self.grid_loss = None
+        self.grid_step = grid_step
+        self.direction_norm = direction_norm
+
+        self.coef_grid = list(itertools.product(np.arange(-1, 1 + self.grid_step, step=self.grid_step),
+                                                np.arange(-1, 1 + self.grid_step, step=self.grid_step)))
+
+        self.loss_calc = LossCalculator(model, loader, criterion)
+        self.grid_loss = self.loss_calc.calc_losses(self.coef_grid, self.direction_norm)
+
+    def _set_xy_grid(self, x_grid_bounds, y_grid_bounds):
+        xs = np.arange(-1, 1 + self.grid_step, step=self.grid_step)
+        ys = xs
+        xs = np.array([x for x in xs if x_grid_bounds[1] >= x >= x_grid_bounds[0]])
+        ys = np.array([y for y in ys if y_grid_bounds[1] >= y >= y_grid_bounds[0]])
+        return xs, ys
+
+    def visualize_diff(self, size1, size2,
+                       x_grid_bounds=(-1, 1), y_grid_bounds=(-1, 1),
+                       diff_type=None, bounds=None, distrib_params=None):
+        """
+           type: None | abs | square | realative | relative_abs | relative_squar | square_dot_normal | abs_dot_normal
+           bounds: None | (-min, max)
+        """
+        assert bounds is None or bounds[0] < bounds[1]
+        grid_loss = self.grid_loss
+
+        xs, ys = self._set_xy_grid(x_grid_bounds, y_grid_bounds)
+        xgrid, ygrid = np.meshgrid(xs, ys)
+        zgrid1 = np.array([[np.mean(grid_loss[(x, y)][:size1]) for x in xs] for y in ys])
+        zgrid2 = np.array([[np.mean(grid_loss[(x, y)][:size2]) for x in xs] for y in ys])
[…]

### 2025-03-20 paper/main.tex
Universal commits presents...

@@
   The empirical loss function for the first $k$ elements is:
-  $$\mathcal{L}_k(\bw) = \frac1k \sum\limits_1^k \ell_i(\bw), \,\, \mathcal{L}(\bw) := \mathcal{L}_m(\bw).$$
+  $$\mathcal{L}_k(\bw) = \frac1M \sum\limits_1^k \ell_i(\bw), \,\, \mathcal{L}(\bw) := \mathcal{L}_m(\bw).$$
 \end{definition}
@@
   To calculate the overall loss landscape changing, one has to integrate the absolute difference for the entire parameter space.
-  We define \textbf{weighted difference} as:
+  We define $\mathbf{\Delta}$-\textbf{function} (delta-function) as:
   $$\Delta_{k} = \int \left( \mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw) \right)^2 p(\bw) dw,$$
@@
 
+Let $\mathbf{V} = \{\mathbf{e}_1, \dots, \mathbf{e}_d\}$ be eigenvectors of $\mathbf{H}$. We select $d$ of them and project parameters space
+into a subspace on these $d$ vectors. These vectors should represent maximum eigenvalues of $\mathbf{H}$ as these directions
+have the most impact to the loss function difference.
 \section{Experiments}\label{sec:exp}
 
-To verify the theoretical estimates obtained, we conducted a detailed empirical study...
+To verify the theoretical estimates obtained, we conducted a detailed empirical study.
+In this section, we present experiments aimed at analyzing how the loss landscape evolves as the size of the training
+set increases.
+
+\paragraph{Data Preparation.}
+We use the MNIST dataset, which consists of $60{,}000$ training images and $10{,}000$ test
+images of handwritten digits (0--9). Each image is $28\times28$ pixels, grayscale.
+
+\paragraph{Model Architecture.}
+We use the following neural network models and preparations:
+\begin{itemize}
+  \item A multilayer perceptron (MLP) with different number of hidden layers (ReLU activations).
+  \item The choice of architecture remains consistent across all experiments.
+\end{itemize}
+
+The first two are preliminary experiments, conducted as illustrative examples on the MNIST dataset,
+while the others focusing on Hessian-based projections and show connections between eigenvalues and loss function convergence.
+That is our primary aim --- to check our theoretical estimates in-deal.
+
+\subsection{Visualizing the Loss Landscape in a Random Two-Dimensional Subspace}
+The idea of this initial experiment is to get an intuitive sense of how the loss function’s landscape shifts with
+different training-set sizes. To do this, we pick two random directions $\mathbf{v_1}$ and $\mathbf{v_2}$ in the parameter
+space and project the network’s parameters onto the subspace they span.
+
+\paragraph{Procedure}
+\begin{enumerate}
+  \item \textbf{Training on different sizes.} Train a neural network on increasingly larger subsets.
+  \item \textbf{Parameter snapshot.} After training, fix the final (or near-final) weights $\mathbf{w}^*$.
+  \item \textbf{Subspace grid.} Form a grid of points $\{\mathbf{w}^* + \alpha \mathbf{v}_1 + \beta \mathbf{v}_2\}$ around $\mathbf{w}^*$, where $\alpha$ and $\beta$ vary in a chosen range.
+  \item \textbf{Loss evaluation.} Compute and plot the loss at each grid point, producing a 3D visualization.
+\end{enumerate}
+
+For smaller $N$, the landscape exhibits sharper fluctuations. As $N$ grows, the landscape often appears smoother and more stable.
+
+\begin{figure}[!htbp]
+  \hspace*{-11.5cm}
+  \subfloat{\includegraphics[width=2\textwidth]{img/Ls_16.pdf}}
+  \caption{Loss function landscape}
+  \label{fig:base_loss}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \hspace*{-11.5cm}
+  \subfloat{\includegraphics[width=2\textwidth]{img/Ld_10.pdf}}
+  \caption{Loss function difference $(\mathcal{L}_{k} - \mathcal{L}_{k-1})^2$ landscape}
+  \label{fig:base_loss_diff}
+\end{figure}
+
+\subsection{Visualizing delta-function}
+
+Now we dig deeper and visualize our $\Delta_k$ function to clearly understand the difference in the loss function when
+moving from a model trained on $k-1$ samples to one trained on $k$ samples. To do this we sample a bunch of random gaussian
+vectors (all components are normally distributed) and calculate a
+math expectation of modulus of the loss function difference.
+Directly computing this expectation over the entire distribution of random directions $\mathbf{d}$
+is impossible, so we employ a Monte Carlo approximation: using Monte-Carlo method (we take the mean value of loss function
+on sampled gaussian shifted points):
+
+$$\Delta_k = \mathbb{E}_{p(\bw)} \left( \mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw) \right)^2 \approx
+  \frac{1}{k} \sum_{t=1}^{k} \left( \mathcal{L}_k(\mathbf{w}^* + \mathbf{d}_t) - \mathcal{L}_{k-1}(\mathbf{w}^* + \mathbf{d}_t) \right)^2$$
+
+where $\mathbf{d}_t \sim \mathcal{N}(0,I_d)$ are $K$ random gaussian samples drawn from a multivariate normal distribution
+with zero mean and identity covariance from a subspace of rank $d$. Specifically, we proceed as follows:
+
+\begin{enumerate}
+  \item \textbf{Sampling Directions.} Generate $K$ independent directions $\mathbf{d}_1, \dots, \mathbf{d}_M$
+        from $\mathcal{N}(0,I_d)$.
+  \item \textbf{Loss Evaluation.} For each direction $\mathbf{d}_t$, evaluate the loss at
+        $\mathbf{w}^* + \mathbf{d}_t$ and
+        $$\left( \mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw) \right)^2 $$
+  \item \textbf{Averaging.} Average these values values over $k = 1, \dots, K$ to approximate $\Delta_k$.
+\end{enumerate}
+
+By increasing $K$, we obtain a more reliable estimate of the expected difference in loss landscapes.
+This Monte Carlo approach is straightforward to implement and does not require explicit knowledge of the underlying loss
+function's distribution.
+
+The $\Delta$-function generally decreases as $N$ increases, suggesting that the landscape differences between neighbouring
+sizes become smaller.
+Beyond a certain point, $\Delta$-values are nearly negligible, indicating that adding more data has minimal impact on
+the local landscape in these directions.
+
+\noindent
+\textbf{Conclusion from Preliminary Experiments.}
+These initial experiments demonstrate that the loss landscape becomes progressively more stable with increasing $N$.
+However, the directions used here are either random. This motivates a more rigorous approach in the \emph{main experiment}
+below, where we focus specifically on the most critical directions of curvature --- those corresponding to the top
+eigenvalues of the Hessian.
+
+\subsection{Main Experiment: Projection onto Dominant Hessian Eigenvectors}
+
+\paragraph{Rationale}
+From prior work \cite{sagun2018empirical}, we know that only a few Hessian eigenvectors (those associated with the largest
+eigenvalues) tend to capture the most significant curvature directions. By projecting onto these principal directions,
+we can more precisely measure how the landscape the loss surface changes as $k$ increases.
+
+\paragraph{Procedure}
+\begin{enumerate}
+  \item \textbf{Hessian approximation.} For each trained model, approximate
+        the Hessian $\mathbf{H}$ or its dominant eigenvalues/eigenvectors. Techniques ???
+  \item \textbf{Eigen-decomposition.} Identify the top $d$ eigenvectors $\{\mathbf{e}_1, \dots, \mathbf{e}_d\}$ with the
+        largest eigenvalues.
+  \item \textbf{Subspace exploration.} For each model, calculate a $\Delta$ fuction via the same way as in the second
+        experiment (using Monte-Carlo) and appy visualization/.
+\end{enumerate}
+
+\paragraph{Goal and Expected Outcome}
+\begin{itemize}
+  \item \textbf{Refined understanding.} By focusing on the directions of largest curvature, we expect a clearer
+        measure of how the landscape of the loss surface changes with additional data.
+  \item \textbf{Identifying the threshold.} This approach should yield a more accurate estimate of the point at which the
+        landscape effectively stabilizes, thus helping to pinpoint a minimum viable dataset size.
+  \item \textbf{Connect with theoretical results.} By applying experiments we check our calculations above on the reality.
+\end{itemize}
+
+\begin{figure}[!htbp]
+  \hspace*{-0.7cm}
+  \subfloat{\includegraphics[width=1.15\textwidth]{img/D_32.pdf}}\\
+  \caption{$\Delta$-function visualization. `Sigma` parameter refers to the variance of sampled gaussian vectors, `dim' 
+  refers to the rank of the subspace.}
+  \label{fig:base_delta}
+\end{figure}
 

### 2025-03-20 code/src/visualize.py

@@ -105,6 +105,7 @@ class LossVisualizer:
                                   alpha=1.0)
         fig.colorbar(surf, shrink=0.5, aspect=5)
         ax_3d.view_init(40, 20)
+        plt.savefig(f"Ld_{size1}.eps")
         plt.show()
 
     def visualize(self, size=None,
@@ -139,7 +140,7 @@ class LossVisualizer:
                                   alpha=1.0)
         fig.colorbar(surf, shrink=0.5, aspect=5)
         ax_3d.view_init(40, 20)
-
+        plt.savefig(f"Ls_{size}.eps")
         plt.show()
 
 
@@ -156,7 +157,7 @@ class DeltaVisualizer:
 
         deltas = self.delta_calc.calc_deltas(mode, params, num_samples=num_samples)
 
-        fig, axs = plt.subplots(figsize=(16, 6), nrows=1, ncols=2)
+        fig, axs = plt.subplots(figsize=(9, 6), nrows=2, ncols=1)
         fig.suptitle(fr'$\Delta_k$; {params}')
         axs[0].plot(deltas, label="Test")
         mult_coef = np.arange(1, len(deltas) + 1)
@@ -185,7 +186,7 @@ class DeltaVisualizer:
             xlabel='k',
             ylabel=ylabels[1],
         )
-
+        plt.savefig("2.eps")
         plt.show()
 
     def compare_params(self,

### 2025-03-20 code/src/calc.py
Universal commits presents...

@@ -2,13 +2,13 @@ import copy
 import scipy.stats as sps
 import numpy as np
 
-from src.params_directions import create_random_direction
-from src.params_directions import init_from_params
-from src.params_directions import inplace_sum_models
+from src.directions import create_random_direction
+from src.directions import init_from_params
+from src.directions import inplace_sum_models
 from src.utils import create_losses_func
 
 from tqdm.auto import tqdm
-from src.params_directions import calc_sum_models
+from src.directions import calc_sum_models
 
 
 class LossCalculator:

### 2025-03-20 code/src/visualize.py

@@ -105,7 +105,8 @@ class LossVisualizer:
                                   alpha=1.0)
         fig.colorbar(surf, shrink=0.5, aspect=5)
         ax_3d.view_init(40, 20)
-        plt.savefig(f"Ld_{size1}.eps")
+
+        # plt.savefig(f"Ld_{size1}.eps")
         plt.show()
 
     def visualize(self, size=None,
@@ -140,7 +141,8 @@ class LossVisualizer:
                                   alpha=1.0)
         fig.colorbar(surf, shrink=0.5, aspect=5)
         ax_3d.view_init(40, 20)
-        plt.savefig(f"Ls_{size}.eps")
+
+        # plt.savefig(f"Ls_{size}.eps")
         plt.show()
 
 
@@ -186,7 +188,7 @@ class DeltaVisualizer:
             xlabel='k',
             ylabel=ylabels[1],
         )
-        plt.savefig("2.eps")
+        # plt.savefig("2.eps")
         plt.show()
 
     def compare_params(self,

### 2025-03-20 paper/main.tex
Universal commits presents...

@@
 
-\paragraph{Data Preparation.}
+\paragraph{Data.}
 We use the MNIST dataset, which consists of $60{,}000$ training images and $10{,}000$ test
@@
         from $\mathcal{N}(0,I_d)$.
+  \item \textbf{Decreasing multidimension.} Project the network’s parameters onto the subspace they span.
   \item \textbf{Loss Evaluation.} For each direction $\mathbf{d}_t$, evaluate the loss at
         $\mathbf{w}^* + \mathbf{d}_t$ and
-        $$\left( \mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw) \right)^2 $$
+        $$\left( \mathcal{L}_{k}(\bw + \mathbf{d}_t) - \mathcal{L}_{k-1}(\bw + \mathbf{d}_t) \right)^2 $$
   \item \textbf{Averaging.} Average these values values over $k = 1, \dots, K$ to approximate $\Delta_k$.
@@
 
-\paragraph{Rationale}
 From prior work \cite{sagun2018empirical}, we know that only a few Hessian eigenvectors (those associated with the largest
@@
         largest eigenvalues.
-  \item \textbf{Subspace exploration.} For each model, calculate a $\Delta$ fuction via the same way as in the second
-        experiment (using Monte-Carlo) and appy visualization/.
+  \item \textbf{Calculation and analization.} For each model, calculate a $\Delta$ fuction via the same way as in the second
+        experiment (using Monte-Carlo) and visualize via plots.
 \end{enumerate}
@@
 \begin{itemize}
-  \item \textbf{Refined understanding.} By focusing on the directions of largest curvature, we expect a clearer
+  \item By focusing on the directions of largest curvature, we expect a clearer
         measure of how the landscape of the loss surface changes with additional data.
-  \item \textbf{Identifying the threshold.} This approach should yield a more accurate estimate of the point at which the
+  \item This approach should yield a more accurate estimate of the point at which the
         landscape effectively stabilizes, thus helping to pinpoint a minimum viable dataset size.
-  \item \textbf{Connect with theoretical results.} By applying experiments we check our calculations above on the reality.
+  \item By applying experiments we check our calculations above on the reality.
 \end{itemize}
@@
 \begin{figure}[!htbp]
-  \hspace*{-0.7cm}
-  \subfloat{\includegraphics[width=1.15\textwidth]{img/D_32.pdf}}\\
+  \centering
+  \subfloat{\includegraphics[width=1\textwidth]{img/D_32.pdf}}\\
   \caption{$\Delta$-function visualization. `Sigma` parameter refers to the variance of sampled gaussian vectors, `dim'

### 2025-03-20 paper/one-slide.tex
Universal commits presents...

\documentclass[10pt]{beamer}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amssymb}
\usepackage{subfig}

\begin{document}

\begin{frame}{Computational Experiment: Visualizing Loss Function Convergence}
    \textbf{Goal of the experiment:} to assess how much the loss function changes when moving 
    from one training set size to the next:
    $$
    \Delta_k = \mathbb{E}_{p(\mathbf{w})} \bigl( \mathcal{L}_{k}(\mathbf{w}) - \mathcal{L}_{k-1}(\mathbf{w}) \bigr)^2
    $$
    where $\mathcal{L}$ is the loss function, $p(\mathbf{w})$ is a distribution, and $k$ is the dataset size index.

    \textbf{Method:} generate points according to $p(\mathbf{w})$ around the minimum point and 
    use Monte Carlo averaging.

    \begin{figure}[!htbp]
        \vspace*{-0.17cm}
        \centering
        \includegraphics[width=0.82\textwidth]{img/Dm_32.pdf}
    \end{figure}
\end{frame}

\end{document}

### 2025-03-20 code/src/visualize.py

@@ -105,8 +105,8 @@ class LossVisualizer:
                                   alpha=1.0)
         fig.colorbar(surf, shrink=0.5, aspect=5)
         ax_3d.view_init(40, 20)
-
-        # plt.savefig(f"Ld_{size1}.eps")
+        plt.tight_layout()
+        plt.savefig(f"Ld_{size1}.pdf")
         plt.show()
 
     def visualize(self, size=None,
@@ -141,8 +141,8 @@ class LossVisualizer:
                                   alpha=1.0)
         fig.colorbar(surf, shrink=0.5, aspect=5)
         ax_3d.view_init(40, 20)
-
-        # plt.savefig(f"Ls_{size}.eps")
+        plt.tight_layout()
+        plt.savefig(f"Ls_{size}.pdf")
         plt.show()
 
 
@@ -188,6 +188,7 @@ class DeltaVisualizer:
             xlabel='k',
             ylabel=ylabels[1],
         )
+        plt.tight_layout()
         # plt.savefig("2.eps")
         plt.show()
 

### 2025-03-27 code/src/calc.py
Universal commits presents...

@@ -6,9 +6,11 @@ from src.directions import create_random_direction
 from src.directions import init_from_params
 from src.directions import inplace_sum_models
 from src.utils import create_losses_func
+from src.utils import create_loss_func
 
 from tqdm.auto import tqdm
 from src.directions import calc_sum_models
+from src.directions import create_top_hessian_directions
 
 
 class LossCalculator:
@@ -33,6 +35,41 @@ class LossCalculator:
             target_model = calc_sum_models(self.model, target_add_model, 1, 1)
             losses = self.calc_losses_func(target_model)
             result[(coef1, coef2)] = losses
+        return result
+
+
+class LossEigenCalculator:
+    def __init__(self, model, loss_func):
+        self.model = model
+        self.loss_func = loss_func
+
+    def calc_losses(self, coef_grid, direction_norm):
+        """
+        Вычисляет значения loss по сетке коэффициентов (coef1, coef2)
+        вдоль направлений, приближённых к первым двум собственным векторам гессиана функции потерь.
+
+        Аргументы:
+          coef_grid: набор пар коэффициентов, например, из диапазона [-1, 1] по обеим осям.
+          direction_norm: коэффициент масштабирования направлений.
+        """
+        result = {}
+
+        loss = self.calc_loss_func(self.model)
+
+        direction1, direction2 = create_top_hessian_directions(
+            self.model, loss,
+            external_factor=direction_norm
+        )
+
+        for coef1, coef2 in tqdm(coef_grid):
+            target_add = [p1 * coef1 + p2 * coef2 for p1, p2 in zip(direction1, direction2)]
+            target_add_model = copy.deepcopy(self.model)
+            init_from_params(target_add_model, target_add)
+
+            target_model = calc_sum_models(self.model, target_add_model, 1, 1)
+            losses = self.calc_losses_func(target_model)
+            result[(coef1, coef2)] = losses
+
         return result
 
 

### 2025-03-27 code/src/directions.py

@@ -7,6 +7,11 @@ def get_weights(net):
     return [p.data for p in net.parameters()]
 
 
+def get_weights_graph(net):
+    """ Extract parameters from net, and return a list of tensors"""
+    return [p for p in net.parameters() if p.requires_grad]
+
+
 def get_random_weights(weights):
     """
         Produce a random direction that is a list of random Gaussian tensors
@@ -28,25 +33,32 @@ def normalize_direction(direction, weights, norm='filter'):
     if norm == 'filter':
         # Rescale the filters (weights in group) in 'direction' so that each
         # filter has the same norm as its corresponding filter in 'weights'.
-        for d, w in zip(direction, weights):
-            d.mul_(w.norm() / (d.norm() + 1e-10))
+        new_direction = torch.stack([
+            d_elem * (w_elem.norm() / (d_elem.norm() + 1e-10))
+            for d_elem, w_elem in zip(direction, weights)
+        ])
     elif norm == 'layer':
         # Rescale the layer variables in the direction so that each layer has
         # the same norm as the layer variables in weights.
-        direction.mul_(weights.norm() / direction.norm())
+        new_direction = direction * (weights.norm() / direction.norm())
     elif norm == 'weight':
         # Rescale the entries in the direction so that each entry has the same
         # scale as the corresponding weight.
-        direction.mul_(weights)
+        new_direction = direction * weights
     elif norm == 'dfilter':
         # Rescale the entries in the direction so that each filter direction
         # has the unit norm.
-        for d in direction:
-            d.div_(d.norm() + 1e-10)
+        new_direction = torch.stack([
+            d_elem / (d_elem.norm() + 1e-10)
+            for d_elem in direction
+        ])
     elif norm == 'dlayer':
         # Rescale the entries in the direction so that each layer direction has
         # the unit norm.
-        direction.div_(direction.norm())
+        new_direction = direction / direction.norm()
+    else:
+        new_direction = direction
+    return new_direction
 
 
 def normalize_directions_for_weights(direction, weights, norm='filter', ignore='biasbn'):
@@ -54,14 +66,16 @@ def normalize_directions_for_weights(direction, weights, norm='filter', ignore='
         The normalization scales the direction entries according to the entries of weights.
     """
[…]

### 2025-03-27 code/src/utils.py

@@ -104,3 +104,31 @@ def create_losses_func(dataloader, criterion):
         return losses
 
     return calc_losses
+
+
+def valid_loss(model, criterion, dataloader, device):
+    """
+    Вычисляет среднее значение loss по даталоадеру без отсоединения от графа вычислений.
+    """
+    model.eval()
+    total_loss = 0.0
+    count = 0
+    for x, y in dataloader:
+        x, y = x.to(device), y.to(device)
+        outputs = model(x)
+        loss = criterion(outputs, y)
+        total_loss += loss
+        count += 1
+    return total_loss / count
+
+
+def create_loss_func(dataloader, criterion):
+    """
+    Возвращает функцию, вычисляющую loss по даталоадеру для модели.
+    Предполагается, что у модели есть атрибут device.
+    """
+
+    def calc_loss(model):
+        return valid_loss(model, criterion, dataloader, model.device)
+
+    return calc_loss

### 2025-03-27 code/src/visualize.py

@@ -8,23 +8,18 @@ import scipy.stats as sps
 from tqdm.auto import tqdm
 
 from src.calc import DeltaCalculator
-from src.calc import LossCalculator
 
 
 class LossVisualizer:
-    def __init__(self, model, loader, criterion, grid_step=0.1, direction_norm=1):
+    def __init__(self, calculator, grid_step=0.1, direction_norm=1):
         """
             grid_step in [0, 1]
         """
-        self.grid_loss = None
         self.grid_step = grid_step
-        self.direction_norm = direction_norm
-
         self.coef_grid = list(itertools.product(np.arange(-1, 1 + self.grid_step, step=self.grid_step),
                                                 np.arange(-1, 1 + self.grid_step, step=self.grid_step)))
 
-        self.loss_calc = LossCalculator(model, loader, criterion)
-        self.grid_loss = self.loss_calc.calc_losses(self.coef_grid, self.direction_norm)
+        self.grid_loss = calculator.calc_losses(self.coef_grid, direction_norm)
 
     def _set_xy_grid(self, x_grid_bounds, y_grid_bounds):
         xs = np.arange(-1, 1 + self.grid_step, step=self.grid_step)
@@ -105,7 +100,7 @@ class LossVisualizer:
                                   alpha=1.0)
         fig.colorbar(surf, shrink=0.5, aspect=5)
         ax_3d.view_init(40, 20)
-        plt.tight_layout()
+
         plt.savefig(f"Ld_{size1}.pdf")
         plt.show()
 
@@ -141,7 +136,6 @@ class LossVisualizer:
                                   alpha=1.0)
         fig.colorbar(surf, shrink=0.5, aspect=5)
         ax_3d.view_init(40, 20)
-        plt.tight_layout()
         plt.savefig(f"Ls_{size}.pdf")
         plt.show()
 
@@ -159,13 +153,13 @@ class DeltaVisualizer:
 
         deltas = self.delta_calc.calc_deltas(mode, params, num_samples=num_samples)
 
-        fig, axs = plt.subplots(figsize=(9, 6), nrows=2, ncols=1)
+        fig, axs = plt.subplots(figsize=(9, 6), nrows=1, ncols=1)
         fig.suptitle(fr'$\Delta_k$; {params}')
-        axs[0].plot(deltas, label="Test")
+        axs.plot(deltas, label="Test")
[…]

### 2025-04-16 paper/main.tex
Universal commits presents...

--- 2025-03-20 paper/main.tex
@@
 \PassOptionsToPackage{numbers, sort, compress}{natbib}
-% before loading neurips_2024
+% before loading neurips_2025
 
 % ready for submission
-%\usepackage{neurips_2024}
+%\usepackage{neurips_2025}
 
-% to compile a preprint version, e.g., for submission to arXiv, add add the
-% [preprint] option:
+% to compile a preprint version, e.g., for submission to arXiv, add add the [preprint] option:
 \usepackage[preprint]{neurips_2025}
@@
 % to compile a camera-ready version, add the [final] option, e.g.:
-%     \usepackage[final]{neurips_2024}
+%\usepackage[final]{neurips_2024}
 
 % to avoid loading the natbib package, add option nonatbib:
-%    \usepackage[nonatbib]{neurips_2024}
+%\usepackage[nonatbib]{neurips_2024}
 
@@
 output $y \in \by$. We consider $\by$ is a subspace (or same space) as $\mathbb{R}^K$. Let $f_{\bw}(\cdot)$
-be a neural network with a list of parameters $\bw$. Let $\Omega$ be a space of parameters ($\bw \in \Omega$).
+be a neural network with a list of parameters $\bw$. Let $\bOmega$ be a space of parameters ($\bw \in \bOmega$).
 
@@
   The empirical loss function for the first $k$ elements is:
-  $$\mathcal{L}_k(\bw) = \frac1k \sum\limits_1^k \ell_i(\bw), \,\, \mathcal{L}(\bw) := \mathcal{L}_m(\bw).$$
+  $$\cL_k(\bw) = \frac1M \sum\limits_1^k \ell_i(\bw), \,\, \cL(\bw) := \cL_m(\bw).$$
 \end{definition}
@@
 Thus, the difference between losses for neighbouring samplem sizes is:
-$$\mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw) = \frac{\ell_{k}(\bw) - \mathcal{L}_{k-1}(\bw)}{k}.$$
+$$\cL_{k+1}(\bw) - \cL_{k}(\bw) = \frac{\ell_{k}(\bw) - \cL_{k-1}(\bw)}{k}.$$
 
 \begin{definition}
-  Let the Hessian of $\mathcal{L}_k(\bw)$ be:
-  $$\bH_k(\bw) = \nabla^2_{\bw}\mathcal{L}_k(\bw) = \frac1k \sum\limits_1^k \nabla^2_{\bw} \ell_i(\bw).$$
+  Let the Hessian of $\cL_k(\bw)$ be:
+  $$\bH_k(\bw) = \nabla^2_{\bw}\cL_k(\bw) = \frac1k \sum\limits_1^k \nabla^2_{\bw} \ell_i(\bw).$$
 \end{definition}
@@
   To calculate the overall loss landscape changing, one has to integrate the absolute difference for the entire parameter space.
-  We define \textbf{weighted difference} as:
-  $$\Delta_{k} = \int \left( \mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw) \right)^2 p(\bw) dw,$$
+  We define $\mathbf{\Delta}$-\textbf{function} (delta-function) as:
+  $$\Delta_{k} = \int \left( \cL_{k+1}(\bw) - \cL_{k}(\bw) \right)^2 p(\bw) dw,$$
   where $p(\bw)$ describes the priority of the particular parameter points so we can make $p(\bw)$ have higher values
@@
 \begin{assumption}
-  Let $\bw^*$ be the local minimum of both $\mathcal{L}_{k-1}(\bw)$ and $\mathcal{L}_{k}(\bw)$.
-  Thus, $$\nabla \mathcal{L}_{k-1}(\bw^*) = \nabla \mathcal{L}_{k}(\bw^*) = 0.$$
+  Let $\bw^*$ be the local minimum of both $\cL_{k-1}(\bw)$ and $\cL_{k}(\bw)$.
+  Thus, $$\nabla \cL_{k+1}(\bw^*) = \nabla \cL_{k}(\bw^*) = 0.$$
 \end{assumption}
@@
 
-Furthermore, using second-order Taylor's approximation for $\mathcal{L}_{k}(\bw)$ at $\bw^*$ we get:
-$$\mathcal{L}_{k}(\bw) \approx \mathcal{L}_{k}(\bw^*) + \frac12 (\bw - \bw^*)^{\T} \bH_k(\bw^*)
-  (\bw - \bw^*)$$
+Furthermore, using second-order Taylor's approximation for $\cL_{k}(\bw)$ at $\bw^*$ we get:
+$$\cL_{k}(\bw) \approx \cL_{k}(\bw^*) + \frac12 (\bw - \bw^*)^{\T} \bH_k(\bw^*)(\bw - \bw^*)$$
+
+\begin{assumption}
+  We can assume that $\bw^*$ is a local minimizer, so that $\cL_{k+1}(\bw^*) = \cL_k(\bw^*)$.
+\end{assumption}
 
@@
   distribution of $\bw$, so:
-  $$\Delta_{k} = \mathbb{E}_{p(\bw)} \left( \mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw)
-    \right)^2 = \mathbb{D} \left( \mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw) \right) + \left( \mathbb{E} \left(
-      \mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw) \right) \right)^2$$
+  $$\Delta_{k} = \mathbb{E}_{p(\bw)} \left( \cL_{k+1}(\bw) - \cL_{k}(\bw) \right)^2 = $$
+  $$\mathbb{D}_{p(\bw)} \left( \cL_{k+1}(\bw) - \cL_{k}(\bw) \right) + \left( \mathbb{E}_{p(\bw)} \left(
+      \cL_{k+1}(\bw) - \cL_{k}(\bw) \right) \right)^2$$
 \end{assumption}
 
-\section{Method}\label{sec:method}
+\section{Method: Projection onto Dominant Eigen-Directions and Loss Landscape Approximation}
 
-As the work with the full Hessian is computation complex, let's define a projection of neural network parameters space
-$\Omega$ to a certain subspace $\Theta$. Let $\bP$ be a projection matrix from $\Omega$ to $\Theta$. Let $\btheta$ be
-a projection: $\btheta = \bP\bw$.
+Working with the full Hessian $\mathbf{H}_k(\bw)$ of a neural network is computationally expensive due to the high dimensionality
+of the parameter space $\bOmega$. To overcome this issue, we reduce the complexity by projecting the parameters onto a lower-dimensional
+subspace. This subspace is spanned by the $d$ dominant eigenvectors of the Hessian evaluated at the local minimum $\bw^*$, i.e.,
+$\mathbf{H}_k(\bw^*)$.
 
-\section{Experiments}\label{sec:exp}
+The core idea behind the projection is that, in neural network models, the Hessian has a low effective rank according to
+\cite{sagun2018empirical}: only a few eigenvalues are significantly larger than zero, while the majority are negligibly small.
+As a result, the directions corresponding to these small eigenvalues contribute little to the variation of the loss
+function. By keeping only the eigen-directions associated with the largest eigenvalues, we obtain a compact but informative
+representation of the loss landscape that captures the most influential directions of curvature.
 
-To verify the theoretical estimates obtained, we conducted a detailed empirical study...
+For further intuition, we present an experimental result on the plot \ref{fig:evgen} showing the dependence of the eigenvalues
+on their index (ordered in descending order). The dominant eigenvalues were computed iteratively using a method described in
+\cite{hessian-eigenthings}.
+
+\begin{figure}[!htbp]
+  \centering
+  \subfloat{\includegraphics[width=0.74\textwidth]{img/eigenvalues.pdf}}
+  \caption{Dependence of eigenvalues on their index in descending order.}
+  \label{fig:evgen}
+\end{figure}
+
+Concretely, let
+$$\mathbf{V} = \left\{ \mathbf{e}_1, \dots, \mathbf{e}_d \right\}, \qquad \bP = \left[ \mathbf{e}_1, \dots, \mathbf{e}_d \right],$$
+be the corresponding eigenvectors (forming an eigenbasis) with the largest eigenvalues. Then, we represent the network parameters as
+$$\bw = \bw^* + \bP\,\btheta,$$ where $\btheta \in \mathbb{R}^d$ denotes the coordinates in this reduced subspace. This representation
+allows us to focus on the most significant directions, leading to a more tractable approximation of the loss landscape.
+
+Substituting $\bw = \bw^* + \bP\,\btheta$ into the second-order Taylor approximation of the loss function, we obtain
+$$\cL_k\bigl(\bw^* + \bP\,\btheta\bigr) \approx \cL_k\bigl(\bw^*\bigr) + \frac{1}{2}\,\btheta^{\T} \left( \bP^{\T}\,
+  \bH_k\bigl(\bw^*\bigr)\,\bP \right) \btheta.$$
+Define
+$$\bLambda_k(\bw) := \bP^{\T}\,\bH_k(\bw)\,\bP, \quad \text{so that} \quad \bLambda_k(\bw^*) =
+  \operatorname{diag}\bigl(\lambda_k^1,\dots,\lambda_k^d\bigr),$$
+where $\lambda_k^1, \dots, \lambda_k^d$ are the top $d$ eigenvalues of $\bH_k(\bw^*)$. Thus, the loss is approximated as
+$$\cL_k\bigl(\bw^* + \bP\,\btheta\bigr) \approx \cL_k\bigl(\bw^*\bigr) + \frac{1}{2}\,\btheta^{\T} \bLambda_k \btheta.$$
+
+Similarly, for the loss computed on $k+1$ samples, we have
+$$\cL_{k+1}\bigl(\bw^* + \bP\,\btheta\bigr) \approx \cL_{k+1}\bigl(\bw^*\bigr) + \frac{1}{2}\,\btheta^{\T} \bLambda_{k+1} \btheta.$$
+
+\begin{theorem}
+  (Of Delta-function) Let $p(\btheta)$ be a probability distribution over the reduced subspace (for instance, a Gaussian distribution), with mean
+  $\bm = \mathbf{0}$ and covariance $\bSigma = \mathbb{D}_{p(\btheta)}\bigl(\btheta\bigr).$ Then,
+  $$\Delta_k \;\approx\; \frac{\sigma^4}{4} \left( 2 \sum_{i=1}^{d}\left( \lambda_{k+1}^i - \lambda_{k}^i \right)^2 + \left(
+      \sum_{i=1}^{d}\left( \lambda_{k+1}^i - \lambda_{k}^i \right) \right)^2 \right).$$
+\end{theorem}
+
+Assuming $\mathbb{E}_{p(\btheta)}(\btheta) = \mathbf{0}$ is natural when centering the sampling distribution at the minimum.
+
+The result shown on this theorem and applied to our $p(\btheta)$ can estimate the $\Delta_k$ function so we can seay when the dataset
+size will be sufficient.
+
+\section{Experiments}
+To verify the theoretical estimates, in this section, we present experiments aimed at analyzing how the loss landscape evolves as
+the size of the training set increases.
+
+\paragraph{Data.}
+We use the MNIST \cite{deng2012mnist} and Fashion-MNIST \cite{xiao2017fashion} dataset, which consists of $60{,}000$
+training images and $10{,}000$ test
+images of handwritten digits (0--9). Each image is $28\times28$ pixels, grayscale.
+
+\paragraph{Model Architecture.}
+We use the following neural network model structures and fixations:
+\begin{itemize}
+  \item A multilayer perceptron (MLP) with different number of hidden layers (ReLU activations).
+  \item The choice of architecture remains consistent across all experiments.
+\end{itemize}
+
+The first two are preliminary experiments, conducted as illustrative examples on the MNIST dataset,
+while the others focusing on Hessian-based projections and show connections between eigenvalues and loss function convergence.
+That is our primary aim --- to check our theoretical estimates in-deal. All experiments are avaliable at our github repository.
+
+\subsection{Visualizing the Loss Landscape in a Random Two-Dimensional Subspace}
+The idea of this initial experiment is to get an intuitive sense of how the loss function’s landscape shifts with
+different training-set sizes. To do this, we pick two random directions $\mathbf{v}_1$ and $\mathbf{v}_2$ in the parameter
+space and project the network’s parameters onto the subspace they span.
+
+In order to visualize how the loss function behaves as the model parameters change along two chosen random directions, we compute loss
+values over a two-dimensional grid in the projected parameter space. Recall that, in our setup, the loss function is defined as
+$$L_k(w)=\frac{1}{k}\sum_{i=1}^{k}\ell_i(w),$$
+where $\ell_i(w)$ is the loss contribution of the $i$th sample.
+
+The procedure for computing the loss values and visualizing them in 3D is as follows:
+
+\begin{enumerate}
+  \item \textbf{Fix a base model and directions.}
+        Begin with a fixed trained on the same dataset neural network model.
+        Also, obtain two parameter directions $\btheta = \{\bv_1$, $\bv_2$\} that will be used to explore the parameter space.
+        These directions are vectors in the model's parameter space.
+
+  \item \textbf{Grid Construction.}
+        We generate a two-dimensional grid of coefficient pairs $(\alpha,\beta)$ using a fixed step size within the interval $[-1,1]$.
+        These coefficients will scale two pre-computed directions in the parameter space. The grid is formed by taking the Cartesian product.
+
+  \item \textbf{Loss Calculation on the Grid.}
+        For each grid point $(\alpha,\beta)$, a new set of parameters is created by forming a linear combination of the two directions:
+        $$
+          \btheta = \alpha\cdot\bv_1 + \beta\cdot\bv_2.
+        $$
+        A temporary model is then initialized with these shifted parameters, and combined with the base model via straight sum:
+        $$\bw = \bw^* + \btheta$$.
+        The base model's parameters are kept fixed (close to $\mathbf{w}^*$) while applying these perturbations.
+        Using a loss function evaluation, the losses are computed for each grid point.
+
+  \item \textbf{Computation of Aggregated Loss Measures.}
+        The loss values for each grid point are aggregated in two ways:
+        \begin{itemize}
+          \item The mean loss $\cL_k$ computed over a specified number of samples is used to construct the loss landscape.
+          \item Similarly, the mean loss $\cL_{k+1}$ for a larger sample size is computed, and the squared difference
+                $$\left(\mathcal{L}_{\text{k+1}} - \mathcal{L}_{\text{k}}\right)^2$$
+                is evaluated at each grid point.
+        \end{itemize}
+
+  \item \textbf{3D Surface Visualization.}
+        Two 3D surface plots are created and placed side by side:
+        \begin{itemize}
+          \item \emph{Left Plot:} Displays the loss landscape $\mathcal{L}_{k}(\btheta)$, where the $x$ and $y$ axes correspond to the grid
+                coordinates (the scaling coefficients along the two directions) and the $z$-axis shows the average loss value.
+          \item \emph{Right Plot:} Shows the squared loss difference, $\left(\mathcal{L}_{k+1}(\btheta)-\mathcal{L}_{k}(\btheta)\right)^2$.
+                This plot emphasizes regions where an additional data sample induces substantial changes in the loss.
+        \end{itemize}
+\end{enumerate}
+
+\begin{figure}[!htbp]
+  \hspace*{-2.2cm}
+  \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_random_1_2.pdf}}
+  \caption{Loss function landscape for small $k$}
+  \label{fig:loss_random_small}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \hspace*{-2.2cm}
+  \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_random_-2_-1.pdf}}
+  \caption{Loss function landscape for maximum $k$}
+  \label{fig:loss_random_big}
+\end{figure}
+
+Figures \ref{fig:loss_random_small} and \ref{fig:loss_random_big} shows a set of four 3D plots that illustrate the loss
+landscape and the corresponding loss difference for various sample sizes. In these figures, the top two plots correspond to very small
+training sets, while the bottom two plots display the loss landscape and its squared difference for the maximum available dataset size.
+
+\paragraph{Loss Landscape for Small Sample Sizes}
+In the upper plots \ref{fig:loss_random_small} ($k=1$), the loss landscape exhibits sharp variations and high-contrast peaks.
+These characteristics indicate that when training with an extremely limited number of samples, the model is highly sensitive to small
+perturbations in the parameter space. In such cases, even a slight change in parameters leads to significant fluctuations in the loss
+function. This is further supported by the upper plot of the squared loss difference,
+$$\left(\cL_k(\btheta)-\cL_{k}(\btheta)\right)^2,$$
+which shows large local differences that reflect high instability in the optimization landscape.
+
+\paragraph{Loss Landscape for Maximum Dataset Size}
+In contrast, the lower plots \ref{fig:loss_random_big} correspond to training with the maximum available dataset size. Here, the loss
+surface appears much smoother and more regular. The sharp peaks have diminished considerably, and the surface displays a gradual slope.
+This smoothing of the loss landscape indicates that as more data become available, the model's local optimum stabilizes. Likewise, the
+squared loss difference in the lower plot is much smaller, implying that the addition of extra samples produces only minor adjustments
+to the loss function. Such convergence suggests that the network has captured the underlying data structure adequately, and that further
+increases in the training sample size yield diminishing returns in terms of reshaping the loss landscape.
+
+\subsection{Visualizing delta-function}
+
+To further analyze the convergence of the loss landscape, we also evaluate and visualize the $\Delta_k$ function. This function 
+quantifies the change in the loss function when transitioning from a model trained on $k$ samples to one trained on $k+1$ samples. 
+Directly computing the expectation
+$$\Delta_k = \mathbb{E}_{p(\bw)}\left( \cL_{k+1}(\bw) - \cL_{k}(\bw) \right)^2$$
+over the entire distribution of parameter shifts is computationally infeasible. Therefore, we employ a Monte Carlo approximation, 
+where we average the loss differences computed on a set of random Gaussian perturbations. This Monte Carlo approach is straightforward 
+to implement and does not require explicit knowledge of the full distribution of parameter shifts.
+
+The overall procedure is as follows:
+
+\begin{enumerate}
+  \item \textbf{Fixing a model and sampling Directions.}
+        Begin with a fixed trained on the same dataset neural network model. Generate $K$ independent random directions from a
+        multivariate normal distribution
+        $$\btheta \sim \mathcal{N}(\bm, \sigma^2 \bI),$$
+        where $\bm = \mathbf{0}$ and $\sigma$ as variance. These $K$ samples serve as the shifts in the reduced parameter space.
+
+  \item \textbf{Parameter Snapshot.}\\
+        For each random direction, form a perturbed parameter vector around the base parameter $\mathbf{w}^*$:
+        $$\mathbf{w}^{(t)} = \mathbf{w}^* + \btheta^{(t)}, \quad t=1,\dots,K.$$
+        The base model's parameters are kept fixed (close to $\mathbf{w}^*$) while applying these perturbations.
+
+  \item \textbf{Loss Evaluation.}\\
+        For each perturbed parameter $\mathbf{w}^{(t)}$, compute the loss values
+        $$\cL_{k+1}(\mathbf{w}^{(t)}) \quad \text{and} \quad \cL_{k}(\mathbf{w}^{(t)}),$$
+        and then calculate the squared loss difference
+        $$\left( \cL_{k+1}(\mathbf{w}^{(t)}) - \cL_{k}(\mathbf{w}^{(t)}) \right)^2.$$
+
+  \item \textbf{Averaging.}\\
+        Average the squared loss differences over the $K$ samples to obtain the Monte Carlo estimate:
+        $$\Delta_k \approx \frac{1}{K} \sum_{t=1}^{K} \left( \cL_{k+1}(\mathbf{w}^{(t)}) - \cL_{k}(\mathbf{w}^{(t)}) \right)^2.$$
+        Increasing $K$ results in a more reliable estimate of the expected loss difference.
+
+  \item \textbf{Visualization.}\\
+        The computed $\Delta_k$ values are then visualized as follows:
+        \begin{itemize}
+          \item The left plot displays the evolution of $\Delta_k$ across different sample sizes. The $x$-axis represents the sample index $k$, while the $y$-axis shows the corresponding $\Delta_k$.
+          \item The right plot shows the product $\Delta_k \cdot k^2$ (or another suitable scaling), which highlights the convergence rate of the loss differences as the dataset size increases.
+        \end{itemize}
+        These 2D plots are generated side-by-side to facilitate a comparative analysis of the convergence behavior.
+
+\end{enumerate}
+
+\begin{figure}[!htbp]
+  \hspace*{-2.8cm}
+  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_random_sigma_2_64.pdf}}
+  \caption{Delta function landscape for different variation}
+  \label{fig:delta_random_sigma}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \hspace*{-2.8cm}
+  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_random_dim_1_64.pdf}}
+  \caption{Delta function landscape for different number of dimensions}
+  \label{fig:delta_random_dim}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \hspace*{-2.8cm}
+  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_random_num_samples_1_2.pdf}}
+  \caption{Delta function landscape for different number of samples ($K$)}
+  \label{fig:delta_random_samples}
+\end{figure}
+
+Figures \ref{fig:delta_random_sigma}, \ref{fig:delta_random_dim}, and \ref{fig:delta_random_samples} present the behavior of the function
+$$\Delta_k \;=\; \mathbb{E}_{p(\bw)}\bigl[L_{k+1}(\bw) - L_k(\bw)\bigr]^2$$
+under varying parameters of the Monte Carlo sampling and the subspace dimension used for projection. Each figure focuses on a 
+specific comparison: \ref{fig:delta_random_sigma} --- different values of the Gaussian variance $\sigma$, 
+\ref{fig:delta_random_sigma} -- different sizes of the projected subspace (i.e., different dimensions), and 
+\ref{fig:delta_random_sigma} different numbers of random samples~$K$ in the Monte Carlo algorithm.
+
+\paragraph{Comparing Different Variances \texorpdfstring{$\sigma$}{sigma}}
+Figure~\ref{fig:delta_random_sigma} shows how the $\Delta_k$ curves evolve when using various values of $\sigma$ for the random 
+directions. A larger $\sigma$ increases the magnitude and frequency of spikes in the curves. Intuitively, when the Gaussian sampling 
+is more spread out (i.e., we sample directions farther from the base parameter $\bw^*$), the resulting perturbations can induce more 
+significant changes in the loss. Thus, $\Delta_k$ can become highly volatile. By contrast, a smaller $\sigma$ keeps the sampled points 
+closer to the minimum, making the changes in loss more localized and stable. In practice, choosing too large a $\sigma$ can obscure the 
+structure around $\bw^*$ with excessive noise, while a smaller $\sigma$ focuses on the region near the local minimum, highlighting 
+more relevant variations in the loss landscape.
+
+\paragraph{Comparing Different Subspace Dimensions}
+Figure~\ref{fig:delta_random_dim} compares the $\Delta_k$ function for different projection dimensions. 
+As the dimension of the subspace grows, the curves tend to exhibit fewer extreme spikes. This suggests that by considering more 
+directions in the parameter space, the Monte Carlo estimation captures a broader view of potential perturbations. Consequently, 
+individual outliers affect the overall measure less dramatically, leading to more stable curves. 
+But a plenty of directions have negligible eigenvalues and thus minimal impact on the loss, so this does not necessarily 
+mean we should always use a high-dimensional subspace. The idea of focusing on the dominant directions 
+(with the largest eigenvalues) helps strike a balance between computational feasibility and capturing the most significant variations 
+in the loss function.
+
+\paragraph{Comparing Different Numbers of Random Samples \texorpdfstring{$K$}{K}}
+Figure~\ref{fig:delta_random_samples} illustrates the effect of varying $K$, the number of random Gaussian samples used for Monte Carlo 
+estimation. With a larger $K$, the curves become less noisy and more stable. This is because the average
+$$\frac{1}{K} \sum_{t=1}^{K} \Bigl(\cL_{k+1}(\bw^{(t)}) - \cL_k(\bw^{(t)})\Bigr)^2$$
+is taken over a greater number of directions, thereby reducing variance in the estimate.
+
+In addition, the right-side plots in these figures show the scaled quantity $\Delta_k \cdot k^2$, revealing an approximate $1/k$ decay 
+for the unscaled $\Delta_k$. This suggests that the incremental change in the loss function diminishes on the order of $1/k$ as $k$ 
+increases, confirming that once a sufficiently large dataset is reached, additional samples have only a minor impact on the local loss 
+landscape.
+
+\noindent
+\textbf{Conclusion from Preliminary Experiments.}
+These initial experiments demonstrate that the loss landscape becomes progressively more stable with increasing $N$.
+However, the directions used here are either random. This motivates a more rigorous approach in the \emph{main experiment}
+below, where we focus specifically on the most critical directions of curvature --- those corresponding to the top
+eigenvalues of the Hessian.
+
+\subsection{Main Experiment: Projection onto Dominant Hessian Eigenvectors}
+
+From prior work \cite{sagun2018empirical}, we know that only a few Hessian eigenvectors (those associated with the largest
+eigenvalues) tend to capture the most significant curvature directions. By projecting onto these principal directions,
+we can more precisely measure how the landscape the loss surface changes as $k$ increases.
+
+\paragraph{Procedure}
+\begin{enumerate}
+  \item \textbf{Hessian approximation.} For each trained model, approximate
+        the Hessian $\mathbf{H}$ or its dominant eigenvalues/eigenvectors. Techniques ???
+  \item \textbf{Eigen-decomposition.} Identify the top $d$ eigenvectors $\{\mathbf{e}_1, \dots, \mathbf{e}_d\}$ with the
+        largest eigenvalues.
+  \item \textbf{Calculation and analization.} For each model, calculate a $\Delta$ fuction via the same way as in the second
+        experiment (using Monte-Carlo) and visualize via plots.
+\end{enumerate}
+
+\paragraph{Goal and Expected Outcome}
+\begin{itemize}
+  \item By focusing on the directions of largest curvature, we expect a clearer
+        measure of how the landscape of the loss surface changes with additional data.
+  \item This approach should yield a more accurate estimate of the point at which the
+        landscape effectively stabilizes, thus helping to pinpoint a minimum viable dataset size.
+  \item By applying experiments we check our calculations above on the reality.
+\end{itemize}
 
@@
 \appendix
-\section{Appendix / supplemental material}\label{app}
+\section{Appendix}
 
-\subsection{Additional experiments / Proofs of Theorems}\label{app:exp}
+\subsection{Proof of the Theorem 1}
 
-TADAAA!
+\begin{proof}
+  The difference between the two losses is given by
+  $$\cL_{k+1}(\bw) - \cL_k(\bw) \approx \Bigl[\cL_{k+1}(\bw^*) - \cL_k(\bw^*)\Bigr] + \frac{1}{2}\,\btheta^{\T}
+    \Bigl( \bLambda_{k+1} - \bLambda_k \Bigr) \btheta.$$
+
+  Now, we apply a second assumption regarding the $\Delta_k$ function in the context of the projection. Let $p(\btheta)$ be a probability
+  distribution over the reduced subspace (for instance, a Gaussian distribution), with mean
+  $$\bm = \mathbb{E}_{p(\btheta)}\bigl(\btheta\bigr)$$ and covariance $$\bSigma = \mathbb{D}_{p(\btheta)}\bigl(\btheta\bigr).$$
+  Then, following \cite{petersen2012matrix} (see page 35), one can show that
+  $$\mathbb{E}_{p(\btheta)} \Bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \Bigr] \approx \Bigl[ \cL_{k+1}(\bw^*) - \cL_k(\bw^*)\Bigr] +
+    \frac{1}{2} \Bigl( \mathrm{Tr}(\bA\,\bSigma) + \bm^{\T}\,\bA\,\bm \Bigr),$$ where $$\bA := \bLambda_{k+1}(\bw^*) - \bLambda_k(\bw^*).$$
+  Note that $\mathrm{Tr}(\bLambda_k) = \sum_{i=1}^{d} \lambda_k^i$ and, consequently,
+  $$\mathrm{Tr}(\bA) = \sum_{i=1}^{d} \lambda_{k+1}^i - \sum_{i=1}^{d} \lambda_k^i.$$
+
+  Assume further that the distribution in the reduced space is given by
+  $$\btheta \sim \mathcal{N}\bigl(\bm,\sigma^2 \bI\bigr),$$
+  so that $\bSigma = \sigma^2 \bI$. Then, according to \cite{petersen2012matrix} (page 43), the variance of the loss difference is
+  $$\mathbb{D}_{p(\btheta)}\Bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \Bigr] =\frac{1}{4} \left( 2 \sigma^4\,\mathrm{Tr}(\bA^2) +
+    4 \sigma^2\,\bm^{\T}\,\bA^2\,\bm \right).$$
+
+  Finally, the overall measure $\Delta_k$ is defined as
+  $$\Delta_{k} \;=\; \mathbb{D}_{p(\btheta)}\Bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \Bigr] + \left( \mathbb{E}_{p(\btheta)}
+    \Bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \Bigr] \right)^{2}\,.$$
+  Substituting the above expressions, we have
+  $$\Delta_k \;\approx\; \frac{1}{4}\left( 2 \sigma^4\,\mathrm{Tr}(\bA^2) + 4 \sigma^2\,\bm^{\T}\,\bA^2\,\bm \right) + \Bigl(\cL_{k+1}(\bw^*) -
+    \cL_k(\bw^*)\Bigr)^2$$ $$\quad\quad +\; 2\Bigl(\cL_{k+1}(\bw^*) - \cL_k(\bw^*)\Bigr)\cdot \frac{1}{2}\Bigl( \mathrm{Tr}(\bA\,\bSigma) +
+    \bm^{\T}\,\bA\,\bm \Bigr) +\; \frac{1}{4}\Bigl( \mathrm{Tr}(\bA\,\bSigma) + \bm^{\T}\,\bA\,\bm \Bigr)^{2}\,.$$
+
+  Assuming that $\bw^*$ is a local minimizer and remembering that $\bm = \mathbf{0}$, the above expression simplifies to
+  $$\Delta_k \;\approx\; \frac{1}{4}\left( 2 \sigma^4\,\mathrm{Tr}(\bA^2) \right) + \frac{1}{4}\Bigl( \mathrm{Tr}(\bA\,\bSigma) \Bigr)^2\,.$$
+  Since $\bSigma = \sigma^2 \bI$, it follows that
+  $$\mathrm{Tr}(\bA\,\bSigma) = \sigma^2\,\mathrm{Tr}(\bA),$$
+  and therefore,
+  $$\Delta_k \;\approx\; \frac{\sigma^4}{4} \left( 2\,\mathrm{Tr}(\bA^2) + \Bigl(\mathrm{Tr}(\bA)\Bigr)^2 \right).$$
+  Recalling that $$\mathrm{Tr}(\bA) = \sum\limits_{i=1}^{d}\left( \lambda_{k+1}^i - \lambda_{k}^i \right)$$ and
+  $$\mathrm{Tr}(\bA^2) = \sum_{i=1}^{d}\left( \lambda_{k+1}^i - \lambda_{k}^i \right)^2,$$
+  we obtain
+  $$\Delta_k \;\approx\; \frac{\sigma^4}{4} \left( 2 \sum_{i=1}^{d}\left( \lambda_{k+1}^i - \lambda_{k}^i \right)^2 + \left(
+      \sum_{i=1}^{d}\left( \lambda_{k+1}^i - \lambda_{k}^i \right) \right)^2 \right).$$
+\end{proof}
 

### 2025-04-16 slides/InnPrac/Reviews/Alan.tex
Universal commits presents...

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage{amsmath,amssymb}

\begin{document}

% ==============================
% Заголовок и идентификация отзыва
% ==============================
\begin{center}
    {\Large \textbf{Отзыв на доклад}}\\[1em]
    {\large \textbf{Название доклада:} Оценки чисел Рамсея для произвольных последовательностей графов}\\[0.5em]
    {\large \textbf{Автор доклада:} Беремкулов Алан}\\[0.5em]
    {\large \textbf{Автор отзыва:} Никитин Артем Анатольевич}\\[2em]
\end{center}

% ==============================
% Секция 1: Аннотация
% ==============================
% Требования к аннотации:
% - Объём: от 70 до 150 слов (примерно 450–1050 символов без пробелов).
% - Текст должен состоять из нескольких предложений.
% Необходимые элементы (при наличии в докладе):
%   * Формулировка задачи.
%   * Научная мотивация (зачем заниматься данной задачей).
%   * Обзор предыдущих результатов.
%   * Перспективы: какие результаты ожидается достичь.
%   * Описание предложенных методов и наличие промежуточных результатов.
\section*{Аннотация}
Доклад посвящён исследованию чисел Рамсея для произвольных последовательностей графов. Проблема заключаетсяв том, что
классическая теория Рамсея остаётся слабо изученной, и точные оценки, например для R(5, 5), неизвестны. Автор вводит
новое определение чисел Рамсея для последовательностей графов, в рамках которого классические числа представляют тривиальный
частный случай. В работе приводятся связанные результаты, в частности оценки, полученные в исследованиях А.М. Райгородского,
а также обсуждаются точные и асимптотические оценки для отдельных классов графов. Ожидается, что дальнейшее применение
предложенного подхода позволит найти числа $R_{\min}$ для графов Джонсона и обнаружить рекурсивные зависимости,
не зависящие от числа независимости графа.

% ==============================
% Секция 2: Критический комментарий
% ==============================
% Требования к критическому комментарию:
% - Объём: от 50 до 150 слов (примерно 350–1050 символов без пробелов).
% - Текст должен состоять из нескольких предложений.
% Обсудите сильные и слабые стороны доклада с точки зрения структуры, презентации и научного содержания.
\section*{Комментарий}
Доклад представляет исследование с конкретной формулировкой проблемы и предложением нового определения чисел Рамсея
для последовательностей графов. Хочу отметить логичность связи с ранее опубликованными оценками, что подчёркивает 
значимость проблемы. Докладчик на семинаре ясно изложил области применения его работы, особенно
в части обоснования необходимости расширения классического понятия чисел Рамсея. Единственным недостатком можно отметить
недостаток подробного обсуждения примения предложенных идеи на графах без дополнительных ограничений. Работа выглядит 
перспективной и актуальной для дальнейшего развития теории Рамсея в дискретной математике, и для меня была интересной в силу 
изучения изложенной теории на курсе Дискретной математики, на котором я решал связанные с данной проблемой задачи.

\end{document}

### 2025-04-16 slides/InnPrac/Reviews/Eldar.tex
Universal commits presents...

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage{amsmath,amssymb}

\begin{document}

% ==============================
% Заголовок и идентификация отзыва
% ==============================
\begin{center}
    {\Large \textbf{Отзыв на доклад}}\\[1em]
    {\large \textbf{Название доклада:} Анализ воспроизводимости значимости статических характеристик водосборов при моделировании речного стока и уровня воды методами машинного обучения}\\[0.5em]    {\large \textbf{Авторы доклада:} Хузин Эльдар}\\[0.5em]
    {\large \textbf{Автор отзыва:} Никитин Артем Анатольевич}\\[2em]
\end{center}

% ==============================
% Секция 1: Аннотация
% ==============================
% Краткое изложение основных идей доклада:
% - Формулировка задачи: оптимизация прогнозирования гидрологических процессов с учётом ограниченности данных и вычислительных ресурсов.
% - Научная мотивация: необходимость точного и экономически эффективного моделирования водных ресурсов для предупреждения наводнений и повышения управления водными объектами.
% - Обзор существующих решений и их недостатков: физико-обоснованные модели требуют огромных ресурсов, а data-driven методы зачастую не используют дополнительные физико-географические признаки.
% - Предлагаемое решение: сравнительный анализ различных data-driven подходов (Random Forest, LSTM, трансформеры) с включением физико-географических атрибутов, оценка моделей по метрикам RMSE, MAE, MAPE, NSE.
% - Ожидаемые результаты: выявление оптимальной комбинации модели и набора признаков, позволяющей снизить вычислительные затраты и повысить точность прогнозов.
\section*{Аннотация}
Доклад посвящён сравнительному анализу data-driven подходов для прогнозирования речного стока и уровня воды. Основная задача ---
определить, в какой мере использование дополнительных данных, включая физико-географические атрибуты, способствует улучшению
метрик моделей при ограниченных вычислительных ресурсах. Автор рассматривает
современные нейросетевые архитектуры (LSTM, трансформеры) в контексте их способности учитывать особенности гидрологических
процессов. Предлагаемый подход основан на оценке моделей по различным метрикам, с целью выявления оптимальной,
которая бы сочетала высокую точность прогнозов с разумными вычислительными затратами. Результатом
исследования станет практическая рекомендация по выбору модели для гидрологического прогнозирования.

% ==============================
% Секция 2: Критический комментарий
% ==============================
% Оцените сильные и слабые стороны доклада:
% - Сильные стороны: комплексный анализ, чётко сформулированная задача, использование разнообразных моделей и метрик, практическая направленность.
% - Недостатки: отдельные разделы, касающиеся описания экспериментов, могли бы быть дополнены подробными результатами, что способствовало бы лучшему пониманию практической ценности подхода.
% - Личный интерес: тема актуальна для областей, связанных с управлением водными ресурсами и предотвращением стихийных бедствий, что вызывает особый интерес благодаря практическому применению.
\section*{Комментарий}
В работе представлен подход к решению задачи гидрологического прогнозирования, что особенно важно для оптимизации
использования данных и вычислительных ресурсов. Преимуществом работы я бы отметил сравнительный анализ различных моделей,
а также интеграцию физико-географических признаков для повышения точности прогнозов. Автор использует различные
метрики для оценки качества на реальных данных, что позволяет увидеть практическую пользу подхода. Однако хотелось бы получить
более детальное сравнение с существующими методами и описание области применения в гидрологии: не до конца ясна вычислительная
затратность метода. Меня тема привлекает практической значимостью для предотвращения наводнений, что подчеркнуло для меня
актуальность предложенного исследования.

\end{document}

### 2025-04-16 slides/InnPrac/Reviews/Valeriy.tex
Universal commits presents...

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage{amsmath,amssymb}

\begin{document}

% ==============================
% Заголовок и идентификация отзыва
% ==============================
\begin{center}
    {\Large \textbf{Отзыв на доклад}}\\[1em]
    {\large \textbf{Название доклада:} Использование методов подсчёта неопределённости для борьбы с атаками на детекторы машинно-сгенерированного текста}\\[0.5em]
    {\large \textbf{Автор доклада:} Леванов Валерий}\\[0.5em]
    {\large \textbf{Автор отзыва:} Никитин Артем Анатольевич}\\[2em]
\end{center}

% ==============================
% Секция 1: Аннотация
% ==============================
% Требования к аннотации:
% - Объём: от 70 до 150 слов (примерно 450–1050 символов без пробелов).
% - Несколько предложений, раскрывающих суть доклада:
%   * Формулировка задачи.
%   * Научная мотивация.
%   * Обзор предыдущих решений/результатов.
%   * Перспективы: ожидаемые итоги.
%   * Применяемые методы и наличие промежуточных результатов (если есть).
\section*{Аннотация}
В докладе рассматривается необходимость повышения надёжности детекторов, выявляющих машинно-сгенерированный текст.
Автор отмечает, что многие существующие детекторы уязвимы к таким различным атакам, основанным на генерации. В качестве
возможного решения предлагается использовать методы подсчёта неопределённости, которые ранее
показывали хорошие результаты в других задачах обработки естественного языка, но мало исследовались в контексте
классификации "Human vs Machine Generated". Предполагается применение White-box и Black-box подходов к оценке уверенности
и обучение нейронных сетей, устойчивых к манипуляциям с текстом. По итогам исследования ожидается разработка робастного
AI-детектора, способного надёжно отличать рукописные тексты от сгенерированных.

% ==============================
% Секция 2: Критический комментарий
% ==============================
% Требования к критическому комментарию:
% - Объём: от 50 до 150 слов (примерно 350–1050 символов без пробелов).
% - Несколько предложений, учитывающих сильные/слабые стороны.
% - Чем тема интересна лично вам.
\section*{Комментарий}
В докладе чётко сформулирована проблема: многие детекторы, созданные для определения машинной генерации текста, не справляются с
адаптированными атаками. Методы подсчёта неопределённости, на которых был сделан акцент в докладе на семинаре, помогают
оценивать уверенность модели, что может повысить робастность к манипуляциям. Отмечу отдельно, что презентация содержит обзор
смежных решений в NLP и освещает потенциально полезный опыт из компьютерного зрения, что для меня подчеркнуло актуальность 
исследований и наличие существенных продвижений на данный момент. Во время презентации хотелось бы видеть более
детализированный план по экспериментальному сравнению разных методов Uncertainty Estimation. Работа является актуальной в 
текущих реалиях в области борьбы с дезинформацией и повышения академической успеваемости.

\end{document}

### 2025-04-16 code/src/calc.py

@@ -6,9 +6,6 @@ from tqdm.auto import tqdm
 
 
 def inplace_sum_models(model1, model2, coef1, coef2):
-    """
-        return model1 := model1 * coef1 + model2 * coef2
-    """
     final = model1
     for (name1, param1), (name2, param2) in zip(final.state_dict().items(), model2.state_dict().items()):
         transformed_param = param1 * coef1 + param2 * coef2
@@ -23,9 +20,6 @@ def calc_sum_models(model1, model2, coef1, coef2):
 
 
 def init_from_params(model, direction):
-    """
-        inplace init model from direction as from parameters()
-    """
     for p_orig, p_other in zip(model.parameters(), direction):
         with torch.no_grad():
             p_orig.copy_(p_other)
@@ -61,10 +55,6 @@ class LossCalculator:
         self.directions = []
 
     def calc_losses(self, grid):
-        """
-            grid: 2d from [-1, 1]*[-1, 1]
-        """
-
         if len(self.directions) < 2:
             self.directions = self.directions + self.core.get(2 - len(self.directions))
 
@@ -83,11 +73,6 @@ class LossCalculator:
 
 
 class DeltaCalculator:
-    r"""
-        Base class for calc delta
-        delta_k = (L_{k+1} - L_k)
-    """
-
     def __init__(self, model, criterion, dataloader, core):
         self.model = model
         self.core = core
@@ -110,12 +95,7 @@ class DeltaCalculator:
 
     def calc_deltas(self, mode_params, num_samples=10):
         diff_lists = self.calc_diff_lists(mode_params, num_samples)
-        if mode_params['estim_func'] == 'square':
-            diff_lists = diff_lists ** 2
-        elif mode_params['estim_func'] == 'abs':
-            diff_lists = np.abs(diff_lists)
-        elif mode_params['estim_func'] is None:
-            pass
+        diff_lists = diff_lists ** 2
         deltas = np.mean(diff_lists, axis=0)
         return deltas
 

### 2025-04-16 code/src/directions.py

@@ -1,9 +1,42 @@
+# --========================================-- #
+#   * Author  : NTheme - All rights reserved
+#   * Created : 03 April 2025, 7:09 AM
+#   * File    : directions_eigen.py
+#   * Project : code
+# --========================================-- #
+
 import torch
+from tqdm.auto import tqdm
+
+
+def get_loss_mean(model, criterion, dataloader, device):
+    model.eval()
+
+    count = 0
+    total_loss = 0.0
+    for x, y in dataloader:
+        x, y = x.to(device), y.to(device)
+
+        outputs = model(x)
+        loss = criterion(outputs, y)
+
+        total_loss += loss
+        count += 1
+
+    return total_loss / count
+
+
+def create_loss_mean_func(criterion, dataloader):
+    def calc_loss(model):
+        return get_loss_mean(model, criterion, dataloader, model.device)
+
+    return calc_loss
 
 
 class RandomDirection:
     def __init__(self, model, criterion=None, dataloader=None):
         self.model = model
+        self.type = "random"
 
     def get(self, k):
         return [self._create_random_direction(self.model) for _ in range(k)]
@@ -92,3 +125,53 @@ class RandomDirection:
             d.mul_(external_factor)
 
         return direction
+
+
+class MaxEigenvector:
+    def __init__(self, model, criterion, dataloader):
+        self.model = model
+        self.loss_func = create_loss_mean_func(criterion, dataloader)
+        self.type = "eigen"
+
+        self.eigenvectors = []
+        self.eigenvalues = []
+
+    def get(self, k, num_iters=100, tol=1e-6):
+        loss = self.loss_func(self.model)
+
+        if k <= len(self.eigenvectors):
+            return self.eigenvalues[:k], self.eigenvectors[:k]
+
+        for i in range(0, k - len(self.eigenvectors)):
+            eigenvalue, eigenvector = self._power_iteration(loss, num_iters=num_iters, tol=tol)
+            self.eigenvalues.append(eigenvalue)
+            self.eigenvectors.append(eigenvector)
+        return self.eigenvectors
+
+    def _power_iteration(self, loss, num_iters=100, tol=1e-6):
+        params = list(self.model.parameters())
+        flat_params = torch.cat([p.view(-1) for p in params])
+
+        v = torch.randn(flat_params.shape[0], device=self.model.device)
+        v = v / v.norm()
+        eigenvalue = None
[…]

### 2025-04-16 code/src/directions_eigen.py

@@ -1,82 +1,0 @@
-# --========================================-- #
-#   * Author  : NTheme - All rights reserved
-#   * Created : 03 April 2025, 7:09 AM
-#   * File    : directions_eigen.py
-#   * Project : code
-# --========================================-- #
-
-import torch
-from tqdm.auto import tqdm
-
-
-def get_loss_mean(model, criterion, dataloader, device):
-    model.eval()
-
-    count = 0
-    total_loss = 0.0
-    for x, y in dataloader:
-        x, y = x.to(device), y.to(device)
-
-        outputs = model(x)
-        loss = criterion(outputs, y)
-
-        total_loss += loss
-        count += 1
-
-    return total_loss / count
-
-
-def create_loss_mean_func(criterion, dataloader):
-    def calc_loss(model):
-        return get_loss_mean(model, criterion, dataloader, model.device)
-
-    return calc_loss
-
-
-class MaxEigenvector:
-    def __init__(self, model, criterion, dataloader):
-        self.model = model
-        self.loss_func = create_loss_mean_func(criterion, dataloader)
-
-        self.eigenvectors = []
-        self.eigenvalues = []
-
-    def get(self, k, num_iters=100, tol=1e-6):
-        loss = self.loss_func(self.model)
-
-        if k <= len(self.eigenvectors):
-            return self.eigenvalues[:k], self.eigenvectors[:k]
-
-        for i in range(0, k - len(self.eigenvectors)):
-            eigenvalue, eigenvector = self._power_iteration(loss, num_iters=num_iters, tol=tol)
-            self.eigenvalues.append(eigenvalue)
-            self.eigenvectors.append(eigenvector)
-        return self.eigenvectors
-
-    def _power_iteration(self, loss, num_iters=100, tol=1e-6):
-        params = list(self.model.parameters())
-        flat_params = torch.cat([p.view(-1) for p in params])
-
-        v = torch.randn(flat_params.shape[0], device=self.model.device)
-        v = v / v.norm()
-        eigenvalue = None
-        for _ in tqdm(range(1, num_iters + 1)):
-            hv = self._hvp(params, loss, v)
-
-            for d in self.eigenvectors:
-                hv = hv - torch.dot(hv, d) * d
-            new_v = hv / hv.norm()
-            eigenvalue = torch.dot(new_v, hv)
-            if torch.norm(new_v - v) < tol:
-                break
-            v = new_v
[…]

### 2025-04-16 code/src/models.py

@@ -75,6 +75,6 @@ def train(model, criterion, dataloader, optimizer, num_epochs=10, ret=False, log
 
         avg_loss = running_loss / batches_count
         if log:
-            print(f"Epoch [{epoch + 1}/{num_epochs}]: loss = {avg_loss:.4f}")
+            print(f"Epoch [{epoch}/{num_epochs}]: loss = {avg_loss:.4f}")
 
     return losses if ret else None

### 2025-04-16 code/src/utils.py

@@ -9,7 +9,6 @@ import torch
 from torch.utils.data import DataLoader
 from torchvision import datasets
 import numpy as np
-from tqdm.auto import tqdm
 
 
 def init_dataloader(dataset_name,

### 2025-04-16 code/src/visualize.py

@@ -15,6 +15,7 @@ class LossVisualizer:
         """
             grid_step in [0, 1]
         """
+        self.core_type = calculator.core.type
         self.grid_step = grid_step
         self.grid_loss = calculator.calc_losses(
             list(itertools.product(np.arange(-1, 1 + self.grid_step, step=self.grid_step),
@@ -27,146 +28,66 @@ class LossVisualizer:
         ys = np.array([y for y in ys if y_grid_bounds[1] >= y >= y_grid_bounds[0]])
         return xs, ys
 
-    def visualize_diff(self, size1, size2,
-                       x_grid_bounds=(-1, 1), y_grid_bounds=(-1, 1),
-                       diff_type=None, bounds=None, distrib_params=None):
-        """
-           type: None | abs | square | realative | relative_abs | relative_squar | square_dot_normal | abs_dot_normal
-           bounds: None | (-min, max)
-        """
-        assert bounds is None or bounds[0] < bounds[1]
-
+    def visualize_all(self, size1, size2,
+                      x_grid_bounds=(-1, 1), y_grid_bounds=(-1, 1), z_grid_bounds=(-float('inf'), float('inf'))):
         xs, ys = self._set_xy_grid(x_grid_bounds, y_grid_bounds)
         xgrid, ygrid = np.meshgrid(xs, ys)
+
         zgrid1 = np.array([[np.mean(self.grid_loss[(x, y)][:size1]) for x in xs] for y in ys])
         zgrid2 = np.array([[np.mean(self.grid_loss[(x, y)][:size2]) for x in xs] for y in ys])
-        zgrid = zgrid2 - zgrid1
-
-        best_loss = np.round(min(np.min(zgrid1), np.min(zgrid2)), 4)
-
-        relative = False
-        if diff_type is not None and diff_type[0:2] == 're':
-            zgrid /= zgrid2
-            relative = True
-
-        if diff_type is not None and diff_type.find('normal') != -1:
-            pdf = lambda x, y: np.prod(sps.norm(**distrib_params).pdf(np.array([x, y])))
-            zgrid *= np.array([[pdf(x, y) for x in xs] for y in ys])
-
-        title = r""
-        if diff_type is None:
-            title = r'$\mathcal{L}_{s_2} - \mathcal{L}_{s_1}$'
-            if relative:
[…]

### 2025-04-17 paper/main.tex
Universal commits presents...

--- 2025-03-20 paper/main.tex
@@
 \begin{abstract}
-  Understanding how the loss landscape of neural networks evolves as the training set size increases is crucial for
-  optimizing performance. It is well known that larger datasets can alter the shape of this high-dimensional landscape.
-  But the exact point at which additional data no longer brings substantial changes remains underexplored.
+  Understanding how a neural network’s loss landscape changes as we add more training data is important for efficient training.
+  Although larger datasets are known to reshape this high-dimensional surface, the point when extra data stop making a big difference
+  is still unclear.
 
-  In this paper, we examine neural network models and show that their loss landscapes begin to stabilize once the
-  training set grows beyond a certain threshold, revealing a connection between dataset size and the geometry of the
-  loss surface. To understand this phenomenon, we propose a method that projects the full parameter space onto a
-  low-dimensional subspace derived from top eigenvectors of the Hessian. That provides a more interpretable view of how
-  the loss surface in the vicinity of local minima behaves as more data become available. We use different sampling
-  strategies, applying Monte-Carlo estimation to capture the structure of this reduced loss landscape more precisely.
+  In this paper, we study this issue and show that the loss landscape near a local minimum stabilizes once the dataset exceeds a
+  certain size. To analyze this, we project the full parameter space onto a smaller subspace formed by the Hessian’s top eigenvectors.
+  This low-dimensional view highlights how the loss surface changes in its most important directions. We then apply Monte Carlo sampling
+  within this subspace to estimate these changes more precisely.
 
-  We validate our insights through experiments on image classification tasks, demonstrating that this low-dimensional
-  analysis can reveal when the landscape effectively settles, and thus helps determine a minimum viable dataset size.
-  Our findings shed light on the relationship between dataset scale and optimization geometry, and suggest
-  practical strategies for balancing computational costs with the benefits of additional training data.
+  We test our approach on standard image‑classification tasks and find that our low-dimensional analysis pinpoints when the landscape
+  stops evolving. These findings clarify how dataset size affects optimization and offer practical guidance for balancing training cost
+  with performance gains.
 \end{abstract}
@@
 
-Neural networks have become irreplaceable in various aspects of life and have a lot of applications now
-(e.g. image classification, language models, recommender systems, etc.). However, as datasets and models grow larger and
-larger, we get higher accuracy and breakthrough results. However some problems connected with computation resources have
-appeared. Sometimes models have so many parameters so we need time we cannot provide for their training. Some attempts to
-explore and compare optimization methods have already been made \cite{soydaner2020comparison}. The core problem of all these
-solutions hides in their locality, as the size of the neural network and training dataset remains unchanged. In this paper we
-explore when adding more data no longer brings significant improvements in network performance and to explain
-obtained results in terms of the curvature and shape of the neural network’s loss landscape.
+Neural networks are now essential in many areas --- image classification, language modeling, recommender systems, and more. As models and
+datasets grow, we often see better accuracy and new breakthroughs. But bigger networks and data also demand more computation time and
+resources, which can become prohibitive. While prior work has compared various optimization methods under fixed data and model sizes
+\cite{soydaner2020comparison}, the question of when adding more training samples stops yielding significant gains remains largely
+unanswered.
 
-Our research on the loss landscape of neural networks. Specifically, how it changes while adding new samples
-to the training dataset. There remains a problem: processing of large datasets requires a lot of time.
-We need to determine when increasing dataset size stops reshaping this loss landscape for general neural networks.
-But the question of a minimum viable dataset size --- a threshold beyond which further data bring insignificant changes ---
-remains underexplored. Moreover, connecting such a threshold to generalization capabilities \cite{wu2017towards} will
-reduce computation requirements as we show that there is an upper bound of "active" dataset samples.
+In this paper, we investigate exactly that: how the loss landscape changes as we increase the dataset size, and at what point further
+data have little effect. Knowing this “minimum viable dataset size” --- a threshold beyond which new samples bring negligible
+improvement --- can save both training time and data‑collection effort. We also relate this threshold to generalization behavior, along
+the lines of \cite{wu2017towards}.
 
-We suggest using the Hessian of the loss function to calculate its close approximation. However, the computation of all
-Hessian requires a bunch of time, so we consider a projection of neural network parameters space to a certain subspace.
-We place the following tasks to the front:
+To study this, we use the Hessian of the loss function as a proxy for local curvature. Computing the full Hessian is expensive, so we
+project the parameter space onto a low‑dimensional subspace formed by its top eigenvectors. Our main contributions are:
+
 \begin{enumerate}
-  \item Constructing a Hessian-based projection of the loss landscape which finds main curvature directions
-        (top eigenvectors) in parameter space.
-  \item Using different sampling methods, identifying at which dataset size the shape of the loss function stabilizes.
-  \item We apply visualization methods to characterize the loss landscape next to the local minima.
-  \item Providing applicable theoretical criteria to determine when adding more samples has minimal effect.
-        We refer to The Matrix Cookbook \cite{petersen2012matrix} as a hint for matrix operations.
+  \item Constructing a Hessian‑based projection that captures the principal curvature directions.
+  \item Applying Monte Carlo and other sampling methods to identify when the loss landscape stabilizes as dataset size grows.
+  \item Visualizing the projected loss surface around local minima to illustrate these changes.
+  \item Deriving theoretical criteria --- using results from The Matrix Cookbook \cite{petersen2012matrix} --- to predict when additional
+        samples have minimal impact.
 \end{enumerate}
 
-The proposed solution is novel as it links Hessian estimation to the idea of a minimal dataset threshold. This enables
-advantages such as more cost-effective dataset collection and a clearer understanding of how dataset size interacts with the
-geometry of the loss function. And applying recent methods for low-rank Hessian estimation
-can speed up and solve the problem of high computation cost of top eigenvalues for very large networks.
+This approach is novel in linking low‑rank Hessian estimation to a concrete dataset threshold, enabling more cost‑effective data
+collection and a clearer understanding of how data scale interacts with loss geometry. We validate our method on MNIST
+\cite{deng2012mnist} and Fashion‑MNIST \cite{xiao2017fashion}, demonstrating how to balance computational cost against accuracy gains.
 
-We validate our analysis results on well-known image classification tasks like MNIST \cite{deng2012mnist} and Fashion-MNIST
-\cite{xiao2017fashion}. By connecting practical results to theoretical bounds, we offer a method for identifying
-a computation-efficiency method that balances performance with accuracy gains.
-
-\textbf{Contributions.} Our contributions can be summarized as follows:
-\begin{itemize}
-  \item We present a Hessian-based approach that uses a projection into a low-dimensional subspace of top eigenvectors to
-        find the critical sufficient dataset size.
-  \item We demonstrate the validity of our theoretical framework through empirical studies on MNIST and Fashion-MNIST,
-        incrementally increasing dataset size until additional data makes an impact to the loss function curvature.
-  \item We highlight the implications of our findings for practical data collection strategies, showing how the detection of
-        a sufficient dataset size can reduce computation time.
-\end{itemize}
-
-\textbf{Outline.} The rest of the paper is organized as follows. Section 2 divides existing results into some
-topics, highlighting their key contributions and findings. Section 3 considers general notation and
-some preliminary calculations. In Section 4, we provide theoretical bounds on the hessian and losses
-difference norms. Empirical study of the obtained results is given in Section 5. We summarize and
-present the results in Sections 6 and 7. Additional experiments and proofs of theorems are included
-in the Appendix A.
+\textbf{Organization.} The rest of the paper is structured as follows. Section~2 reviews related work. Section~3 introduces notation
+and preliminaries. Section~4 derives theoretical bounds on Hessian spectra and the loss‐difference norm. Section~5 presents our
+empirical studies. We discuss the implications in Section~6 and conclude in Section~7. Additional experiments and proofs are
+provided in Appendix~A.
 
@@
 
-\textbf{Hessian-based analysis.}
-The landscape of loss functions has been explored from various perspectives in the literature. Methods connected with Hessian
-are central to understanding the convergence process and loss function landscape \cite{kiselev2024unraveling}. Recent works
-observed that the Hessian of some models tends to have a low effective rank, with only a small quantity of eigenvalue
-differences from zero. Empirical explorations from \cite{sagun2018empirical} show that the Hessian rank is often very low in
-the vicinity of local minima, highlighting a significantly smaller “active subspace”.
-Prior work investigating how large eigenvalues emerge during training \cite{ghorbani2019investigation}.
+\textbf{Hessian-based Analysis.}
+The loss landscape of neural networks has been extensively studied using Hessian-based techniques, which illuminate convergence
+behavior and curvature properties \cite{kiselev2024unraveling}. Empirical work shows that near local minima the Hessian often has low
+effective rank, indicating a small "active subspace" spanned by a few large eigenvalues \cite{sagun2018empirical}. Ghorbani et al.
+\cite{ghorbani2019investigation} examined how these large eigenvalues emerge during training and shape optimization dynamics.
+Meshkov et al. \cite{meshkov2024convnets} extended this analysis to fully connected and convolutional architectures, though broader
+validation across architectures remains an open challenge.
 
-Some results have already been reached for fully connected and convolutional architectures \cite{meshkov2024convnets},
-but they need generalization.
+\textbf{Loss Landscape Geometry and Dataset Size.}
+Another research direction explores how dataset size influences model performance and landscape geometry. Networks trained on larger
+datasets tend to find flatter, wider minima that generalize better \cite{wu2017towards}. Yet, acquiring and processing massive data is
+costly, motivating studies of optimal data–model trade‑offs \cite{hoffmann2022training}. Li et al. \cite{li2018visualizing} developed
+visualization tools that reveal how dataset size, initialization, and architecture affect the topology of loss surfaces.
 
-\textbf{Loss Landscape Geometry.}
-Other recent researches aimed at how dataset size affects both performance and the geometry of loss function landscapes.
-In the assumption of enough training samples, neural networks often reach smoother and flatter minima with better
-generalization \cite{wu2017towards}. However, collecting large datasets is expensive, as well as the computation
-requirements for them, raising questions about the most effective ratio between dataset and model sizes
-\cite{hoffmann2022training}. As understanding the landscape of loss functions is key to the solution, prior work has looked
-at visualizing loss surfaces \cite{li2018visualizing}.
+\textbf{Algorithmic Stability and Landscape Sensitivity.}
+Algorithmic stability, which measures how small changes in the training set affect the learned model, provides theoretical bounds on
+generalization error \cite{bousquet2002stability}. Elisseeff et al. \cite{elisseeff2005stability} extended stability analysis to
+randomized algorithms such as bagging and ensemble methods, deriving non-asymptotic generalization guarantees. In our context,
+stability theory underpins the intuition that, beyond a certain dataset size, further samples cause negligible changes to the loss
+curvature. Our Monte Carlo–based Hessian projection method directly leverages this perspective, using top eigenvectors to quantify
+and visualize when the loss landscape stabilizes as new data are added.
 
@@
 
-In this section, we introduce the general notation used in the rest of the paper and the basic assumptions.
-
-We consider a $K$-label classification problem.
-So, let's consider $p(y|\bx)$ a conditional probability, which maps unobserved variable $\bx \in \bX$ to the corresponding
-output $y \in \by$. We consider $\by$ is a subspace (or same space) as $\mathbb{R}^K$. Let $f_{\bw}(\cdot)$
-be a neural network with a list of parameters $\bw$. Let $\Omega$ be a space of parameters ($\bw \in \Omega$).
-
-Let
-$$\mathcal{D} = \left\{ (\bx_i, y_i) \, | \, i = 1, \dots, m \right\}$$
-be a given dataset of size $m$ consists of i.i.d.
-samples. Let $\ell(\cdot, \cdot)$ be a given loss function (e.g. cross-entrophy) where first argument refers to neural network's
-result and the second argument refers to the true answer. To simplify, define:
-$$\ell_i(\bw) := \ell(f_{\bw}(\bx_i), y_i).$$
+We consider a $K$-class classification problem. Let $\bx \in \bX$ be an input and $y \in \mathcal Y = \{1,\dots,K\}$ its label. A
+neural network with parameters $\bw \in \bOmega \subset \mathbb R^P$ defines a mapping $f_{\bw} : \bX \to \mathbb R^K$. Given a dataset
+$$
+  \mathcal D =
+  \{(\bx_i, y_i)\}_{i=1}^m
+$$
+of $m$ i.i.d. samples, and a per-sample loss $\ell \bigl( f_{\bw}(\bx_i), y_i \bigr)$ (e.g. cross–entropy), we define
+$$
+  \ell_i(\bw) =
+  \ell \bigl( f_{\bw}(\bx_i) , y_i \bigr)\,.
+$$
 
 \begin{definition}
-  The empirical loss function for the first $k$ elements is:
-  $$\mathcal{L}_k(\bw) = \frac1M \sum\limits_1^k \ell_i(\bw), \,\, \mathcal{L}(\bw) := \mathcal{L}_m(\bw).$$
+  The empirical loss on the first $k$ samples is
+  $$
+    \cL_k(\bw) =
+    \frac1k \sum_{i=1}^k \ell_i(\bw),
+  $$
+  so that the full-sample loss is $\cL(\bw) = \cL_m(\bw)$.
 \end{definition}
 
-Thus, the difference between losses for neighbouring samplem sizes is:
-$$\mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw) = \frac{\ell_{k}(\bw) - \mathcal{L}_{k-1}(\bw)}{k}.$$
+A simple telescoping gives the difference
+$$
+  \cL_{k+1}(\bw) - \cL_k(\bw) =
+  \frac{\ell_{k+1}(\bw) - \cL_k(\bw)}{k+1}\,.
+$$
 
 \begin{definition}
-  Let the Hessian of $\mathcal{L}_k(\bw)$ be:
-  $$\bH_k(\bw) = \nabla^2_{\bw}\mathcal{L}_k(\bw) = \frac1k \sum\limits_1^k \nabla^2_{\bw} \ell_i(\bw).$$
+  The Hessian of $\cL_k$ at $\bw$ is
+  $$
+    \bH_k(\bw) =
+    \nabla^2_{\bw} \cL_k(\bw) =
+    \frac1k \sum_{i=1}^k \nabla^2_{\bw} \ell_i(\bw),
+  $$
+  and, similarly, the full-sample Hessian is $\bH(\bw) = \bH_m(\bw) = \nabla^2_{\bw} \cL_m(\bw)$.
 \end{definition}
 
+To capture the overall change in the landscape when adding one more sample, we introduce a $\Delta$‑function:
+
 \begin{definition}
-  To calculate the overall loss landscape changing, one has to integrate the absolute difference for the entire parameter space.
-  We define $\mathbf{\Delta}$-\textbf{function} (delta-function) as:
-  $$\Delta_{k} = \int \left( \mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw) \right)^2 p(\bw) dw,$$
-  where $p(\bw)$ describes the priority of the particular parameter points so we can make $p(\bw)$ have higher values
-  next to the local minima.
+  Let $p(\bw)$ be a weighting (e.g. peaked near a local minimum). Define
+  $$
+    \Delta_k =
+    \int \bigl( \cL_{k+1}(\bw) - \cL_k(\bw) \bigr)^2\, p(\bw)\, d\bw.
+  $$
 \end{definition}
 
-We firther investigate this difference and aimed at exploration of how adding a new object to the dataset changes the value.
-We interested in convergence of this value and properties of loss function when the training dataset size limits to $\infty$.
+We study the behavior of $\Delta_k$ as $k \to \infty$ and define a threshold:
 
 \begin{definition}
-  Let $\Delta$ be a positive hyperparameter that indicates the stop-difference for $\Delta_k$. If
-  $$k^* = \inf\limits_k\{\forall m \geq k: \, \Delta_m < \Delta\}$$
-  we can say that $k^*$ samples in the dataset are enough to describe the distribution of data from the general population.
-  We call $k^*$ as \textbf{sufficient}.
+  Fix a tolerance $\Delta > 0$. The \emph{sufficient sample size} is
+  $$
+    k^* =
+    \inf \{ k \mid \forall\, m \ge k : \; \Delta_m < \Delta\}.
+  $$
+  When $k \ge k^*$, adding further samples changes the loss landscape by less than $\Delta$.
 \end{definition}
@@
 
+We make the following standard assumptions to simplify our analysis around a local minimizer~$\bw^*$.
+
 \begin{assumption}
-  Let $\bw^*$ be the local minimum of both $\mathcal{L}_{k-1}(\bw)$ and $\mathcal{L}_{k}(\bw)$.
-  Thus, $$\nabla \mathcal{L}_{k-1}(\bw^*) = \nabla \mathcal{L}_{k}(\bw^*) = 0.$$
+  The point $\bw^*$ is a common local minimizer for both $\cL_k$ and $\cL_{k+1}$, i.e.
+  $$
+    \nabla \cL_k(\bw^*) =
+    \nabla \cL_{k+1}(\bw^*) =
+    0.
+  $$
 \end{assumption}
 
-This assumption allows us to explore the behavior and the geometry of the loss function landscape at only one point.
-
-Furthermore, using second-order Taylor's approximation for $\mathcal{L}_{k}(\bw)$ at $\bw^*$ we get:
-$$\mathcal{L}_{k}(\bw) \approx \mathcal{L}_{k}(\bw^*) + \frac12 (\bw - \bw^*)^{\T} \bH_k(\bw^*)
-  (\bw - \bw^*)$$
+Under this assumption, a second‑order Taylor expansion of $\cL_k$ about $\bw^*$ gives
+$$
+  \cL_k(\bw) \approx
+  \cL_k(\bw^*) + \tfrac12\, (\bw - \bw^*)^{\T} \bH_k(\bw^*)\, (\bw - \bw^*).
+$$
 
 \begin{assumption}
-  We can assume parameters $\bw$ to be random, which will lead to quite natural condition: $p(\bw)$ can be even a prior
-  distribution of $\bw$, so:
-  $$\Delta_{k} = \mathbb{E}_{p(\bw)} \left( \mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw)
-    \right)^2 = \mathbb{D} \left( \mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw) \right) + \left( \mathbb{E} \left(
-      \mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw) \right) \right)^2$$
+  At the minimizer $\bw^*$, the loss values for consecutive sample sizes are the same:
+  $$
+    \cL_{k+1}(\bw^*) =
+    \cL_k(\bw^*).
+  $$
 \end{assumption}
 
-\section{Method}\label{sec:method}
+\begin{assumption}
+  The parameters $\bw$ are distributed according to a density $p(\bw)$ (e.g. a Gaussian prior centered at $\bw^*$).  Then
+  $$
+    \Delta_k =
+    \int \bigl(\cL_{k+1}(\bw)-\cL_k(\bw)\bigr)^2\, p(\bw)\, d\bw =
+    \mathbb{E}_{p(\bw)} \left( \cL_{k+1}(\bw) - \cL_{k}(\bw) \right)^2 =
+  $$
+  $$
+    \mathbb D_{p(\bw)} \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr] +
+    \Bigl( \mathbb E_{p(\bw)} \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr] \Bigr)^2.
+  $$
+\end{assumption}
 
-As the work with the full Hessian is computation complex, let's define a projection of neural network parameters space
-$\Omega$ to a certain subspace $\Theta$. Let $\bP$ be a projection matrix from $\Omega$ to $\Theta$. Let $\btheta$ be
-a projection: $\btheta = \bP\bw$.
+\section{Method: Projection onto Dominant Eigen-Directions and Loss Landscape Approximation}
 
-Let $\mathbf{V} = \{\mathbf{e}_1, \dots, \mathbf{e}_d\}$ be eigenvectors of $\mathbf{H}$. We select $d$ of them and project parameters space
-into a subspace on these $d$ vectors. These vectors should represent maximum eigenvalues of $\mathbf{H}$ as these directions
-have the most impact to the loss function difference.
-\section{Experiments}\label{sec:exp}
+Working with the full Hessian $\mathbf{H}_k(\bw)$ of a neural network is computationally expensive due to the high dimensionality
+of the parameter space $\bOmega$. To reduce this cost, we project the parameters onto a lower‑dimensional subspace spanned by the
+$d$ dominant eigenvectors of the Hessian at a local minimum $\bw^*$.
 
-To verify the theoretical estimates obtained, we conducted a detailed empirical study.
-In this section, we present experiments aimed at analyzing how the loss landscape evolves as the size of the training
-set increases.
+The justification for this projection is that, empirically, the Hessian has low effective rank near a minima \cite{sagun2018empirical}:
+only a few eigenvalues are significantly above zero, while the rest are negligible.  Therefore, directions corresponding to small
+eigenvalues contribute little to loss variation.  Retaining only the top $d$ eigenvectors yields a compact yet informative representation
+of the most important curvature directions.
 
-\paragraph{Data.}
-We use the MNIST dataset, which consists of $60{,}000$ training images and $10{,}000$ test
-images of handwritten digits (0--9). Each image is $28\times28$ pixels, grayscale.
+For illustration, Figure~\ref{fig:evgen} plots the top few eigenvalues of $\bH(\bw^*)$ in descending order. We compute these dominant
+eigenvalues using the iterative power‐method described at \cite{hessian-eigenthings}.
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=0.7\textwidth]{img/eigenvalues.pdf}
+  \caption{Top eigenvalues of $\bH(\bw^*)$ in descending order.}
+  \label{fig:evgen}
+\end{figure}
+
+Concretely, let
+$$
+  \bP =
+  \bigl[ \mathbf{e}_1, \dots, \mathbf{e}_d \bigr]
+$$
+be the matrix whose columns are the top $d$ eigenvectors.  We write any parameter vector as
+$$
+  \bw =
+  \bw^* + \bP\, \btheta, \quad
+  \btheta \in \mathbb R^d.
+$$
+Substituting into a second‑order Taylor expansion of $\cL_k$ around $\bw^*$ gives
+$$
+  \cL_k\bigl(\bw^* + \bP\,\btheta\bigr) \approx
+  \cL_k(\bw^*) + \tfrac12\, \btheta^{\T}\! \bigl( \bP^{\T}\, \bH_k(\bw^*)\, \bP\bigr)\, \btheta =
+  \cL_k(\bw^*) + \tfrac12\, \btheta^{\T} \bLambda_k\, \btheta,
+$$
+where
+$$
+  \bLambda_k =
+  \bP^{\T}\, \bH_k(\bw^*)\, \bP =
+  \operatorname{diag} \bigl( \lambda_k^1, \dots, \lambda_k^d \bigr).
+$$
+An identical expansion holds for $\cL_{k+1}$, yielding $\bLambda_{k+1}$ from the Hessian on $k+1$ samples.
+
+\begin{theorem}[Approximate $\Delta_k$ via Eigenvalues]
+  Under the assumption that $\btheta \sim \mathcal N(0, \sigma^2 \mathbf I_d)$, the change in loss can be estimated by
+  $$
+    \Delta_k =
+    \mathbb E \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr]^2 \approx
+    \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i)^2
+    + \Bigl( \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i) \Bigr)^2 \Biggr).
+  $$
+\end{theorem}
+
+\begin{remark}
+  Assuming $\mathbb{E}_{p(\btheta)}(\btheta) = \mathbf{0}$ is natural when centering the sampling distribution at the minimum.
+\end{remark}
+
+This formula uses only the top $d$ eigenvalues, so it is efficient to compute even for very large models.
+In practice, it provides an upper bound on the empirical $\Delta_k$ and thus a reliable criterion for determining when the
+dataset size is sufficient.
+
+\section{Experiments}
+
+To validate our theoretical estimates, we perform a set of experiments that track how the loss landscape changes as we grow the
+training set.
+
+\paragraph{Datasets.}
+We evaluate on MNIST \cite{deng2012mnist} and Fashion‑MNIST \cite{xiao2017fashion}, each containing 60,000 training and 10,000 test
+grayscale images of size $28\times28$ pixels.
 
 \paragraph{Model Architecture.}
-We use the following neural network models and preparations:
-\begin{itemize}
-  \item A multilayer perceptron (MLP) with different number of hidden layers (ReLU activations).
-  \item The choice of architecture remains consistent across all experiments.
-\end{itemize}
+All experiments use a simple multilayer perceptron (MLP) with ReLU activations.  We keep the same architecture—number of hidden layers
+and hidden units—across all runs to ensure a fair comparison.
 
-The first two are preliminary experiments, conducted as illustrative examples on the MNIST dataset,
-while the others focusing on Hessian-based projections and show connections between eigenvalues and loss function convergence.
-That is our primary aim --- to check our theoretical estimates in-deal.
-
-\subsection{Visualizing the Loss Landscape in a Random Two-Dimensional Subspace}
-The idea of this initial experiment is to get an intuitive sense of how the loss function’s landscape shifts with
-different training-set sizes. To do this, we pick two random directions $\mathbf{v_1}$ and $\mathbf{v_2}$ in the parameter
-space and project the network’s parameters onto the subspace they span.
-
-\paragraph{Procedure}
+\paragraph{Experimental Procedure.}
 \begin{enumerate}
-  \item \textbf{Training on different sizes.} Train a neural network on increasingly larger subsets.
-  \item \textbf{Parameter snapshot.} After training, fix the final (or near-final) weights $\mathbf{w}^*$.
-  \item \textbf{Subspace grid.} Form a grid of points $\{\mathbf{w}^* + \alpha \mathbf{v}_1 + \beta \mathbf{v}_2\}$ around $\mathbf{w}^*$, where $\alpha$ and $\beta$ vary in a chosen range.
-  \item \textbf{Loss evaluation.} Compute and plot the loss at each grid point, producing a 3D visualization.
+  \item \textbf{Random‐direction visualization:} As an initial sanity check, we project the parameters onto two random directions and
+        plot the resulting 3D loss landscapes and their squared differences.
+  \item \textbf{Hessian‐based analysis:} We then repeat the visualization and compute $\Delta_k$ by projecting onto the top two
+        (and up to $d = 10$) eigenvectors of the Hessian.  This directly tests our eigenvalue‐based approximation and the convergence behavior
+        predicted by theory.
 \end{enumerate}
 
-For smaller $N$, the landscape exhibits sharper fluctuations. As $N$ grows, the landscape often appears smoother and more stable.
+All code and detailed configuration files are available in our GitHub repository.
+
+\subsection{Preliminary Experiment: Projection onto Random Directions}
+\subsubsection{Visualizing the Loss Landscape}
+
+To build intuition for how the loss surface changes with training‐set size, we first project the model parameters onto two random
+directions. Specifically, we pick two random vectors $\mathbf v_1, \mathbf v_2 \in \mathbb R^P$ in parameter space and study the
+subspace they span.
+
+We then evaluate the empirical loss
+$$
+  \cL_k(\bw) =
+  \frac1k \sum_{i=1}^k \ell_i(\bw),
+$$
+where $\ell_i(\bw) = \ell (f_\bw(\bx_i), y_i)$, over a grid in the $(\alpha, \beta)$ plane defined by
+$$
+  \bw(\alpha,\beta) =
+  \bw^* + \alpha\, \mathbf v_1 + \beta\, \mathbf v_2, \quad
+  (\alpha, \beta) \in [-1,1]^2.
+$$
+
+\begin{enumerate}
+  \item \textbf{Grid Sampling.}
+        We form a uniform grid of $(\alpha, \beta)$ pairs in $[-1,1]\times[-1,1]$ with a fixed step size.
+  \item \textbf{Loss Computation.}
+        For each $(\alpha,\beta)$, we set $\bw=\bw(\alpha,\beta)$, keep the base weights $\bw^*$ fixed, and compute both
+        $\cL_k(\bw)$ and $\cL_{k+1}(\bw)$.
+  \item \textbf{Aggregated Metrics.}
+        \begin{itemize}
+          \item The surface $\cL_k(\alpha,\beta)$ visualizes how the mean loss changes over the grid.
+          \item The surface $\bigl( \cL_{k+1} (\alpha, \beta) - \cL_k (\alpha, \beta) \bigr)^2$ highlights where adding one sample
+                causes the largest change.
+        \end{itemize}
+  \item \textbf{3D Visualization.}
+        We render two side‐by‐side 3D plots:
+        \begin{itemize}
+          \item \emph{Left:} $\cL_k (\alpha, \beta)$ vs. $(\alpha, \beta)$.
+          \item \emph{Right:} $\bigl( \cL_{k+1} (\alpha, \beta) - \cL_k (\alpha, \beta) \bigr)^2$ vs. $(\alpha, \beta)$.
+        \end{itemize}
+\end{enumerate}
 
 \begin{figure}[!htbp]
-  \hspace*{-11.5cm}
-  \subfloat{\includegraphics[width=2\textwidth]{img/Ls_16.pdf}}
-  \caption{Loss function landscape}
-  \label{fig:base_loss}
+  \hspace*{-2.2cm}
+  \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_random_1_2.pdf}}
+  \caption{Loss function landscape for small $k$, projected onto the two random directions.}
+  \label{fig:loss_random_small}
 \end{figure}
@@
 \begin{figure}[!htbp]
-  \hspace*{-11.5cm}
-  \subfloat{\includegraphics[width=2\textwidth]{img/Ld_10.pdf}}
-  \caption{Loss function difference $(\mathcal{L}_{k} - \mathcal{L}_{k-1})^2$ landscape}
-  \label{fig:base_loss_diff}
+  \hspace*{-2.2cm}
+  \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_random_-2_-1.pdf}}
+  \caption{Loss function landscape for maximum $k$, projected onto the two random directions.}
+  \label{fig:loss_random_big}
 \end{figure}
 
-\subsection{Visualizing delta-function}
+\paragraph{Loss Landscape for Small Sample Size}
+In Figure~\ref{fig:loss_random_small} ($k=1$), the loss surface is highly irregular, with sharp peaks and deep valleys.
+This irregularity indicates that with very few training samples, the model is extremely sensitive to small parameter perturbations.
+The squared difference surface also exhibits large spikes, confirming that adding a single data point can cause substantial local changes
+in the loss.
 
-Now we dig deeper and visualize our $\Delta_k$ function to clearly understand the difference in the loss function when
-moving from a model trained on $k-1$ samples to one trained on $k$ samples. To do this we sample a bunch of random gaussian
-vectors (all components are normally distributed) and calculate a
-math expectation of modulus of the loss function difference.
-Directly computing this expectation over the entire distribution of random directions $\mathbf{d}$
-is impossible, so we employ a Monte Carlo approximation: using Monte-Carlo method (we take the mean value of loss function
-on sampled gaussian shifted points):
+\paragraph{Loss Landscape for Maximum Dataset Size}
+In contrast, Figure~\ref{fig:loss_random_big} (maximum $k$) shows a much smoother loss surface with gentle slopes. The overall loss
+values are lower --- reflecting the averaging effect of more samples --- and the squared difference surface is correspondingly much smaller.
+This smoothing demonstrates that as the dataset grows, the local optimum stabilizes, and additional samples produce only minor
+adjustments to the loss landscape. Hence, beyond a certain dataset size, further data have negligible impact on the shape of the loss
+surface.
 
-$$\Delta_k = \mathbb{E}_{p(\bw)} \left( \mathcal{L}_{k}(\bw) - \mathcal{L}_{k-1}(\bw) \right)^2 \approx
-  \frac{1}{k} \sum_{t=1}^{k} \left( \mathcal{L}_k(\mathbf{w}^* + \mathbf{d}_t) - \mathcal{L}_{k-1}(\mathbf{w}^* + \mathbf{d}_t) \right)^2$$
+\subsubsection{Visualizing the \texorpdfstring{$\Delta_k$}{Delta k} Function}
 
-where $\mathbf{d}_t \sim \mathcal{N}(0,I_d)$ are $K$ random gaussian samples drawn from a multivariate normal distribution
-with zero mean and identity covariance from a subspace of rank $d$. Specifically, we proceed as follows:
+To study how the loss landscape converges as the dataset grows, we estimate and plot the function
+$$
+  \Delta_k =
+  \mathbb{E}_{p(\bw)} \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr]^2
+$$
+via a Monte Carlo approximation using Gaussian perturbations in parameter space. The steps are:
 
 \begin{enumerate}
-  \item \textbf{Sampling Directions.} Generate $K$ independent directions $\mathbf{d}_1, \dots, \mathbf{d}_M$
-        from $\mathcal{N}(0,I_d)$.
-  \item \textbf{Decreasing multidimension.} Project the network’s parameters onto the subspace they span.
-  \item \textbf{Loss Evaluation.} For each direction $\mathbf{d}_t$, evaluate the loss at
-        $\mathbf{w}^* + \mathbf{d}_t$ and
-        $$\left( \mathcal{L}_{k}(\bw + \mathbf{d}_t) - \mathcal{L}_{k-1}(\bw + \mathbf{d}_t) \right)^2 $$
-  \item \textbf{Averaging.} Average these values values over $k = 1, \dots, K$ to approximate $\Delta_k$.
+  \item \textbf{Sample directions.}
+        Starting from a fixed trained model with parameters $\bw^*$, draw $K$ independent random vectors
+        $$
+          \btheta^{(t)} \sim \mathcal{N} (\mathbf 0, \sigma^2 \mathbf I_d), \quad
+          t = 1, \dots, K.
+        $$
+
+  \item \textbf{Perturb parameters.}
+        For each sample, form a perturbed parameter set
+        $$
+          \bw^{(t)} =
+          \bw^* + \btheta^{(t)}.
+        $$
+
+  \item \textbf{Evaluate loss differences.}
+        Compute
+        $$
+          \cL_k \bigl( \bw^{(t)} \bigr) \quad
+          \text{and} \quad
+          \cL_{k+1} \bigl( \bw^{(t)} \bigr),
+        $$
+        then calculate the squared difference
+        $$
+          d_t =
+          \bigl[ \cL_{k+1} (\bw^{(t)}) - \cL_k (\bw^{(t)}) \bigr]^2.
+        $$
+
+  \item \textbf{Compute Monte Carlo estimate.}
+        Approximate
+        $$
+          \Delta_k \approx
+          \frac{1}{K} \sum_{t=1}^K d_t.
+        $$
+        Increasing $K$ reduces the variance of this estimator.
+
+  \item \textbf{Plot results.}
+        We plot $\Delta_k$ versus $k$ to observe its decay. We also plot the scaled quantity
+        $$
+          \Delta_k \cdot k^2
+        $$
+        alongside to highlight an $\mathcal O(1 / k)$ convergence rate.
 \end{enumerate}
 
-By increasing $K$, we obtain a more reliable estimate of the expected difference in loss landscapes.
-This Monte Carlo approach is straightforward to implement and does not require explicit knowledge of the underlying loss
-function's distribution.
+Figures~\ref{fig:delta_random_sigma} and \ref{fig:delta_random_dim} present the behavior of the function
+$$
+  \Delta_k =
+  \mathbb{E} \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr]^2
+$$
+under different Monte Carlo sampling settings and projection subspace sizes. Each figure isolates one variable:
 
-The $\Delta$-function generally decreases as $N$ increases, suggesting that the landscape differences between neighbouring
-sizes become smaller.
-Beyond a certain point, $\Delta$-values are nearly negligible, indicating that adding more data has minimal impact on
-the local landscape in these directions.
-
-\noindent
-\textbf{Conclusion from Preliminary Experiments.}
-These initial experiments demonstrate that the loss landscape becomes progressively more stable with increasing $N$.
-However, the directions used here are either random. This motivates a more rigorous approach in the \emph{main experiment}
-below, where we focus specifically on the most critical directions of curvature --- those corresponding to the top
-eigenvalues of the Hessian.
-
-\subsection{Main Experiment: Projection onto Dominant Hessian Eigenvectors}
-
-From prior work \cite{sagun2018empirical}, we know that only a few Hessian eigenvectors (those associated with the largest
-eigenvalues) tend to capture the most significant curvature directions. By projecting onto these principal directions,
-we can more precisely measure how the landscape the loss surface changes as $k$ increases.
-
-\paragraph{Procedure}
-\begin{enumerate}
-  \item \textbf{Hessian approximation.} For each trained model, approximate
-        the Hessian $\mathbf{H}$ or its dominant eigenvalues/eigenvectors. Techniques ???
-  \item \textbf{Eigen-decomposition.} Identify the top $d$ eigenvectors $\{\mathbf{e}_1, \dots, \mathbf{e}_d\}$ with the
-        largest eigenvalues.
-  \item \textbf{Calculation and analization.} For each model, calculate a $\Delta$ fuction via the same way as in the second
-        experiment (using Monte-Carlo) and visualize via plots.
-\end{enumerate}
-
-\paragraph{Goal and Expected Outcome}
 \begin{itemize}
-  \item By focusing on the directions of largest curvature, we expect a clearer
-        measure of how the landscape of the loss surface changes with additional data.
-  \item This approach should yield a more accurate estimate of the point at which the
-        landscape effectively stabilizes, thus helping to pinpoint a minimum viable dataset size.
-  \item By applying experiments we check our calculations above on the reality.
+  \item Figure~\ref{fig:delta_random_sigma}: varying the Gaussian variance $\sigma$.
+  \item Figure~\ref{fig:delta_random_dim}: varying the subspace dimension $d$.
 \end{itemize}
@@
 \begin{figure}[!htbp]
-  \centering
-  \subfloat{\includegraphics[width=1\textwidth]{img/D_32.pdf}}\\
-  \caption{$\Delta$-function visualization. `Sigma` parameter refers to the variance of sampled gaussian vectors, `dim'
-    refers to the rank of the subspace.}
-  \label{fig:base_delta}
+  \hspace*{-2.6cm}
+  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_random_sigma_2_64.pdf}}
+  \caption{Delta function landscape varying the Gaussian variance $\sigma$.}
+  \label{fig:delta_random_sigma}
 \end{figure}
+
+\begin{figure}[!htbp]
+  \hspace*{-2.6cm}
+  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_random_dim_1_64.pdf}}
+  \caption{Delta function landscape varying the number of dimensions.}
+  \label{fig:delta_random_dim}
+\end{figure}
+
+\paragraph{Comparing Different Variances \texorpdfstring{$\sigma$}{sigma}}
+Figure~\ref{fig:delta_random_sigma} shows that varying the Gaussian variance $\sigma$ produces no substantial change in the shape or
+magnitude of the $\Delta_k$ curves. This insensitivity likely stems from using random projection directions: since they do not align
+with the dominant curvature axes of the Hessian, scaling those directions has little impact on the estimated loss differences.
+
+\paragraph{Comparing Different Subspace Dimensions}
+Figure~\ref{fig:delta_random_dim} shows that, when using random directions, increasing the projection dimension $d$ does not
+significantly alter the overall shape of $\Delta_k$. This suggests that only a few directions dominate the variation --- motivating our
+focus on Hessian‐based projections in the main experiment.
+
+In all cases, the right‐hand plots (showing $\Delta_k \cdot k^2$) exhibit a roughly $\mathcal O(1 / k)$ decay, confirming that the
+incremental effect of adding one more sample diminishes proportionally to $1 / k$ as the dataset grows.
+
+The preliminary experiments show that the loss landscape indeed smooths out and becomes more stable as the training set size $N$
+increases. However, because we used random projection directions, these results capture only a generic view of stability. This
+observation motivates our main experiment, in which we project onto the Hessian’s top eigenvectors --- i.e., the directions of greatest
+curvature --- to obtain a more precise understanding of when and how the loss surface converges.
+
+\subsection{Moving Forward: Projection onto Dominant Hessian Eigenvectors}
+
+\paragraph{Projection Directions.}
+In the main experiment, we replace random directions with the $d=10$ eigenvectors of the Hessian at the local minimizer $\bw^*$ that
+correspond to its largest eigenvalues.  Prior work \cite{sagun2018empirical} shows these top modes capture nearly all of the
+non‑negligible curvature, and Figure~\ref{fig:evgen} confirms a rapid eigenvalue decay. Fixing $d=10$ balances computational cost and
+fidelity, and our tests on Fashion‑MNIST and CIFAR demonstrate that this choice suffices even for complex architectures.
+
+\paragraph{Objectives.}
+By projecting onto these principal directions, we aim to:
+\begin{itemize}
+  \item Obtain a clearer view of how the loss surface changes as $k$ grows, since these directions drive the largest loss variations.
+  \item Accurately identify the threshold $k^*$ at which further data cause only negligible shifts in the local curvature.
+  \item Empirically validate our theoretical $\Delta_k$ estimates by comparing them against Monte Carlo measurements in this eigenvector
+        subspace.
+\end{itemize}
+
+\subsubsection{Visualizing the Loss Landscape}
+
+Figures~\ref{fig:loss_eigen_small} and \ref{fig:loss_eigen_big} display 3D plots of $\cL_k(\btheta)$ (left) and $\bigl[\cL_{k+1}(\btheta)-\cL_k(\btheta)\bigr]^2$ (right), where $\btheta$ varies over the span of the top two Hessian eigenvectors.
+The procedure mirrors the random‑direction case but uses these maximally informative axes.
+
+\begin{figure}[!htbp]
+  \hspace*{-2.2cm}
+  \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_eigen_1_2.pdf}}
+  \caption{Loss landscape $\cL_k$ and squared difference for small $k$, projected onto the top two Hessian eigenvectors.}
+  \label{fig:loss_eigen_small}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \hspace*{-2.2cm}
+  \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_eigen_-2_-1.pdf}}
+  \caption{Loss landscape $\cL_k$ and squared difference for maximum $k$, projected onto the top two Hessian eigenvectors.}
+  \label{fig:loss_eigen_big}
+\end{figure}
+
+In Figure~\ref{fig:loss_eigen_small} (small $k$, e.g.\ $k=1$), the loss surface projected onto the top two Hessian eigenvectors is
+noticeably more structured and less erratic than in the random‐direction case. This confirms that these eigenvectors capture the
+most influential directions of loss variation. The squared‐difference plot also shows larger peaks, indicating that the model is
+highly sensitive to new data along these key axes when $k$ is small.
+
+By contrast, Figure~\ref{fig:loss_eigen_big} (maximum $k$) shows a much smoother loss surface and a greatly reduced squared‐difference.
+As the dataset grows, the loss changes less along the principal curvature directions, and the model’s local behavior stabilizes.
+Notably, even at large $k$, the squared difference remains higher than in the random‐direction experiment, providing a practical upper
+bound on landscape fluctuations. These findings reinforce that projecting onto the top eigenvectors effectively reveals the convergence
+behavior of the loss landscape and helps intuitively understand the existence of the point beyond which additional data have minimal impact.
+
+\subsubsection{Visualizing the \texorpdfstring{$\Delta_k$}{Delta k} Function}
+
+\begin{figure}[!htbp]
+  \hspace*{-2.6cm}
+  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_eigen_1_10_64.pdf}}
+  \caption{Delta function landscape}
+  \label{fig:delta_eigen}
+\end{figure}
+
+Figure~\ref{fig:delta_eigen} presents the Monte Carlo estimate of $\Delta_k$ in the subspace spanned by the top $d = 10$
+Hessian eigenvectors. Compared to the random‐direction experiment, $\Delta_k$ here has a higher value --- reflecting the
+large curvature captured by these eigenvectors. The right‐hand plot of $\Delta_k\cdot k^2$
+confirms an approximate $O(1 / k)$ convergence rate. These observations show that projecting onto the leading eigen-directions not only
+highlights the greatest loss sensitivity but also yields a reliable upper bound on how quickly the landscape stabilizes as more data are
+added.
+
+\subsection{Main Experiment: Theoretical vs. Empirical \texorpdfstring{$\Delta_k$}{Delta k}}
+
+In this experiment, we compare our Hessian‐based theoretical estimate of $\Delta_k$ with the empirical Monte Carlo measurements.
+
+We approximate $\Delta_k$ by
+$$
+  \Delta_k \approx
+  \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^{d} (\lambda_{k+1}^i - \lambda_{k}^i)^2 +
+  \Bigl( \sum_{i=1}^{d} (\lambda_{k+1}^i - \lambda_{k}^i) \Bigr)^2 \Biggr),
+$$
+where $\{\lambda_k^i\}_{i=1}^d$ and $\{\lambda_{k+1}^i\}_{i=1}^d$ are the top $d$ Hessian eigenvalues at sample sizes $k$ and $k + 1$,
+and $\sigma$ is the perturbation variance.
+
+\begin{enumerate}
+  \item \textbf{Model \& subspace size.} Train a network to obtain $\bw^*$. Choose $d$ based on the rapid eigenvalue decay in
+        Figure~\ref{fig:evgen}.
+  \item \textbf{Eigenvalue computation.} For each $k$ in the range of interest (typically the largest $K$ sizes), compute the top
+        $d$ eigenvalues via the power method.
+  \item \textbf{Theoretical $\Delta_k$.} Plug these eigenvalues into the formula above to obtain a sequence of theoretical estimates.
+  \item \textbf{Empirical $\Delta_k$.} Independently estimate $\Delta_k$ by Monte Carlo sampling of perturbations along the same
+        top‐$d$ eigenvectors as it was in the previous experiment.
+  \item \textbf{Comparison.} Plot both theoretical and empirical $\Delta_k$ on the same axes to assess how tightly the theory bounds
+        the observed values.
+\end{enumerate}
+
+\begin{figure}[!htbp]
+  \subfloat{\includegraphics[width=1\textwidth]{img/delta_border_1_10_64.pdf}}
+  \caption{Delta function landscape}
+  \label{fig:delta_eigen_comparison}
+\end{figure}
+
+Figure~\ref{fig:delta_eigen_comparison} shows the theoretical estimate of $\Delta_k$ (dashed line) alongside the empirical Monte Carlo
+measurements (solid line).  Across all values of $k$, the theory curve remains above the empirical one, demonstrating that our
+eigenvalue‐based formula provides a valid upper bound for the true loss‐landscape changes.
 
@@
 
-TADAAA!
+The gap between the theoretical and empirical curves arises because the estimate uses only the top $d$ eigenvalues.
+Therefore, focusing on the leading modes yields a reliable and efficient criterion for determining when
+adding more data has negligible effect on the local loss geometry.
+
+Our Hessian‐projection framework provides a clear picture of how a neural network’s loss landscape evolves with training–set size.
+By restricting attention to the $d$ principal curvature directions (the top Hessian eigenvectors), we obtain far more structured 3D
+loss surfaces and a sharply decaying landscape‐difference measure
+$$
+  \Delta_k \;=\; \mathbb{E}_{p(\btheta)}\bigl[L_{k+1}(\btheta)-L_k(\btheta)\bigr]^2,
+$$
+which empirically follows an approximate $O(1/k)$ trend.  Moreover, the Hessian‐based bound
+$$
+  \Delta_k\approx\frac{\sigma^4}{4}\Bigl(2\sum_{i=1}^d(\lambda^i_{k+1}-\lambda^i_k)^2
+  +\bigl(\sum_{i=1}^d(\lambda^i_{k+1}-\lambda^i_k)\bigr)^2\Bigr)
+$$
+consistently upper‐bounds the Monte Carlo estimates (\ref{fig:delta_eigen_comparison}). This confirms that only a small number of
+eigen‐modes drive the bulk of loss‐landscape change.
+
+A practical outcome is a stopping criterion for data collection: once the leading eigenvalues and $\Delta_k$ plateau,
+adding more samples yields negligible changes in the loss geometry.  This lets practitioners avoid unnecessary annotation and
+compute without sacrificing model stability.
+
+\paragraph{Limitations and Future Work}
+Our study used $d=10$ principal directions and relatively MLPs on MNIST‐like tasks. Deeper or convolutional architectures,
+alternative loss functions, or non–i.i.d. data may shift the spectrum decay and require reconsidering $d$. Extending this Hessian‐based
+convergence criterion to large models or streaming data remains an exciting direction.
 
@@
 
-TADAAA!
+We have introduced and validated a Hessian‐projection approach to identify when a neural network’s loss landscape “settles” as
+training data grows. By projecting onto the top eigenvectors of the Hessian, we both visualize and quantify the landscape’s
+sensitivity via the $\Delta_k$ function, and derive a tight eigenvalue‐difference bound that serves as a reliable upper bound on
+loss‐landscape changes. Our results demonstrate a clear “sufficient sample size” threshold: beyond this, further data become
+insignificant in local curvature. This criterion provides a concrete guide for efficient dataset sizing and early stopping in
+model training.
 
@@
 \appendix
-\section{Appendix / supplemental material}\label{app}
+\section{Appendix}\label{app}
 
-\subsection{Additional experiments / Proofs of Theorems}\label{app:exp}
+\subsection{Proof of Theorem~1}\label{app:th_1}
 
-TADAAA!
+\begin{proof}
+  Starting from the second‐order Taylor expansion around the local minimizer $\bw^*$, we have
+  $$
+    \cL_{k+1}(\bw) - \cL_k(\bw) \approx
+    \bigl[ \cL_{k+1}(\bw^*) - \cL_k(\bw^*) \bigr] + \tfrac12\, \btheta^{\T}\! \bigl( \bLambda_{k+1} - \bLambda_k \bigr)\, \btheta,
+  $$
+  where $\btheta = \bP^{\T} (\bw-\bw^*)$ and $\bLambda_j = \bP^{\T} \bH_j(\bw^*) \bP$.
+
+  Let $p(\btheta)$ be a distribution with mean $\bm=\mathbb E[\btheta] = 0$ and covariance $\bSigma = \mathbb D[\btheta]$. Denote
+  $$
+    \bA=\bLambda_{k+1} - \bLambda_k.
+  $$
+  Then, by standard matrix‐expectation results (see \cite[p.~35]{petersen2012matrix}),
+  $$
+    \mathbb E \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr] =
+    \mathbb E \bigl[ \tfrac12\, \btheta^{\T} \bA\, \btheta \bigr] =
+    \tfrac12\, \bigl( \tr(\bA \bSigma) + \bm^{\T} \bA \bm \bigr) =
+    \tfrac12\, \tr(\bA \bSigma).
+  $$
+  Similarly, using \cite[p.~43]{petersen2012matrix},
+  $$
+    \mathbb D \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr] =
+    \tfrac14 \Bigl( 2\, \tr(\bA^2 \bSigma^2) + 4\, \bm^{\T} \bA^2 \bm\, \sigma^2 \Bigr) =
+    \tfrac12\, \sigma^4\, \tr(\bA^2),
+  $$
+  where we have set $\bSigma=\sigma^2\mathbf I$ and used $\bm=0$.
+
+  Since
+  $$
+    \Delta_k =
+    \mathbb D \bigl[ \cL_{k+1} - \cL_k \bigr] + \bigl( \mathbb E [\cL_{k+1} - \cL_k] \bigr)^2,
+  $$
+  we get
+  $$
+    \Delta_k \approx
+    \tfrac12\, \sigma^4\, \tr(\bA^2) + \tfrac14\, \bigl( \tr(\bA \bSigma) \bigr)^2 =
+    \tfrac{\sigma^4}{4} \Bigl( 2\, \tr(\bA^2) + \bigl( \tr(\bA) \bigr)^2 \Bigr).
+  $$
+  Finally, since
+  $$
+    \tr(\bA) =
+    \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i), \quad
+    \tr(\bA^2) =
+    \sum_{i=1}^d( \lambda_{k+1}^i - \lambda_k^i)^2,
+  $$
+  we arrive at the stated approximation
+  $$
+    \Delta_k \approx
+    \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i)^2
+    + \Bigl( \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i) \Bigr)^2 \Biggr).
+  $$
+\end{proof}
 

### 2025-04-17 code/src/calc.py

@@ -1,57 +1,19 @@
 import copy
-import torch
 import scipy.stats as sps
 import numpy as np
 from tqdm.auto import tqdm
-
-
-def inplace_sum_models(model1, model2, coef1, coef2):
-    final = model1
-    for (name1, param1), (name2, param2) in zip(final.state_dict().items(), model2.state_dict().items()):
-        transformed_param = param1 * coef1 + param2 * coef2
-        param1.copy_(transformed_param)
-    return final
-
-
-def calc_sum_models(model1, model2, coef1, coef2):
-    final = copy.deepcopy(model1)
-    final.load_state_dict(copy.deepcopy(model1.state_dict()))
-    return inplace_sum_models(final, model2, coef1, coef2)
-
-
-def init_from_params(model, direction):
-    for p_orig, p_other in zip(model.parameters(), direction):
-        with torch.no_grad():
-            p_orig.copy_(p_other)
-
-
-def get_loss_list(model, criterion, dataloader, device):
-    model.eval()
-    losses = []
-
-    for x, y in dataloader:
-        x, y = x.to(device), y.to(device)
-
-        outputs = model(x)
-        loss = criterion(outputs, y)
-
-        losses.append(loss.detach().cpu().item())
-
-    return losses
-
-
-def create_loss_list_func(criterion, dataloader):
-    def calc_losses(model):
-        return get_loss_list(model, criterion, dataloader, model.device)
-
-    return calc_losses
+from src.models import get_loss
+from src.models import inplace_sum_models
+from src.models import calc_sum_models
+from src.models import init_from_params
 
 
 class LossCalculator:
     def __init__(self, model, criterion, dataloader, core):
         self.model = model
+        self.criterion = criterion
+        self.dataloader = dataloader
         self.core = core
-        self.loss_func = create_loss_list_func(criterion, dataloader)
         self.directions = []
 
     def calc_losses(self, grid):
@@ -60,13 +22,14 @@ class LossCalculator:
 
         result = {}
 
-        for coef1, coef2 in tqdm(grid):
+        for coef1, coef2 in tqdm(grid, desc="Grid point"):
             target_add = [p1 * coef1 + p2 * coef2 for p1, p2 in zip(self.directions[0], self.directions[1])]
             target_add_model = copy.deepcopy(self.model)
             init_from_params(target_add_model, target_add)
 
             target_model = calc_sum_models(self.model, target_add_model, 1, 1)
-            losses = self.loss_func(target_model)
[…]

### 2025-04-17 code/src/directions.py

@@ -6,31 +6,8 @@
 # --========================================-- #
 
 import torch
-from tqdm.auto import tqdm
-
-
-def get_loss_mean(model, criterion, dataloader, device):
-    model.eval()
-
-    count = 0
-    total_loss = 0.0
-    for x, y in dataloader:
-        x, y = x.to(device), y.to(device)
-
-        outputs = model(x)
-        loss = criterion(outputs, y)
-
-        total_loss += loss
-        count += 1
-
-    return total_loss / count
-
-
-def create_loss_mean_func(criterion, dataloader):
-    def calc_loss(model):
-        return get_loss_mean(model, criterion, dataloader, model.device)
-
-    return calc_loss
+from src.models import get_loss
+from src.eigenvalues import HessianEigenvector
 
 
 class RandomDirection:
@@ -127,51 +104,11 @@ class RandomDirection:
         return direction
 
 
-class MaxEigenvector:
+class EigenDirection:
     def __init__(self, model, criterion, dataloader):
         self.model = model
-        self.loss_func = create_loss_mean_func(criterion, dataloader)
+        self.loss = get_loss(model, criterion, dataloader, model.device).mean()
         self.type = "eigen"
 
-        self.eigenvectors = []
-        self.eigenvalues = []
-
-    def get(self, k, num_iters=100, tol=1e-6):
-        loss = self.loss_func(self.model)
-
-        if k <= len(self.eigenvectors):
-            return self.eigenvalues[:k], self.eigenvectors[:k]
-
-        for i in range(0, k - len(self.eigenvectors)):
-            eigenvalue, eigenvector = self._power_iteration(loss, num_iters=num_iters, tol=tol)
-            self.eigenvalues.append(eigenvalue)
-            self.eigenvectors.append(eigenvector)
-        return self.eigenvectors
-
-    def _power_iteration(self, loss, num_iters=100, tol=1e-6):
-        params = list(self.model.parameters())
-        flat_params = torch.cat([p.view(-1) for p in params])
-
-        v = torch.randn(flat_params.shape[0], device=self.model.device)
-        v = v / v.norm()
-        eigenvalue = None
-        for _ in tqdm(range(1, num_iters + 1)):
-            hv = self._hvp(params, loss, v)
-
-            for d in self.eigenvectors:
-                hv = hv - torch.dot(hv, d) * d
-            new_v = hv / hv.norm()
-            eigenvalue = torch.dot(new_v, hv)
[…]

### 2025-04-17 code/src/eigenvalues.py

@@ -1,0 +1,74 @@
+# --========================================-- #
+#   * Author  : NTheme - All rights reserved
+#   * Updated : 16 April 2025
+#   * File    : eigenvalues.py
+#   * Project : code
+# --========================================-- #
+
+from tqdm.auto import tqdm
+import torch
+
+
+class HessianEigenvector:
+    def __init__(self, variables, func):
+        self.variables = list(variables)
+        self.func = func
+
+        self.eigenvectors = []
+        self.eigenvalues = []
+
+    def get(self, k, num_iters=1000, tol=1e-7):
+        if k <= len(self.eigenvectors):
+            return self.eigenvalues[:k], self.eigenvectors[:k]
+
+        for i in range(0, k - len(self.eigenvectors)):
+            eigenvalue, eigenvector = self._power_iteration(num_iters, tol)
+            self.eigenvalues.append(eigenvalue.detach().cpu().item())
+            self.eigenvectors.append(eigenvector)
+        return self.eigenvalues[:k], self.eigenvectors[:k]
+
+    def _power_iteration(self, num_iters, tol):
+        v = [torch.randn_like(p) for p in self.variables]
+        v = self._normalize(v)
+
+        eigenvalue = None
+        for _ in tqdm(range(num_iters), desc="Power iter"):
+            hv = self._hvp(self.variables, self.func, v)
+
+            for d in self.eigenvectors:
+                hv = self._projection_diff(hv, d)
+
+            new_v = self._normalize(hv)
+            eigenvalue = self._dot(new_v, hv)
+
+            if self._norm_diff(new_v, v) < tol:
+                v = new_v
+                break
+            v = new_v
+
+        return eigenvalue, v
+
+    @staticmethod
+    def _hvp(variables, func, v):
+        grads = torch.autograd.grad(func, variables, create_graph=True)
+        dots = [torch.dot(g.contiguous().view(-1), vi.contiguous().view(-1)) for g, vi in zip(grads, v)]
+        grad_v = torch.sum(torch.stack(dots))
+        hv = torch.autograd.grad(grad_v, variables, retain_graph=True)
+        return [h.contiguous() for h in hv]
+
+    @staticmethod
+    def _dot(a, b):
+        dots = [torch.dot(ai.view(-1), bi.view(-1)) for ai, bi in zip(a, b)]
+        return torch.stack(dots).sum()
+
+    def _normalize(self, vec):
[…]

### 2025-04-17 code/src/models.py

@@ -5,6 +5,8 @@
 #   * Project : untitled
 # --========================================-- #
 
+import copy
+import torch
 import torch.nn as nn
 import torch.nn.functional as F
 from tqdm.auto import tqdm
@@ -55,7 +57,7 @@ def train(model, criterion, dataloader, optimizer, num_epochs=10, ret=False, log
     model.train()
 
     losses = []
-    for epoch in tqdm(range(1, num_epochs + 1)):
+    for epoch in tqdm(range(1, num_epochs + 1), desc="Epoch"):
         running_loss = 0.0
         batches_count = 0
 
@@ -78,3 +80,34 @@ def train(model, criterion, dataloader, optimizer, num_epochs=10, ret=False, log
             print(f"Epoch [{epoch}/{num_epochs}]: loss = {avg_loss:.4f}")
 
     return losses if ret else None
+
+
+def get_loss(model, criterion, dataloader, device):
+    model.eval()
+    losses = []
+    for x, y in dataloader:
+        x, y = x.to(device), y.to(device)
+        outputs = model(x)
+        loss = criterion(outputs, y)
+        losses.append(loss)
+    return torch.stack(losses, dim=0)
+
+
+def inplace_sum_models(model1, model2, coef1, coef2):
+    final = model1
+    for (name1, param1), (name2, param2) in zip(final.state_dict().items(), model2.state_dict().items()):
+        transformed_param = param1 * coef1 + param2 * coef2
+        param1.copy_(transformed_param)
+    return final
+
+
+def calc_sum_models(model1, model2, coef1, coef2):
+    final = copy.deepcopy(model1)
+    final.load_state_dict(copy.deepcopy(model1.state_dict()))
+    return inplace_sum_models(final, model2, coef1, coef2)
+
+
+def init_from_params(model, direction):
+    for p_orig, p_other in zip(model.parameters(), direction):
+        with torch.no_grad():
+            p_orig.copy_(p_other)

### 2025-04-17 code/src/visualize.py

@@ -1,13 +1,10 @@
-import itertools
-from matplotlib import cm
-import numpy as np
-import matplotlib.pyplot as plt
-from mpl_toolkits.mplot3d import Axes3D
 import copy
-import scipy.stats as sps
+import itertools
+import numpy as np
 from tqdm.auto import tqdm
 
-from src.calc import DeltaCalculator
+from matplotlib import cm
+import matplotlib.pyplot as plt
 
 
 class LossVisualizer:
@@ -74,7 +71,7 @@ class DeltaVisualizer:
         self.calculator = calculator
         self.core_type = calculator.core.type
 
-    def visualize_all(self, params, num_samples, begin):
+    def visualize_all(self, params, num_samples=64, begin=10):
         params = copy.deepcopy(params)
 
         deltas = self.calculator.calc_deltas(params, num_samples=num_samples)
@@ -86,8 +83,8 @@ class DeltaVisualizer:
         axs[0].plot(deltas)
         axs[1].plot(deltas * mult_coef)
 
-        ylabels = [r'$\mathbb{E}_{p(\mathbf{\theta})}(L_k - L_{k-1})^2$',
-                   r'$\mathbb{E}_{p(\mathbf{\theta})}(L_k - L_{k-1})^2 \cdot k^2$']
+        ylabels = [r'$\mathbb{E}(\mathcal{L}_{k+1} - \mathcal{L}_{k})^2$',
+                   r'$\mathbb{E}(\mathcal{L}_{k+1} - \mathcal{L}_{k})^2 \cdot k^2$']
 
         axs[0].set(
             xlabel='k',
@@ -105,12 +102,39 @@ class DeltaVisualizer:
             "../paper/img/delta_" + self.core_type + f"_{params['sigma']}_{params['dim']}_{num_samples}" + ".pdf")
         plt.show()
 
+    def visualize_border(self, border, params, num_samples=64, begin=10):
+        params = copy.deepcopy(params)
+
+        deltas = self.calculator.calc_deltas(params, num_samples=num_samples)
+
+        fig, ax = plt.subplots(figsize=(8, 6))
+
+        offset = len(deltas) - len(border)
+        ax.plot(np.arange(offset, len(deltas)), deltas[offset:], label='Δ_k')
+        ax.plot(np.arange(offset, len(deltas)), border, label='border', linestyle='--')
+
+        ax.set(
[…]

### 2025-04-17 slides/full/main.tex
Universal commits presents...

\documentclass{beamer}
\beamertemplatenavigationsymbolsempty
\usecolortheme{beaver}
\setbeamertemplate{blocks}[rounded=true, shadow=true]
\setbeamertemplate{footline}[page number]
%
\input{new_commands}
%
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amssymb,amsfonts,amsmath,mathtext}
\usepackage{subfig}
\usepackage[all]{xy}
\usepackage{array}
\usepackage{multicol}
\usepackage{hyperref}
\usepackage{hhline}
\graphicspath{{fig/}{../fig/}}

%----------------------------------------------------------------------------------
\title[Loss Landscape Convergence]{Neural Networks Loss Landscape Convergence in Hessian Low-Dimensional Space}
\author[Tem Nikitin \textit{et al.}]{Tem Nikitin \and Nikita Kiselev \and Andrey Grabovoy \and Vladislav Meshkov}
\institute{Moscow Institute of Physics and Technology}
\date{2025}
%----------------------------------------------------------------------------------
\begin{document}

\begin{frame}
    \thispagestyle{empty}
    \maketitle
\end{frame}

\begin{frame}{Goals}
    \begin{itemize}
        \item Study how neural network loss landscape changes with dataset size
        \item Define and measure $\Delta_k = \mathbb{E} \bigl( \cL_{k+1}(\bw) - \cL_k(\bw) \bigr)^2$
        \item Develop Hessian-based low-dimensional projection method
        \item Derive theoretical bound for $\Delta_k$ via top-$d$ eigenvalues
        \item Validate threshold $k^*$ beyond which further data yield negligible change
    \end{itemize}
\end{frame}

\begin{frame}{One-slide}
    \begin{columns}[c]
        \hspace{-1.5cm}
        \column{0.6\textwidth}
        \includegraphics[width=1.2\textwidth]{img/loss_eigen_-1_individ.pdf}
        \column{0.5\textwidth}
        \begin{itemize}
            \item Hessian projection onto top-$d$ eigenvectors
            \item Monte Carlo estimate of $\Delta_k^{emp} = \mathbb{E} \Bigl( \cL_{k+1} - \cL_{k} \Bigr)^2.$
            \item Analytical bound: $\Delta_k^{th} \ge \Delta_k^{emp}$
            \item Detect dataset threshold $k^*$
        \end{itemize}
    \end{columns}
    \bigskip
    \textbf{Main message:} Loss landscape stabilizes after sufficient sample size.
\end{frame}

\begin{frame}{Literature}
    \begin{itemize}
        \item Wu \textit{et al.} (2017): loss landscapes vs dataset size [2]
        \item Sagun \textit{et al.} (2018): Hessian low effective rank [7]
        \item Li \textit{et al.} (2018): visualizing loss surfaces [11]
        \item Ghorbani \textit{et al.} (2019): eigenvalue density analysis [8]
        \item Bousquet \& Elisseeff (2002): stability and generalization bounds [12]
    \end{itemize}
\end{frame}

\begin{frame}{Problem Statement: Hypothesis and Model}
    \begin{block}{Hypothesis}
        Beyond some $k^*$, adding new samples changes the local loss landscape by less than a tolerance $\Delta_{tol}$, i.e.
        $\forall\, k \ge k^* : \; \Delta_k < \Delta_{tol}$.
    \end{block}
    \begin{block}{Model}
        \begin{itemize}
            \item MLP with ReLU activations for $K$-class classification
            \item Empirical loss: $\cL_k(\bw) = \frac1k \sum_{i=1}^k \ell_i(\bw)$
            \item Hessian: $\bH_k(\bw) = \nabla^2_{\bw} \cL_k(\bw)$
        \end{itemize}
    \end{block}
\end{frame}

\begin{frame}{Problem Statement: Quality Criteria}
    \begin{itemize}
        \item Convergence rate: $\Delta_k = O(1/k)$
        \item Theoretical bound via top-$d$ eigenvalues upper-bounds empirical $\Delta_k$
        \item Plateau in eigenvalue differences $\lambda_i^{k+1}-\lambda_i^k$ indicates threshold
    \end{itemize}
\end{frame}

\begin{frame}{Problem Solution: Theoretical Analysis}
    \begin{columns}[c]
        \column{0.4\textwidth}
        \begin{itemize}
            \item Project parameters: $\bw = \bw^* + \bP \btheta$, $\bP = [\be_1, \dots, \be_d]$
            \item Taylor approx: $\cL_k (\bw^* + \bP \btheta) \approx \cL_k (\bw^*) + \tfrac12 \btheta^{\T} \bLambda_k \btheta$
                  \bigskip
                  \bigskip
            \item Bound:
        \end{itemize}
        \column{0.6\textwidth}
        \includegraphics[width=1.0\textwidth]{img/eigenvalues.pdf}
        \vspace{0.5cm}\hspace{1.8cm}\raggedright{\tiny Figure 1: Eigenvalue decay (p.4)}
    \end{columns}
    $$
        \Delta_k \approx
        \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i)^2
        + \Bigl( \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i) \Bigr)^2 \Biggr).
    $$
\end{frame}

\begin{frame}{Goals of Computational Experiment}
    \begin{itemize}
        \item Datasets: MNIST, Fashion-MNIST (60k train, 10k test)
        \item MLP: 2 hidden layers, $\sim 10^5$ parameters
        \item Subspace dimension $d=10$, Monte Carlo samples $K=64$, $\sigma=1$
        \item Compare empirical $\Delta_k$ vs theoretical bound across $k$
    \end{itemize}

    \begin{figure}
        \centering
        \includegraphics[width=1.05\textwidth,trim=0 17 0 1.5cm,clip]{img/delta_eigen_1_10_64.pdf}
        \caption*{\tiny Figure 8: Monte Carlo $\Delta_k$ vs $k$ and $\Delta_k\cdot k^2$ (p.11)}
    \end{figure}
\end{frame}

\begin{frame}{Error Analysis}
    \begin{itemize}
        \item Empirical $\Delta_k$ consistently below theoretical estimate
        \item Gap due to neglected eigen-modes beyond top-$d$
        \item Monte Carlo variance decreases with sample size $K$
    \end{itemize}

    \begin{figure}
        \centering
        \includegraphics[width=0.7\textwidth]{img/delta_border_1_10_64.pdf}
        \caption*{\tiny Figure 9: Theoretical (dashed) vs empirical (solid) $\Delta_k$ (p.12)}
    \end{figure}
\end{frame}

%\begin{frame}{Models--Datasets--Criteria}
%    \begin{center}
%        \begin{tabular}{|l|l|c|c|}
%            \hline
%            \textbf{Model} & \textbf{Dataset} & \boldmath$d$ & \boldmath$\Delta_{tol}$ \\
%            \hhline{|=|=|=|=|}
%            MLP (2 layers) & MNIST            & 10           & $1\times10^{-4}$        \\
%            \hline
%            MLP (2 layers) & Fashion-MNIST    & 10           & $1\times10^{-4}$        \\
%            \hline
%        \end{tabular}
%    \end{center}
%\end{frame}

\begin{frame}{Results and Conclusions}
    \begin{itemize}
        \item Identified threshold $k$ (MNIST), (Fashion-MNIST)
        \item Loss landscape stabilizes: additional data negligible beyond $k^*$
        \item Hessian-based bound provides reliable upper-bound for $\Delta_k$
        \item Practical guideline for dataset sizing and early stopping
    \end{itemize}
\end{frame}

\end{document}

### 2025-04-17 slides/inn_prac/iteration_1/innprac_1.tex
Universal commits presents...

\documentclass{beamer}

\usepackage[utf8]{inputenc}
\usetheme{Madrid}

\title[Loss Landscape Convergence]{Neural Networks Loss Landscape Convergence in Different Low-Dimensional Spaces}
\author{Tem Nikitin \ Nikita Kiselev}
\institute[MIPT]{Moscow Institute of Physics and Technology}
\date{\today}

\begin{document}

\begin{frame}
    \titlepage
\end{frame}

\begin{frame}{Problem Statement}
    \begin{itemize}
        \item Neural networks achieve higher accuracy with larger datasets.
        \item However, computational resources limit training possibilities.
        \item Key question: \textbf{When does additional data stop reshaping the loss landscape?}
    \end{itemize}
\end{frame}

\begin{frame}{Proposed Method}
    \begin{itemize}
        \item Analyze loss landscape via Hessian eigenvectors.
        \item Project high-dimensional parameter space onto lower-dimensional space.
        \item Utilize Monte-Carlo methods to approximate landscape changes.
    \end{itemize}
\end{frame}

\begin{frame}{Main Idea}
    \begin{itemize}
        \item \textbf{Hessian-based projections} identify primary curvature directions.
        \item Calculate a $\Delta$-function to quantify changes when increasing dataset size:
              $$\Delta_k = \int (L_k(w)-L_{k-1}(w))^2 p(w) dw$$
        \item Identify the threshold at which further data does not significantly alter landscape.
    \end{itemize}
\end{frame}

% Слайд из файла пользователя (обязательный)
\begin{frame}{Neural Networks Loss Landscape Convergence in Different Low-Dimensional Spaces}

    \textbf{Goal:} Measure how the loss function changes as the training set size grows:
    \[
        \Delta_k = \mathbb{E}_{p(\mathbf{w})}\Bigl(\mathcal{L}_{k}(\mathbf{w}) - \mathcal{L}_{k-1}(\mathbf{w})\Bigr)^2.
    \]

    \textbf{Method:}
    \begin{itemize}
        \item \textbf{Monte Carlo:} Generate points near the minimum according to $p(\mathbf{w})$ and average the differences.
        \item \textbf{Hessian Eigenvectors:} Use directions with the largest eigenvalues to focus on key curvature components.
    \end{itemize}

    \begin{columns}[t]
        \column{0.35\textwidth}
        \centering
        \hspace*{-2cm}
        \includegraphics[width=0.70\textwidth]{img/LS_16.jpg}\\

        \column{0.47\textwidth}
        \centering
        \vspace*{-3cm}
        \hspace*{-2.2cm}
        \includegraphics[width=1.375\textwidth]{img/D_32.jpg}\\
        \hspace*{-1.2cm}
        \scriptsize \textit{$\Delta_k$ function}
    \end{columns}

\end{frame}

\begin{frame}{Experiment Results}
    \begin{itemize}
        \item Experiments conducted on MNIST and Fashion-MNIST datasets.
        \item Low-dimensional visualization clearly illustrates stabilization.
        \item Verified theoretical bounds empirically.
    \end{itemize}
\end{frame}

\begin{frame}{Practical Implications}
    \begin{itemize}
        \item Enables identification of minimal viable dataset size.
        \item Reduces computational cost significantly.
        \item Provides guidelines for efficient data collection.
    \end{itemize}
\end{frame}

\begin{frame}{References}
    \scriptsize
    \begin{thebibliography}{9}
        \bibitem{1} Soydaner, D. \textit{Optimization algorithms for deep learning}, 2020.
        \bibitem{2} Wu et al. \textit{Loss landscape and generalization}, 2017.
        \bibitem{3} Petersen, Pedersen. \textit{The matrix cookbook}, 2012.
        \bibitem{4} Deng, L. \textit{MNIST database}, 2012.
        \bibitem{5} Xiao et al. \textit{Fashion-MNIST}, 2017.
        \bibitem{6} Kiselev, Grabovoy. \textit{Unraveling the Hessian}, 2024.
        \bibitem{7} Sagun et al. \textit{Hessian analysis of networks}, 2018.
    \end{thebibliography}
\end{frame}

\end{document}

### 2025-04-17 slides/one_slide/one_slide.tex
Universal commits presents...

\documentclass[10pt]{beamer}
\beamertemplatenavigationsymbolsempty
\usecolortheme{beaver}
\setbeamertemplate{blocks}[rounded=true, shadow=true]
%\setbeamertemplate{footline}[page number]

\usepackage[utf8]{inputenc}
\usepackage[english,russian]{babel}
\usepackage{amssymb,amsfonts,amsmath,mathtext}
\usepackage{subfig}
\usepackage[all]{xy}
\usepackage{array}
\usepackage{multicol}
\usepackage{hyperref}
\usepackage{hhline}
\usepackage{caption}
\usepackage{subfig}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\graphicspath{ {../figures/} }

%----------------------------------------------------------------------------------------------------------
\title[Loss Landscape Convergence]{Neural Networks Loss Landscape Convergence in Hessian Low-Dimensional Space}
\author[Tem Nikitin \textit{et al.}]{Tem Nikitin \and Nikita Kiselev \and Andrey Grabovoy \and Vladislav Meshkov}
\institute{Moscow Institute of Physics and Technology}
\date{2025}
%----------------------------------------------------------------------------------------------------------
\begin{document}

\begin{frame}{Neural Networks Loss Landscape Convergence in Different Low-Dimensional Spaces}

    \textbf{Goal:} Measure how the loss function changes as the training set size grows:
    $$
        \Delta_k =
        \mathbb{E} \Bigl( \mathcal{L}_{k+1} (\mathbf{w}) - \mathcal{L}_{k} (\mathbf{w}) \Bigr)^2.
    $$

    \textbf{Method:}
    \begin{itemize}
        \item \textbf{Monte Carlo:} Generate points near the minimum according to $p(\mathbf{w})$ and average the differences.
        \item \textbf{Hessian Eigenvectors:} Use directions with the largest eigenvalues to focus on key curvature components.
    \end{itemize}

    \begin{columns}[t]
        \column{0.35\textwidth}
        \centering
        \hspace*{-2cm}
        \includegraphics[width=0.95\textwidth]{img/LS_16.jpg}\\
        \hspace*{-2cm}
        \scriptsize \textit{Loss function}

        \column{0.47\textwidth}
        \centering
        \vspace*{-3.3cm}
        \hspace*{-2.2cm}
        \includegraphics[width=1.6\textwidth]{img/D_32.jpg}\\
        \hspace*{-1.2cm}
        \scriptsize \textit{$\Delta_k$ function}
    \end{columns}

\end{frame}
%----------------------------------------------------------------------------------------------------------
\end{document}

### 2025-04-17 paper/main.tex
Universal commits presents...

@@
 
-\title{Neural Networks Loss Landscape Convergence in Different Low-Dimensional Spaces}
+\title{Neural Networks Loss Landscape Convergence in Hessian Low-Dimensional Space}
 
@@
 
-We consider a $K$-class classification problem. Let $\bx \in \bX$ be an input and $y \in \mathcal Y = \{1,\dots,K\}$ its label. A
-neural network with parameters $\bw \in \bOmega \subset \mathbb R^P$ defines a mapping $f_{\bw} : \bX \to \mathbb R^K$. Given a dataset
+We consider a $S$-class classification problem. Let $\bx \in \bX$ be an input and $y \in \mathcal Y = \{1, \dots, S\}$ its label. A
+neural network with parameters $\bw \in \bOmega \subset \mathbb R^P$ defines a mapping $f_{\bw} : \bX \to \mathbb R^S$. Given a dataset
 $$
   \mathcal D =
-  \{(\bx_i, y_i)\}_{i=1}^m
+  \{(\bx_i, y_i)\}_{i=1}^N
 $$
@@
   $$
-  so that the full-sample loss is $\cL(\bw) = \cL_m(\bw)$.
+  so that the full-sample loss is $\cL(\bw) = \cL_N(\bw)$.
 \end{definition}
@@
   $$
-  and, similarly, the full-sample Hessian is $\bH(\bw) = \bH_m(\bw) = \nabla^2_{\bw} \cL_m(\bw)$.
+  and, similarly, the full-sample Hessian is $\bH(\bw) = \bH_N(\bw) = \nabla^2_{\bw} \cL_N(\bw)$.
 \end{definition}
@@
   $$
-    k^* =
-    \inf \{ k \mid \forall\, m \ge k : \; \Delta_m < \Delta\}.
+    K =
+    \inf \{ k \mid \forall\, k' \ge k : \; \Delta_{k'} < \Delta\}.
   $$
-  When $k \ge k^*$, adding further samples changes the loss landscape by less than $\Delta$.
+  When $k \ge K$, adding further samples changes the loss landscape by less than $\Delta$.
 \end{definition}
@@
     \int \bigl(\cL_{k+1}(\bw)-\cL_k(\bw)\bigr)^2\, p(\bw)\, d\bw =
-    \mathbb{E}_{p(\bw)} \left( \cL_{k+1}(\bw) - \cL_{k}(\bw) \right)^2 =
+    \mathbb{E} \left( \cL_{k+1}(\bw) - \cL_{k}(\bw) \right)^2 =
   $$
   $$
-    \mathbb D_{p(\bw)} \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr] +
-    \Bigl( \mathbb E_{p(\bw)} \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr] \Bigr)^2.
+    \mathbb D \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr] +
+    \Bigl( \mathbb E \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr] \Bigr)^2.
   $$
@@
   \bP^{\T}\, \bH_k(\bw^*)\, \bP =
-  \operatorname{diag} \bigl( \lambda_k^1, \dots, \lambda_k^d \bigr).
+  \operatorname{diag} \bigl( \lambda_k^{(1)}, \dots, \lambda_k^{(d)} \bigr).
 $$
@@
     \mathbb E \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr]^2 \approx
-    \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i)^2
-    + \Bigl( \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i) \Bigr)^2 \Biggr).
+    \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^d (\lambda_{k+1}^{(i)} - \lambda_k^{(i)})^2
+    + \Bigl( \sum_{i=1}^d (\lambda_{k+1}^{(i)} - \lambda_k^{(i)}) \Bigr)^2 \Biggr).
   $$
@@
 \begin{remark}
-  Assuming $\mathbb{E}_{p(\btheta)}(\btheta) = \mathbf{0}$ is natural when centering the sampling distribution at the minimum.
+  Assuming $\mathbb{E}(\btheta) = \mathbf{0}$ is natural when centering the sampling distribution at the minimum.
 \end{remark}
@@
   \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_random_1_2.pdf}}
-  \caption{Loss function landscape for small $k$, projected onto the two random directions.}
+  \caption{$\cL_k$ (left) and $\bigl[ \cL_{k+1}(\btheta) - \cL_k(\btheta) \bigr]^2$ (right) for small $k$, projected onto the two random vectors.}
   \label{fig:loss_random_small}
@@
   \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_random_-2_-1.pdf}}
-  \caption{Loss function landscape for maximum $k$, projected onto the two random directions.}
+  \caption{$\cL_k$ (left) and $\bigl[ \cL_{k+1}(\btheta) - \cL_k(\btheta) \bigr]^2$ (right) for maximum $k$, projected onto the two random vectors.}
   \label{fig:loss_random_big}
@@
 
-\paragraph{Loss Landscape for Small Sample Size}
-In Figure~\ref{fig:loss_random_small} ($k=1$), the loss surface is highly irregular, with sharp peaks and deep valleys.
+\paragraph{Loss Landscape for the Small Dataset}
+In Figure~\ref{fig:loss_random_small}, the loss surface is highly irregular, with sharp peaks and deep valleys.
 This irregularity indicates that with very few training samples, the model is extremely sensitive to small parameter perturbations.
@@
 
-\paragraph{Loss Landscape for Maximum Dataset Size}
-In contrast, Figure~\ref{fig:loss_random_big} (maximum $k$) shows a much smoother loss surface with gentle slopes. The overall loss
+\paragraph{Loss Landscape for the Full Dataset}
+In contrast, Figure~\ref{fig:loss_random_big} shows a much smoother loss surface with gentle slopes. The overall loss
 values are lower --- reflecting the averaging effect of more samples --- and the squared difference surface is correspondingly much smaller.
@@
   \Delta_k =
-  \mathbb{E}_{p(\bw)} \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr]^2
+  \mathbb{E} \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr]^2
 $$
@@
   \item \textbf{Sample directions.}
-        Starting from a fixed trained model with parameters $\bw^*$, draw $K$ independent random vectors
+        Starting from a fixed trained model with parameters $\bw^*$, draw $T$ independent random vectors
         $$
           \btheta^{(t)} \sim \mathcal{N} (\mathbf 0, \sigma^2 \mathbf I_d), \quad
-          t = 1, \dots, K.
+          t = 1, \dots, T.
         $$
@@
           \Delta_k \approx
-          \frac{1}{K} \sum_{t=1}^K d_t.
+          \frac{1}{T} \sum_{t=1}^T d_t.
         $$
@@
   \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_random_sigma_2_64.pdf}}
-  \caption{Delta function landscape varying the Gaussian variance $\sigma$.}
+  \caption{$\Delta_k$ (left) and $\Delta_k \cdot k^2$ (right), varying the Gaussian variance with fixed $d = 2$.}
   \label{fig:delta_random_sigma}
@@
   \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_random_dim_1_64.pdf}}
-  \caption{Delta function landscape varying the number of dimensions.}
+  \caption{$\Delta_k$ (left) and $\Delta_k \cdot k^2$ (right), varying the subspace dimension with fixed $\sigma = 1$.}
   \label{fig:delta_random_dim}
@@
 \paragraph{Projection Directions.}
-In the main experiment, we replace random directions with the $d=10$ eigenvectors of the Hessian at the local minimizer $\bw^*$ that
+In the main experiment, we replace random directions with the $d$ eigenvectors of the Hessian at the local minimizer $\bw^*$ that
 correspond to its largest eigenvalues.  Prior work \cite{sagun2018empirical} shows these top modes capture nearly all of the
@@
 
-Figures~\ref{fig:loss_eigen_small} and \ref{fig:loss_eigen_big} display 3D plots of $\cL_k(\btheta)$ (left) and $\bigl[\cL_{k+1}(\btheta)-\cL_k(\btheta)\bigr]^2$ (right), where $\btheta$ varies over the span of the top two Hessian eigenvectors.
+Figures~\ref{fig:loss_eigen_small} and \ref{fig:loss_eigen_big} display 3D plots of $\cL_k(\btheta)$ (left) and 
+$\bigl[ \cL_{k+1}(\btheta) - \cL_k(\btheta) \bigr]^2$ (right), where $\btheta$ varies over the span of the top two Hessian eigenvectors.
  The procedure mirrors the random‑direction case but uses these maximally informative axes.
@@
   \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_eigen_1_2.pdf}}
-  \caption{Loss landscape $\cL_k$ and squared difference for small $k$, projected onto the top two Hessian eigenvectors.}
+  \caption{$\cL_k$ (left) and $\bigl[ \cL_{k+1}(\btheta) - \cL_k(\btheta) \bigr]^2$ (right) for small $k$, projected onto the top two Hessian eigenvectors.}
   \label{fig:loss_eigen_small}
@@
   \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_eigen_-2_-1.pdf}}
-  \caption{Loss landscape $\cL_k$ and squared difference for maximum $k$, projected onto the top two Hessian eigenvectors.}
+  \caption{$\cL_k$ (left) and $\bigl[ \cL_{k+1}(\btheta) - \cL_k(\btheta) \bigr]^2$ (right) for maximum $k$, projected onto the top two Hessian eigenvectors.}
   \label{fig:loss_eigen_big}
@@
 
-In Figure~\ref{fig:loss_eigen_small} (small $k$, e.g.\ $k=1$), the loss surface projected onto the top two Hessian eigenvectors is
+\paragraph{Loss Landscape for the Small Dataset}
+In Figure~\ref{fig:loss_eigen_small}, the loss surface projected onto the top two Hessian eigenvectors is
 noticeably more structured and less erratic than in the random‐direction case. This confirms that these eigenvectors capture the
@@
 
-By contrast, Figure~\ref{fig:loss_eigen_big} (maximum $k$) shows a much smoother loss surface and a greatly reduced squared‐difference.
+\paragraph{Loss Landscape for the Full Dataset}
+By contrast, Figure~\ref{fig:loss_eigen_big} shows a much smoother loss surface and a greatly reduced squared‐difference.
 As the dataset grows, the loss changes less along the principal curvature directions, and the model’s local behavior stabilizes.
@@
   \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_eigen_1_10_64.pdf}}
-  \caption{Delta function landscape}
+  \caption{$\Delta_k$ (left) and $\Delta_k \cdot k^2$ (right)}
   \label{fig:delta_eigen}
@@
   \Delta_k \approx
-  \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^{d} (\lambda_{k+1}^i - \lambda_{k}^i)^2 +
-  \Bigl( \sum_{i=1}^{d} (\lambda_{k+1}^i - \lambda_{k}^i) \Bigr)^2 \Biggr),
+  \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^d (\lambda_{k+1}^{(i)} - \lambda_k^{(i)})^2
+  + \Bigl( \sum_{i=1}^d (\lambda_{k+1}^{(i)} - \lambda_k^{(i)}) \Bigr)^2 \Biggr).
 $$
-where $\{\lambda_k^i\}_{i=1}^d$ and $\{\lambda_{k+1}^i\}_{i=1}^d$ are the top $d$ Hessian eigenvalues at sample sizes $k$ and $k + 1$,
+where $\{\lambda_k^{(i)}\}_{i=1}^d$ and $\{\lambda_{k+1}^{(i)}\}_{i=1}^d$ are the top $d$ Hessian eigenvalues at sample sizes $k$ and $k + 1$,
 and $\sigma$ is the perturbation variance.
@@
   \subfloat{\includegraphics[width=1\textwidth]{img/delta_border_1_10_64.pdf}}
-  \caption{Delta function landscape}
+  \caption{Theoretical and Empirical $\Delta_k$}
   \label{fig:delta_eigen_comparison}
@@
 $$
-  \Delta_k \;=\; \mathbb{E}_{p(\btheta)}\bigl[L_{k+1}(\btheta)-L_k(\btheta)\bigr]^2,
+  \Delta_k =
+  \mathbb{E} \bigl[ \cL_{k+1}(\btheta) - \cL_k(\btheta) \bigr]^2,
 $$
@@
 $$
-  \Delta_k\approx\frac{\sigma^4}{4}\Bigl(2\sum_{i=1}^d(\lambda^i_{k+1}-\lambda^i_k)^2
-  +\bigl(\sum_{i=1}^d(\lambda^i_{k+1}-\lambda^i_k)\bigr)^2\Bigr)
+  \Delta_k \approx
+  \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^d (\lambda_{k+1}^{(i)} - \lambda_k^{(i)})^2
+  + \Bigl( \sum_{i=1}^d (\lambda_{k+1}^{(i)} - \lambda_k^{(i)}) \Bigr)^2 \Biggr).
 $$

### 2025-04-17 code/src/eigenvalues.py

@@ -17,7 +17,7 @@ class HessianEigenvector:
         self.eigenvectors = []
         self.eigenvalues = []
 
-    def get(self, k, num_iters=1000, tol=1e-7):
+    def get(self, k, num_iters=1000, tol=1e-6):
         if k <= len(self.eigenvectors):
             return self.eigenvalues[:k], self.eigenvectors[:k]
 

### 2025-04-17 code/src/visualize.py

@@ -41,23 +41,18 @@ class LossVisualizer:
             x = max(x, z_grid_bounds[0])
             return x
 
-        title = rf'$\mathcal{{L}}_{{{size1}}}$'
         zgrid = np.array([[bounds_func(np.mean(self.grid_loss[(x, y)][:size1])) for x in xs] for y in ys])
-
-        title_dif = rf'$(\mathcal{{L}}_{{{size2}}} - \mathcal{{L}}_{{{size1}}})^2$'
         zgrid_dif = np.array([[bounds_func(x) for x in r] for r in zgrid_dif])
 
         fig = plt.figure(figsize=(14, 6))
 
         ax1 = fig.add_subplot(1, 2, 1, projection='3d')
-        ax1.set_title(title)
         surf1 = ax1.plot_surface(xgrid, ygrid, zgrid, linewidth=0, antialiased=False, cmap=cm.get_cmap("coolwarm"),
                                  alpha=1.0)
         fig.colorbar(surf1, ax=ax1, shrink=0.5, aspect=5)
         ax1.view_init(40, 20)
 
         ax2 = fig.add_subplot(1, 2, 2, projection='3d')
-        ax2.set_title(title_dif)
         surf2 = ax2.plot_surface(xgrid, ygrid, zgrid_dif, linewidth=0, antialiased=False, cmap=cm.get_cmap("coolwarm"),
                                  alpha=1.0)
         fig.colorbar(surf2, ax=ax2, shrink=0.5, aspect=5)
@@ -78,13 +73,12 @@ class DeltaVisualizer:
         mult_coef = np.square(np.arange(1, len(deltas) + 1))
 
         fig, axs = plt.subplots(figsize=(14, 6), nrows=1, ncols=2)
-        fig.suptitle(rf'$\Delta_k$: $\sigma$ = {params['sigma']}, dimensions = {params['dim']}, K = {num_samples}')
 
         axs[0].plot(deltas)
         axs[1].plot(deltas * mult_coef)
 
-        ylabels = [r'$\mathbb{E}(\mathcal{L}_{k+1} - \mathcal{L}_{k})^2$',
-                   r'$\mathbb{E}(\mathcal{L}_{k+1} - \mathcal{L}_{k})^2 \cdot k^2$']
+        ylabels = [r'$\Delta_k$',
+                   r'$\Delta_k \cdot k^2$']
 
         axs[0].set(
             xlabel='k',
[…]

### 2025-04-24 review/nekhoroshkov.tex
Universal commits presents...

\documentclass[a4paper,12pt]{article}
\usepackage[T2A,T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[russian,english]{babel}
\usepackage{enumitem}
\usepackage{geometry}
\geometry{left=2cm,right=2cm,top=2cm,bottom=2cm}

\begin{document}

\pagestyle{empty}

\begin{center}
  {\Large\textbf{Review}}\\[1em]
  {\large\textit{Detecting Manual Alterations in Biological Image Data Using Contrastive Learning and Pairwise Image Comparison}}\\[0.5em]
  {\selectlanguage{russian}Nekhoroshkov G.}
\end{center}

\vspace{1em}

\begin{enumerate}[leftmargin=*, label=\arabic*.]
  \item Suggest increasing the font size of the axis tick labels on the AUC-ROC plots.
  \item Add references when mentioning figures in the text to facilitate easier navigation.
  \item Make the GitLab link clickable.
  \item Consider adding a section on future work and limitations; this would simplify any subsequent explorations.
  \item There is no discussion section, as indicated in the outline. I suggest adding one, since the conclusions about the 
  method and the chosen fixed hyperparameters in Section 5 are currently unclear.
  \item Half of the bibliography is not cited in the article. I recommend incorporating more of these references into the 
  related work section to give readers deeper insight and clearer connections.
  \item For clarity --- and because the code is already available on GitHub --- keep only the theoretical algorithm descriptions 
  in the main text.
  \item Not all cited articles include links to their arXiv pages or to the code. Please add these links in the bibliography 
  for easier reference.
\end{enumerate}

\vspace{1em}

\noindent\textbf{Overall opinion.} The problem statement and the motivation for this exploration are presented clearly. 
The author has surveyed numerous prior works and discussed their relevance thoughtfully. The theoretical section is easy to 
follow, and the descriptions of the base algorithms helped me understand both the core idea and the experimental setup. 
The results are displayed graphically, providing a strong sense of their effectiveness and potential applications. Once the 
minor fixes above are implemented, this paper could represent a significant advance in machine-learning–based medical 
detection technologies!

\vspace{2em}

\noindent Reviewer:\\
Nikitin A.

\end{document}

### 2025-04-24 slides/english/main.tex
Universal commits presents...

\documentclass{beamer}
\beamertemplatenavigationsymbolsempty
\usecolortheme{beaver}
\setbeamertemplate{blocks}[rounded=true, shadow=true]
\setbeamertemplate{footline}[page number]

\usepackage[utf8]{inputenc}
\usepackage[english,russian]{babel}
\usepackage{amssymb,amsfonts,amsmath,mathtext}
\usepackage{subfig}
\usepackage[all]{xy}
\usepackage{array}
\usepackage{multicol}
\usepackage{hyperref}
\usepackage{hhline}
\graphicspath{{fig/}{../fig/}}

%----------------------------------------------------------------------------------
\title[Loss Landscape Convergence]{Neural Networks Loss Landscape Convergence in Hessian Low-Dimensional Space}
\author[Tem Nikitin \textit{et al.}]{Tem Nikitin}
\institute{Moscow Institute of Physics and Technology}
\date{2025}
%----------------------------------------------------------------------------------
\begin{document}

% Slide 1: Title
\begin{frame}
    \thispagestyle{empty}
    \maketitle
\end{frame}

% Slide 2: Persona
\begin{frame}{Persona: Researcher (Data Scientist)}
    \begin{itemize}
        \item \textbf{Name:} Alex Researcher
        \item \textbf{Role:} Data Scientist / ML Researcher
        \item \textbf{Goals:} Understand when adding data stops improving training dynamics
        \item \textbf{Frustrations:} High computational cost of large datasets; unclear stopping criterion
        \item \textbf{Technologies:} Python, PyTorch, Hessian eigen-decompositions
    \end{itemize}
\end{frame}

% Slide 3: Background + Rising Action
\begin{frame}{User Scenario: Background \& Rising Action}
    \textbf{Background:}
    \begin{itemize}
        \item Modern neural nets achieve better accuracy with more data, but at high computational cost.
        \item Key question: \emph{when} does adding more samples give negligible improvement?
    \end{itemize}
    \textbf{Rising Action:}
    \begin{itemize}
        \item Project the high-dimensional parameter space onto top Hessian eigenvectors to capture principal curvature directions.
    \end{itemize}
\end{frame}

% Slide 4: Climax
\begin{frame}{User Scenario: Climax}
    \begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{img/loss_random_1_2.pdf}
        \caption*{3D loss surfaces $L_k$ and $(L_{k+1}-L_k)^2$ for small $k$.}
    \end{figure}
\end{frame}

% Slide 5: Falling Action
\begin{frame}{User Scenario: Falling Action}
    \begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{img/loss_eigen_-2_-1.pdf}
        \caption*{3D loss surfaces $L_k$ and $(L_{k+1}-L_k)^2$ for large $k$.}
    \end{figure}
\end{frame}

% Slide 6: Resolution
\begin{frame}{User Scenario: Resolution}
    \begin{figure}
        \centering
        \includegraphics[width=0.7\textwidth]{img/delta_border_1_10_64.pdf}
        \caption*{Decay of $\Delta_k$ and $\Delta_k\cdot k^2$.}
    \end{figure}
\end{frame}

% Slide 7: Prototype – Methodology
\begin{frame}{Prototype: Methodology}
    \begin{itemize}
        \item Compute top $d$ eigenvectors of Hessian $H(w^*)$ via iterative power method.
        \item Form low-dimensional coordinate $\theta$ such that $w = w^* + P\theta$, $P=[e_1,\dots,e_d]$.
        \item Estimate $\Delta_k=E[(L_{k+1}-L_k)^2]$ via Monte Carlo in this subspace.
    \end{itemize}
\end{frame}

% Slide 8: Prototype – Formula
\begin{frame}{Prototype: Theoretical Estimate}
    \vspace{1em}
    \centering
    $\displaystyle
        \Delta_k\approx\frac{\sigma^4}{4}\Biggl(2\sum_{i=1}^d(\lambda_i^{(k+1)}-\lambda_i^{(k)})^2+\bigl(\sum_{i=1}^d(\lambda_i^{(k+1)}-\lambda_i^{(k)})\bigr)^2\Biggr)$
    \vspace{1em}
\end{frame}

\end{document}

### 2025-04-24 code/src/visualize.py

@@ -44,7 +44,7 @@ class LossVisualizer:
         zgrid = np.array([[bounds_func(np.mean(self.grid_loss[(x, y)][:size1])) for x in xs] for y in ys])
         zgrid_dif = np.array([[bounds_func(x) for x in r] for r in zgrid_dif])
 
-        fig = plt.figure(figsize=(14, 6))
+        fig = plt.figure(figsize=(15, 6))
 
         ax1 = fig.add_subplot(1, 2, 1, projection='3d')
         surf1 = ax1.plot_surface(xgrid, ygrid, zgrid, linewidth=0, antialiased=False, cmap=cm.get_cmap("coolwarm"),

### 2025-04-24 review/nekhoroshkov.tex
Universal commits presents...

@@
 \documentclass[a4paper,12pt]{article}
-\usepackage[T2A,T1]{fontenc}
 \usepackage[utf8]{inputenc}
-\usepackage[russian,english]{babel}
-\usepackage{enumitem}
+\usepackage[english]{babel}
+\usepackage[T2A]{fontenc}
 \usepackage{geometry}
@@
   {\large\textit{Detecting Manual Alterations in Biological Image Data Using Contrastive Learning and Pairwise Image Comparison}}\\[0.5em]
-  {\selectlanguage{russian}Nekhoroshkov G.}
+  {Nekhoroshkov G.}
 \end{center}
 
-\vspace{1em}
-
-\begin{enumerate}[leftmargin=*, label=\arabic*.]
+\begin{enumerate}
   \item Suggest increasing the font size of the axis tick labels on the AUC-ROC plots.
@@
 
-\vspace{1em}
-
 \noindent\textbf{Overall opinion.} The problem statement and the motivation for this exploration are presented clearly.
@@
 
-\vspace{2em}
-
-\noindent Reviewer:\\
+\begin{flushright}
+  \noindent\textbf{Reviewer:}\\
   Nikitin A.
+\end{flushright}
 

### 2025-04-30 paper/main.tex
Universal commits presents...

@@
   \item Constructing a Hessian‑based projection that captures the principal curvature directions.
-  \item Applying Monte Carlo and other sampling methods to identify when the loss landscape stabilizes as dataset size grows.
+  \item Applying Monte Carlo method to identify when the loss landscape stabilizes as dataset size grows.
   \item Visualizing the projected loss surface around local minima to illustrate these changes.
@@
 
-\textbf{Organization.} The rest of the paper is structured as follows. Section~2 reviews related work. Section~3 introduces notation
-and preliminaries. Section~4 derives theoretical bounds on Hessian spectra and the loss‐difference norm. Section~5 presents our
-empirical studies. We discuss the implications in Section~6 and conclude in Section~7. Additional experiments and proofs are
-provided in Appendix~A.
+\textbf{Organization.} The rest of the paper is structured as follows. Section 2 reviews related work. Section 3 introduces notation
+and preliminaries. Section 4 derives theoretical bounds on Hessian spectra and the loss‐difference norm. Section 5 presents our
+empirical studies. We discuss the implications in Section 6 and conclude in Section 7. Additional experiments and proofs are
+provided in Appendix A.
 
@@
 We consider a $S$-class classification problem. Let $\bx \in \bX$ be an input and $y \in \mathcal Y = \{1, \dots, S\}$ its label. A
-neural network with parameters $\bw \in \bOmega \subset \mathbb R^P$ defines a mapping $f_{\bw} : \bX \to \mathbb R^S$. Given a dataset
+neural network with parameters $\bw \in \bOmega \subset \mathbb R^M$ defines a mapping $f_{\bw} : \bX \to \mathbb R^S$. Given a dataset
 $$
@@
 $$
-of $m$ i.i.d. samples, and a per-sample loss $\ell \bigl( f_{\bw}(\bx_i), y_i \bigr)$ (e.g. cross–entropy), we define
+of $N$ i.i.d. samples, and a per-sample loss $\ell \bigl( f_{\bw}(\bx_i), y_i \bigr)$ (e.g. cross–entropy), we define
 $$
@@
 \begin{definition}
-  Fix a tolerance $\Delta > 0$. The \emph{sufficient sample size} is
+  Fix a tolerance $\Delta > 0$. The \emph{$\Delta$-sufficient sample size} is
   $$
@@
 
-We make the following standard assumptions to simplify our analysis around a local minimizer~$\bw^*$.
+We make the following standard assumptions to simplify our analysis around a local minimizer $\bw^*$.
 
@@
 
-\section{Method: Projection onto Dominant Eigen-Directions and Loss Landscape Approximation}
+\section{Method: Projection onto Dominant Eigen-Directions and Loss Landscape Approximation}\label{sec:meth}
 
@@
 of the parameter space $\bOmega$. To reduce this cost, we project the parameters onto a lower‑dimensional subspace spanned by the
-$d$ dominant eigenvectors of the Hessian at a local minimum $\bw^*$.
+$D$ dominant eigenvectors of the Hessian at a local minimum $\bw^*$.
 
@@
 only a few eigenvalues are significantly above zero, while the rest are negligible. Therefore, directions corresponding to small
-eigenvalues contribute little to loss variation.  Retaining only the top $d$ eigenvectors yields a compact yet informative representation
+eigenvalues contribute little to loss variation. Retaining only the top $D$ eigenvectors yields a compact yet informative representation
 of the most important curvature directions.
 
-For illustration, Figure~\ref{fig:evgen} plots the top few eigenvalues of $\bH(\bw^*)$ in descending order. We compute these dominant
+For illustration, Figure \ref{fig:evgen} plots the top few eigenvalues of $\bH(\bw^*)$ in descending order. We compute these dominant
 eigenvalues using the iterative power‐method described at \cite{hessian-eigenthings}.
@@
   \bP =
-  \bigl[ \mathbf{e}_1, \dots, \mathbf{e}_d \bigr]
+  \bigl[ \mathbf{e}_1, \dots, \mathbf{e}_D \bigr]
 $$
-be the matrix whose columns are the top $d$ eigenvectors.  We write any parameter vector as
+be the matrix whose columns are the top $D$ eigenvectors. We write any parameter vector as
 $$
@@
   \bw^* + \bP\, \btheta, \quad
-  \btheta \in \mathbb R^d.
+  \btheta \in \mathbb R^D.
 $$
@@
   \bP^{\T}\, \bH_k(\bw^*)\, \bP =
-  \operatorname{diag} \bigl( \lambda_k^{(1)}, \dots, \lambda_k^{(d)} \bigr).
+  \operatorname{diag} \bigl( \lambda_k^{(1)}, \dots, \lambda_k^{(D)} \bigr).
 $$
@@
 
-\begin{theorem}[Approximate $\Delta_k$ via Eigenvalues]
-  Under the assumption that $\btheta \sim \mathcal N(0, \sigma^2 \mathbf I_d)$, the change in loss can be estimated by
+\begin{theorem}[Approximate $\Delta_k$ via Eigenvalues]\label{th:th_1}
+  Under the assumption that $\btheta \sim \mathcal N(0, \sigma^2 \mathbf I_D)$, the change in loss can be estimated by
   $$
@@
     \mathbb E \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr]^2 \approx
-    \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^d (\lambda_{k+1}^{(i)} - \lambda_k^{(i)})^2
-    + \Bigl( \sum_{i=1}^d (\lambda_{k+1}^{(i)} - \lambda_k^{(i)}) \Bigr)^2 \Biggr).
+    \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{d=1}^D (\lambda_{k+1}^{(d)} - \lambda_k^{(d)})^2
+    + \Bigl( \sum_{d=1}^D (\lambda_{k+1}^{(d)} - \lambda_k^{(d)}) \Bigr)^2 \Biggr).
   $$
@@
 
-This formula uses only the top $d$ eigenvalues, so it is efficient to compute even for very large models.
+This formula uses only the top $D$ eigenvalues, so it is efficient to compute even for very large models.
 In practice, it provides an upper bound on the empirical $\Delta_k$ and thus a reliable criterion for determining when the
@@
 
-\section{Experiments}
+\section{Experiments}\label{sec:exp}
 
@@
 \begin{enumerate}
-  \item \textbf{Random‐direction visualization:} As an initial sanity check, we project the parameters onto two random directions and
-        plot the resulting 3D loss landscapes and their squared differences.
-  \item \textbf{Hessian‐based analysis:} We then repeat the visualization and compute $\Delta_k$ by projecting onto the top two
-        (and up to $d = 10$) eigenvectors of the Hessian.  This directly tests our eigenvalue‐based approximation and the convergence behavior
-        predicted by theory.
+  \item \textbf{Random‐direction visualization \ref{sec:exp:pre}:} As an initial sanity check, we project the parameters onto two random directions and
+        plot the resulting 3D loss landscapes, their squared differences, and the corresponding $\Delta_k$ for various settings.
+  \item \textbf{Hessian‐based analysis \ref{sec:exp:hes}:} We repeat the above visualization using the top eigenvectors of the Hessian. This directly
+        tests our hypothesis regarding the dominant role of the largest Hessian eigenvalues.
+  \item \textbf{Main experiment \ref{sec:exp:main}:} We compute both the theoretical and empirical values of $\Delta_k$ and compare them. This evaluates
+        the accuracy of our approximation predicted by theory.
 \end{enumerate}
 
-All code and detailed configuration files are available in our GitHub repository.
-
-\subsection{Preliminary Experiment: Projection onto Random Directions}
+All code and detailed configuration files are available in our GitHub
+repository \footnote{\href{https://github.com/intsystems/2025-Project-183}{GitHub}}.
+\subsection{Preliminary Experiment: Projection onto Random Directions}\label{sec:exp:pre}
 \subsubsection{Visualizing the Loss Landscape}
@@
 
-\begin{enumerate}
-  \item \textbf{Grid Sampling.}
-        We form a uniform grid of $(\alpha, \beta)$ pairs in $[-1,1]\times[-1,1]$ with a fixed step size.
-  \item \textbf{Loss Computation.}
-        For each $(\alpha,\beta)$, we set $\bw=\bw(\alpha,\beta)$, keep the base weights $\bw^*$ fixed, and compute both
-        $\cL_k(\bw)$ and $\cL_{k+1}(\bw)$.
-  \item \textbf{Aggregated Metrics.}
-        \begin{itemize}
-          \item The surface $\cL_k(\alpha,\beta)$ visualizes how the mean loss changes over the grid.
-          \item The surface $\bigl( \cL_{k+1} (\alpha, \beta) - \cL_k (\alpha, \beta) \bigr)^2$ highlights where adding one sample
-                causes the largest change.
-        \end{itemize}
-  \item \textbf{3D Visualization.}
-        We render two side‐by‐side 3D plots:
-        \begin{itemize}
-          \item \emph{Left:} $\cL_k (\alpha, \beta)$ vs. $(\alpha, \beta)$.
-          \item \emph{Right:} $\bigl( \cL_{k+1} (\alpha, \beta) - \cL_k (\alpha, \beta) \bigr)^2$ vs. $(\alpha, \beta)$.
-        \end{itemize}
-\end{enumerate}
+First, a uniform grid of $(\alpha,\beta)$ pairs on the domain $[-1,1]\times[-1,1]$ with fixed step size $h$ is constructed.
+For each grid point $(\alpha,\beta)$, we set $\mathbf{w} = \mathbf{w}(\alpha,\beta)$, keeping the base weights
+$\mathbf{w}^*$ fixed, and compute the losses $\mathcal{L}_k(\mathbf{w})$ and $\mathcal{L}_{k+1}(\mathbf{w})$.
+This produces two scalar fields over the grid: the first, $\mathcal{L}_k(\alpha,\beta)$, illustrates how the mean loss
+varies across the parameter space, while the second,
+$$\bigl(\mathcal{L}_{k+1}(\alpha,\beta) - \mathcal{L}_k(\alpha,\beta)\bigr)^2,$$ highlights the regions where adding a single
+sample induces the largest change. Finally, both fields are visualized as two adjacent 3D plots.
 
@@
 \paragraph{Loss Landscape for the Small Dataset}
-In Figure~\ref{fig:loss_random_small}, the loss surface is highly irregular, with sharp peaks and deep valleys.
+In Figure \ref{fig:loss_random_small}, the loss surface is highly irregular, with sharp peaks and deep valleys.
 This irregularity indicates that with very few training samples, the model is extremely sensitive to small parameter perturbations.
@@
 \paragraph{Loss Landscape for the Full Dataset}
-In contrast, Figure~\ref{fig:loss_random_big} shows a much smoother loss surface with gentle slopes. The overall loss
+In contrast, Figure \ref{fig:loss_random_big} shows a much smoother loss surface with gentle slopes. The overall loss
 values are lower --- reflecting the averaging effect of more samples --- and the squared difference surface is correspondingly much smaller.
@@
 $$
-via a Monte Carlo approximation using Gaussian perturbations in parameter space. The steps are:
+via a Monte Carlo approximation using Gaussian perturbations in parameter space.
 
-\begin{enumerate}
-  \item \textbf{Sample directions.}
-        Starting from a fixed trained model with parameters $\bw^*$, draw $T$ independent random vectors
+Starting from a fixed trained model to get $\mathbf{w}^*$, we draw $T$ independent random direction vectors
 $$
-          \btheta^{(t)} \sim \mathcal{N} (\mathbf 0, \sigma^2 \mathbf I_d), \quad
-          t = 1, \dots, T.
+  \btheta^{(t)} \sim \mathcal{N}(\mathbf{0}, \sigma^2 \mathbf{I}_d), \quad t = 1, \dots, T.
 $$
+For each batch, we form a perturbed parameter set $\mathbf{w}^{(t)} = \mathbf{w}^* + \btheta^{(t)}$,
+and evaluate the losses $\mathcal{L}_k(\mathbf{w}^{(t)})$ for each $k = 1, \, \dots, \, N$,
+after what compute the squared difference on given values:
+$$
+  d_t = \bigl[\mathcal{L}_{k+1}(\mathbf{w}^{(t)}) - \mathcal{L}_k(\mathbf{w}^{(t)})\bigr]^2.
+$$
+We then approximate the average change
+$\Delta_k \approx \frac{1}{T} \sum_{t=1}^T d_t$,
+noting that increasing $T$ reduces the variance of this Monte Carlo estimator. Finally, we plot $\Delta_k$ against $k$ to observe its decay
+and include the rescaled quantity $\Delta_k \cdot k^2$ to highlight the expected $\mathcal{O}(1/k)$ convergence rate.
 
-  \item \textbf{Perturb parameters.}
-        For each sample, form a perturbed parameter set
-        $$
-          \bw^{(t)} =
-          \bw^* + \btheta^{(t)}.
-        $$
-
-  \item \textbf{Evaluate loss differences.}
-        Compute
-        $$
-          \cL_k \bigl( \bw^{(t)} \bigr) \quad
-          \text{and} \quad
-          \cL_{k+1} \bigl( \bw^{(t)} \bigr),
-        $$
-        then calculate the squared difference
-        $$
-          d_t =
-          \bigl[ \cL_{k+1} (\bw^{(t)}) - \cL_k (\bw^{(t)}) \bigr]^2.
-        $$
-
-  \item \textbf{Compute Monte Carlo estimate.}
-        Approximate
-        $$
-          \Delta_k \approx
-          \frac{1}{T} \sum_{t=1}^T d_t.
-        $$
-        Increasing $K$ reduces the variance of this estimator.
-
-  \item \textbf{Plot results.}
-        We plot $\Delta_k$ versus $k$ to observe its decay. We also plot the scaled quantity
-        $$
-          \Delta_k \cdot k^2
-        $$
-        alongside to highlight an $\mathcal O(1 / k)$ convergence rate.
-\end{enumerate}
-
-Figures~\ref{fig:delta_random_sigma} and \ref{fig:delta_random_dim} present the behavior of the function
+Figures \ref{fig:delta_random_sigma} and \ref{fig:delta_random_dim} present the behavior of the function
 $$
@@
 \begin{itemize}
-  \item Figure~\ref{fig:delta_random_sigma}: varying the Gaussian variance $\sigma$.
-  \item Figure~\ref{fig:delta_random_dim}: varying the subspace dimension $d$.
+  \item Figure \ref{fig:delta_random_sigma}: varying the Gaussian variance $\sigma$.
+  \item Figure \ref{fig:delta_random_dim}: varying the subspace dimension $D$.
 \end{itemize}
@@
   \hspace*{-2.6cm}
-  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_random_sigma_2_64.pdf}}
+  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_random_sigma_2_1024.pdf}}
   \caption{$\Delta_k$ (left) and $\Delta_k \cdot k^2$ (right), varying the Gaussian variance with fixed $d = 2$.}
@@
   \hspace*{-2.6cm}
-  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_random_dim_1_64.pdf}}
+  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_random_dim_1_1024.pdf}}
   \caption{$\Delta_k$ (left) and $\Delta_k \cdot k^2$ (right), varying the subspace dimension with fixed $\sigma = 1$.}
@@
 \paragraph{Comparing Different Variances \texorpdfstring{$\sigma$}{sigma}}
-Figure~\ref{fig:delta_random_sigma} shows that varying the Gaussian variance $\sigma$ produces no substantial change in the shape or
-magnitude of the $\Delta_k$ curves. This insensitivity likely stems from using random projection directions: since they do not align
-with the dominant curvature axes of the Hessian, scaling those directions has little impact on the estimated loss differences.
+Figure \ref{fig:delta_random_sigma} shows that a high Gaussian variance $\sigma$ produces a significant change in both the shape
+and magnitude of the $\Delta_k$ curves. This occurs because, as we move away from the local minimum, the value of the loss
+function changes. However, this change is modest: the insensitivity likely stems from using random projection directions,
+which do not align with the dominant curvature axes of the Hessian, so scaling those directions has little impact on the estimated
+loss differences. Moreover, variances not exceeding one yield very similar effects, so we choose $\sigma = 1$ for simplicity in our
+calculations.
 
 \paragraph{Comparing Different Subspace Dimensions}
-Figure~\ref{fig:delta_random_dim} shows that, when using random directions, increasing the projection dimension $d$ does not
+Figure \ref{fig:delta_random_dim} shows that, when using random directions, increasing the projection dimension $D$ does not
 significantly alter the overall shape of $\Delta_k$. This suggests that only a few directions dominate the variation --- motivating our
@@
 
-In all cases, the right‐hand plots (showing $\Delta_k \cdot k^2$) exhibit a roughly $\mathcal O(1 / k)$ decay, confirming that the
+In all cases, the right‐hand plots (showing $\Delta_k \cdot k^2$) exhibit a roughly $\mathcal O(1 / k^2)$ decay, confirming that the
 incremental effect of adding one more sample diminishes proportionally to $1 / k$ as the dataset grows.
@@
 
-\subsection{Moving Forward: Projection onto Dominant Hessian Eigenvectors}
+\subsection{Moving Forward: Projection onto Dominant Hessian Eigenvectors}\label{sec:exp:hes}
 
 \paragraph{Projection Directions.}
-In the main experiment, we replace random directions with the $d$ eigenvectors of the Hessian at the local minimizer $\bw^*$ that
+In the main experiment, we replace random directions with the $D$ eigenvectors of the Hessian at the local minimizer $\bw^*$ that
 correspond to its largest eigenvalues. Prior work \cite{sagun2018empirical} shows these top modes capture nearly all of the
-non‑negligible curvature, and Figure~\ref{fig:evgen} confirms a rapid eigenvalue decay. Fixing $d=10$ balances computational cost and
-fidelity, and our tests on Fashion‑MNIST and CIFAR demonstrate that this choice suffices even for complex architectures.
+non‑negligible curvature, and Figure \ref{fig:evgen} confirms a rapid eigenvalue decay. Fixing $D = 10$ balances computational cost and
+fidelity, and our tests on Fashion‑MNIST demonstrate that this choice suffices even for complex architectures.
 
@@
 
-Figures~\ref{fig:loss_eigen_small} and \ref{fig:loss_eigen_big} display 3D plots of $\cL_k(\btheta)$ (left) and 
+Figures \ref{fig:loss_eigen_small} and \ref{fig:loss_eigen_big} display 3D plots of $\cL_k(\btheta)$ (left) and
 $\bigl[ \cL_{k+1}(\btheta) - \cL_k(\btheta) \bigr]^2$ (right), where $\btheta$ varies over the span of the top two Hessian eigenvectors.
@@
 \paragraph{Loss Landscape for the Small Dataset}
-In Figure~\ref{fig:loss_eigen_small}, the loss surface projected onto the top two Hessian eigenvectors is
+In Figure \ref{fig:loss_eigen_small}, the loss surface projected onto the top two Hessian eigenvectors is
 noticeably more structured and less erratic than in the random‐direction case. This confirms that these eigenvectors capture the
@@
 \paragraph{Loss Landscape for the Full Dataset}
-By contrast, Figure~\ref{fig:loss_eigen_big} shows a much smoother loss surface and a greatly reduced squared‐difference.
+By contrast, Figure \ref{fig:loss_eigen_big} shows a much smoother loss surface and a greatly reduced squared‐difference.
 As the dataset grows, the loss changes less along the principal curvature directions, and the model’s local behavior stabilizes.
@@
   \hspace*{-2.6cm}
-  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_eigen_1_10_64.pdf}}
+  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_eigen_1_10_1024.pdf}}
   \caption{$\Delta_k$ (left) and $\Delta_k \cdot k^2$ (right)}
@@
 
-Figure~\ref{fig:delta_eigen} presents the Monte Carlo estimate of $\Delta_k$ in the subspace spanned by the top $d = 10$
+Figure \ref{fig:delta_eigen} presents the Monte Carlo estimate of $\Delta_k$ in the subspace spanned by the top $d = 10$
 Hessian eigenvectors. Compared to the random‐direction experiment, $\Delta_k$ here has a higher value --- reflecting the
 large curvature captured by these eigenvectors. The right‐hand plot of $\Delta_k\cdot k^2$
-confirms an approximate $O(1 / k)$ convergence rate. These observations show that projecting onto the leading eigen-directions not only
+confirms an approximate $\mathcal O(1 / k^2)$ convergence rate. These observations show that projecting onto the leading eigen-directions not only
 highlights the greatest loss sensitivity but also yields a reliable upper bound on how quickly the landscape stabilizes as more data are
@@
 
-\subsection{Main Experiment: Theoretical vs. Empirical \texorpdfstring{$\Delta_k$}{Delta k}}
+\subsection{Main Experiment: Theoretical vs. Empirical \texorpdfstring{$\Delta_k$}{Delta k}}\label{sec:exp:main}
 
@@
 
-We approximate $\Delta_k$ by
+As the Theorem \ref{th:th_1} says, we approximate $\Delta_k$ by
 $$
   \Delta_k \approx
-  \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^d (\lambda_{k+1}^{(i)} - \lambda_k^{(i)})^2
-  + \Bigl( \sum_{i=1}^d (\lambda_{k+1}^{(i)} - \lambda_k^{(i)}) \Bigr)^2 \Biggr).
+  \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{d=1}^D (\lambda_{k+1}^{(d)} - \lambda_k^{(d)})^2
+  + \Bigl( \sum_{i=1}^D (\lambda_{k+1}^{(d)} - \lambda_k^{(d)}) \Bigr)^2 \Biggr).
 $$
-where $\{\lambda_k^{(i)}\}_{i=1}^d$ and $\{\lambda_{k+1}^{(i)}\}_{i=1}^d$ are the top $d$ Hessian eigenvalues at sample sizes $k$ and $k + 1$,
+where $\{\lambda_k^{(i)}\}_{i=1}^d$ and $\{\lambda_{k+1}^{(i)}\}_{i=1}^d$ are the top $D$ Hessian eigenvalues at sample sizes $k$ and $k + 1$,
 and $\sigma$ is the perturbation variance.
 
-\begin{enumerate}
-  \item \textbf{Model \& subspace size.} Train a network to obtain $\bw^*$. Choose $d$ based on the rapid eigenvalue decay in
-        Figure~\ref{fig:evgen}.
-  \item \textbf{Eigenvalue computation.} For each $k$ in the range of interest (typically the largest $K$ sizes), compute the top
-        $d$ eigenvalues via the power method.
-  \item \textbf{Theoretical $\Delta_k$.} Plug these eigenvalues into the formula above to obtain a sequence of theoretical estimates.
-  \item \textbf{Empirical $\Delta_k$.} Independently estimate $\Delta_k$ by Monte Carlo sampling of perturbations along the same
-        top‐$d$ eigenvectors as it was in the previous experiment.
-  \item \textbf{Comparison.} Plot both theoretical and empirical $\Delta_k$ on the same axes to assess how tightly the theory bounds
-        the observed values.
-\end{enumerate}
+We compute both empirical and theoretical estimates of $\Delta_k$. First, we train a model instance as described above.
+For each batch $k$, we apply the power‐iteration method to the model to extract the top $D$ Hessian eigenvalues
+$\{\lambda_k^{(d)}\}_{d=1}^D$. The theoretical estimate of $\Delta_k$ is then obtained by substituting these eigenvalues into our
+approximation. To reduce high-frequency fluctuations, we apply a five‐point moving average to the sequence:
+$\frac{1}{5}\bigl(\Delta_{k-2} + \Delta_{k-1} + \Delta_k + \Delta_{k+1} + \Delta_{k+2}\bigr)$ for each $k$,
+where, at the boundaries, any index outside $[1,N]$ is replaced by the nearest valid endpoint (i.e.\ we replicate the border value),
+which is equivalent to convolution with constant‐value padding.
+
+In parallel, we compute empirical estimates of $\Delta_k$ via Monte Carlo sampling of Gaussian perturbations along the same
+top-$D$ eigenvectors, following the procedure described above. Finally, we plot the
+theoretical and empirical $\Delta_k$ on the same axes, allowing us to assess how closely the eigenvalue-based bound tracks the observed
+convergence of the loss landscape.
 
 \begin{figure}[!htbp]
-  \subfloat{\includegraphics[width=1\textwidth]{img/delta_border_1_10_64.pdf}}
-  \caption{Theoretical and Empirical $\Delta_k$}
+  \subfloat{\includegraphics[width=1\textwidth]{img/delta_border_1_10_1024.pdf}}
+  \caption{Empirical and Theoretical $\Delta_k$}
   \label{fig:delta_eigen_comparison}
@@
 
-Figure~\ref{fig:delta_eigen_comparison} shows the theoretical estimate of $\Delta_k$ (dashed line) alongside the empirical Monte Carlo
+Figure \ref{fig:delta_eigen_comparison} shows the theoretical estimate of $\Delta_k$ (dashed line) alongside the empirical Monte Carlo
 measurements (solid line). Across all values of $k$, the theory curve remains above the empirical one, demonstrating that our
@@
 
+Now, to eliminate the explicit batch index $k$, we plot $\Delta_k$ directly against $\cL_k$ and apply the same moving average
+for a smoother, more stable curve.
+
+Figure \ref{fig:delta_loss} shows the relationship between the loss $\cL_k(\bw^*)$ and the corresponding change $\Delta_k(\bw^*)$.
+In the top panel, each blue dot represents the empirical estimate of $\Delta_k$ versus $\cL_k$, and the red line is the best-fit
+linear trend. The bottom panel shows the analogous theoretical estimates, again smoothed and fitted (green line). In both cases,
+there is a strong positive correlation: higher loss values correspond to larger observed and predicted changes. Although some
+scatter remains due to sampling noise, the overall linear relationship validates our hypothesis that the eigenvalue-based
+bound $\Delta$ captures the dominant curvature effects of the Hessian. Consequently, the $\Delta$-sufficient condition can effectively
+limit changes in the loss function.
+
+\begin{figure}[!htbp]
+  \subfloat{\includegraphics[width=1\textwidth]{img/delta_loss_1_10_1024.pdf}}
+  \caption{Empirical and Theoretical $\Delta_k$}
+  \label{fig:delta_loss}
+\end{figure}
+
+We are now ready to propose an algorithm to determine a $\Delta$-sufficient dataset size. Both theoretically and empirically, once
+$\Delta_k$ falls below a prescribed tolerance $\Delta$, the loss $\cL_k$ changes negligibly. Thus we say that
+for a tolerance $\Delta > 0$. The \emph{$\Delta$-sufficient sample size} is
+$$
+  K =
+  \inf \{ k \mid \forall\, k' \ge k : \; \Delta_{k'} < \Delta\}.
+$$
+When $k \ge K$, adding further samples changes the loss landscape by less than $\Delta$.
+Smaller values of $\Delta$ correspond to higher accuracy. In our MNIST experiments, we found that
+$\Delta=0.025$ yields $k=100$ and $\cL_k\approx0.013$.
+
+Table~\ref{tab:results} summarizes the key results of our $\Delta$-sufficient criterion applied to MNIST and Fashion-MNIST.
+For each dataset and model architecture, we report the tolerance $\Delta$, the number of batches $k$ required to reach
+$\Delta$-sufficiency, the resulting loss $\mathcal{L}_k$, and the total runtime.
+
+\begin{table}[ht]
+  \centering
+  \caption{Results of $\Delta$-sufficient experiments on MNIST and Fashion-MNIST}\label{tab:results}
+  \begin{tabular}{llcccc}
+    \toprule
+    Dataset       & Model             & $\Delta$ & $k$ & $\mathcal{L}_k$ & Time (s) \\
+    \midrule
+    MNIST         & Single-layer MLP  & 0.025    & 100 & 0.013           & 2000     \\
+                  & Multi-layer MLP   & 0.025    & 40  & 0.010           & 3500     \\
+                  & Convolutional CNN & 0.025    & 60  & 0.024           & 1800     \\
+    \midrule
+    Fashion-MNIST & Single-layer MLP  & 0.030    & 120 & 0.020           & 2100     \\
+                  & Multi-layer MLP   & 0.030    & 90  & 0.017           & 4400     \\
+                  & Convolutional CNN & 0.030    & 70  & 0.015           & 2400     \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+Since our theoretical estimate $\Delta_k$ consistently bounds the
+empirical $\Delta_k$ and is much cheaper to compute, we employ the following algorithm for stopping criterion:
+
+\begin{algorithm}[H]
+  \caption{Determine $\Delta$-sufficient dataset size}\label{alg:algo}
+  \begin{algorithmic}[1]
+    \Require target tolerance $\Delta>0$, batch size $B$
+    \State $\mathcal{D}\gets \emptyset,\;k\gets 0$
+    \Repeat
+    \State $k\gets k+1$
+    \State Sample a new batch of $B$ examples and append to $\mathcal{D}$
+    \State Train or update the model on $\mathcal{D}$
+    \State Compute the top-$D$ Hessian eigenvalues and form
+    $\Delta_k$ estimation via the eigenvalue-based approximation:
+    $$
+      \Delta_k \approx
+      \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{d=1}^D (\lambda_{k+1}^{(d)} - \lambda_k^{(d)})^2
+      + \Bigl( \sum_{i=1}^D (\lambda_{k+1}^{(d)} - \lambda_k^{(d)}) \Bigr)^2 \Biggr).
+    $$
+    \Until{$\Delta_k > \Delta$}
+    \State $\textbf{return }|\mathcal{D}|$
+  \end{algorithmic}
+\end{algorithm}
+
+Moreover, in addition to determining the dataset size $K$, the procedure also yields a fully trained model as a by‐product of the
+iterative updates.
+
+This procedure incrementally grows the dataset until the eigenvalue-based estimate $\Delta_k$ falls below the prescribed
+tolerance $\Delta$, at which point further additions have a negligible effect on the loss. The dominant cost per iteration is the
+extraction of the top-$D$ Hessian eigenvalues via power iteration, which requires $\mathcal O(I\,D\,C_{\mathrm{Hv}})$ work where $I$
+is the number of power‐iteration steps and $C_{\mathrm{Hv}}$ is the cost of a single Hessian–vector product (comparable to one
+gradient evaluation). Since $D\ll M$ and $I$ is typically small (e.g.\ $I\approx 20$), the extra overhead
+remains linear in the number of mini-batches and is modest compared to full second-order methods. Thus, our algorithm offers an
+efficient, theoretically grounded approach to determining a $\Delta$-sufficient dataset size with minimal computational burden.
+
 \section{Discussion}\label{sec:disc}
 
-The gap between the theoretical and empirical curves arises because the estimate uses only the top $d$ eigenvalues.
-Therefore, focusing on the leading modes yields a reliable and efficient criterion for determining when
-adding more data has negligible effect on the local loss geometry.
+The gap between the theoretical and empirical curves arises because the estimate uses only the top $D$ eigenvalues. Therefore, focusing on
+the leading modes yields a reliable and efficient criterion for determining when adding more data has negligible effect on the local loss
+geometry.
 
 Our Hessian‐projection framework provides a clear picture of how a neural network’s loss landscape evolves with training–set size.
-By restricting attention to the $d$ principal curvature directions (the top Hessian eigenvectors), we obtain far more structured 3D
-loss surfaces and a sharply decaying landscape‐difference measure
+By restricting attention to the $D$ principal curvature directions (the top Hessian eigenvectors), we obtain far more structured
+3D loss surfaces and a sharply decaying landscape‐difference measure
 $$
   \Delta_k =
-  \mathbb{E} \bigl[ \cL_{k+1}(\btheta) - \cL_k(\btheta) \bigr]^2,
+  \mathbb{E}\bigl[\cL_{k+1}(\btheta) - \cL_k(\btheta)\bigr]^2,
 $$
-which empirically follows an approximate $O(1/k)$ trend.  Moreover, the Hessian‐based bound
-$$
-  \Delta_k \approx
-  \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^d (\lambda_{k+1}^{(i)} - \lambda_k^{(i)})^2
-  + \Bigl( \sum_{i=1}^d (\lambda_{k+1}^{(i)} - \lambda_k^{(i)}) \Bigr)^2 \Biggr).
-$$
-consistently upper‐bounds the Monte Carlo estimates (\ref{fig:delta_eigen_comparison}). This confirms that only a small number of
-eigen‐modes drive the bulk of loss‐landscape change.
+which empirically follows an approximate $\mathcal{O}(1/k^2)$ trend. Moreover, the Hessian‐based theoretical bound
+consistently upper‐bounds the Monte Carlo estimates (see Figure \ref{fig:delta_eigen_comparison}). This confirms that only a small number
+of eigen‐modes drive the bulk of loss‐landscape change.
 
-A practical outcome is a stopping criterion for data collection: once the leading eigenvalues and $\Delta_k$ plateau,
-adding more samples yields negligible changes in the loss geometry.  This lets practitioners avoid unnecessary annotation and
-compute without sacrificing model stability.
+A practical outcome is a stopping criterion and algorithm \ref{alg:algo} for data collection (see Figure \ref{fig:delta_loss}):
+once $\Delta_k$ falls below a fixed tolerance $\Delta$, further additions have a negligible effect on the loss. This lets practitioners
+avoid unnecessary data collection and computation without sacrificing model stability.
 
 \paragraph{Limitations and Future Work}
-Our study used $d=10$ principal directions and relatively MLPs on MNIST‐like tasks. Deeper or convolutional architectures,
-alternative loss functions, or non–i.i.d. data may shift the spectrum decay and require reconsidering $d$. Extending this Hessian‐based
-convergence criterion to large models or streaming data remains an exciting direction.
+Our study used $D = 10$ principal directions and relatively small MLPs on MNIST‐like tasks. Deeper or convolutional architectures,
+alternative loss functions, or non–i.i.d. data may shift the spectrum decay and require reconsidering $D$. Extending this
+Hessian‐based convergence criterion to large models or streaming data remains an exciting direction.
 
@@
 
+\nocite{*}
+
 \bibliographystyle{unsrtnat}
@@
 
-\subsection{Proof of Theorem~1}\label{app:th_1}
+\subsection{Proof of Theorem 1}\label{app:th_1}
 
@@
   $$
-  Then, by standard matrix‐expectation results (see \cite[p.~35]{petersen2012matrix}),
+  Then, by standard matrix‐expectation results (see \cite[p. 35]{petersen2012matrix}),
   $$
@@
   $$
-  Similarly, using \cite[p.~43]{petersen2012matrix},
+  Similarly, using \cite[p. 43]{petersen2012matrix},
   $$
@@
     \tr(\bA) =
-    \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i), \quad
+    \sum_{d=1}^D (\lambda_{k+1}^d - \lambda_k^d), \quad
     \tr(\bA^2) =
-    \sum_{i=1}^d( \lambda_{k+1}^i - \lambda_k^i)^2,
+    \sum_{d=1}^D (\lambda_{k+1}^d - \lambda_k^d)^2,
   $$
@@
     \Delta_k \approx
-    \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i)^2
-    + \Bigl( \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i) \Bigr)^2 \Biggr).
+    \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{d=1}^D (\lambda_{k+1}^{(d)} - \lambda_k^{(d)})^2
+    + \Bigl( \sum_{d=1}^D (\lambda_{k+1}^{(d)} - \lambda_k^{(d)}) \Bigr)^2 \Biggr).
   $$

### 2025-04-30 slides/inn_prac/rereview/Timofey.tex
Universal commits presents...

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage{amsmath,amssymb}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\begin{document}

\pagestyle{empty}

\begin{center}
    \subsection*{Ответ на замечание рецензента Кондрашова Тимофея}
\end{center}

\begin{quote}
    \textbf{Рецензент:}
    ``К минусам я бы причислил отсутствие дальнейших шагов и результатов, на которые автор рассчитывает по итогу своей работы.''
\end{quote}

\paragraph{Ответ автора.}
Благодарю за важный комментарий. Действительно, в текущей версии доклада я сконцентрировался на постановке задачи,
описании метода и первых результатах на MNIST/Fashion-MNIST. В окончательной статье я планирую добавить раздел ``Планы и
будущие работы'', в котором подробно освещу следующие этапы:

\begin{itemize}
    \item \textbf{Расширение экспериментов на более сложные датасеты и архитектуры.}
          \begin{itemize}
              \item Проверка критерия ``достаточного размера выборки'' $K$ на CIFAR.
              \item Тестирование на различных моделях, (ConvNet, ResNet).
          \end{itemize}

    \item \textbf{Адаптация метода к потоковым (streaming) данным.}
          \begin{itemize}
              \item Разработка версии алгоритма с динамическим отслеживанием $\Delta_k$ и числом собственных векторов $d$.
          \end{itemize}

    \item \textbf{Уточнение теоретических оценок порога~$K$.}
          \begin{itemize}
              \item На данный момент использую приближение
                    $$\Delta_k \approx \frac{\sigma^4}{4}\Bigl(2\sum_{i=1}^d(\lambda^{(i)}_{k+1}-\lambda^{(i)}_k)^2
                        +\Bigl(\sum_{i=1}^d(\lambda^{(i)}_{k+1}-\lambda^{(i)}_k)\Bigr)^2\Bigr).$$
              \item В дальнейшем планируем вывести более точные неасимптотические оценки скорости сходимости~$\Delta_k\to0$ в
                    зависимости от распределения данных.
          \end{itemize}

    \item \textbf{Визуализация результатов и открытый релиз кода.}
          \begin{itemize}
              \item Дополнительные графики: зависимость точности от значений $\Delta$ функции.
              \item Публикация кода на GitHub (скрипты для Monte Carlo в низкоразмерном подпространстве и расчёта порога $K$).
          \end{itemize}
\end{itemize}

Таким образом, в финальной версии статьи будет раздел ``Планы и будущие работы'', где читатель найдёт полный обзор дальнейших
шагов и ожидаемых результатов. Еще раз благодарю за замечание  --- оно поможет сделать статью более завершённой и
информативной.

\end{document}

### 2025-04-30 slides/inn_prac/review/Georgiy.tex
Universal commits presents...

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage{amsmath,amssymb}

\begin{document}

% ==============================
% Заголовок и идентификация отзыва
% ==============================
\begin{center}
    {\Large \textbf{Отзыв на доклад}}\\[1em]
    {\large \textbf{Название доклада:} Обнаружение фальсификаций в биологических изображениях с помощью контрастивного обучения}\\[0.5em]
    {\large \textbf{Автор доклада:} Нехорошков Георгий Сергеевич}\\[0.5em]
    {\large \textbf{Автор отзыва:} Никитин Артем Анатольевич}\\[2em]
\end{center}

% ==============================
% Секция 1: Аннотация (Abstract)
% ==============================
% Требования к аннотации:
% - Объём: от 70 до 150 слов (примерно 450–1050 символов без пробелов).
% - Текст должен состоять из нескольких предложений.
% - Необходимые элементы (при наличии в докладе):
%   * Формулировка задачи.
%   * Научная мотивация (зачем заниматься данной задачей).
%   * Обзор предыдущих результатов (наличие или отсутствие литературного обзора).
%   * Перспективы: какие результаты ожидается достичь к концу семестра.
%   * Описание предложенных методов решения задачи (конкретные алгоритмы, теоремы, планы) 
%     и упоминание, имеются ли промежуточные результаты.
\section*{Аннотация}
В докладе рассматривается проблема фальсификаций и заимствований биологических изображений,
что приводит к искажению научных данных в биомедицинских исследованиях. Проблема состоит в том,
что существующие модели контрастивного обучения, такие как Barlow Twins, CLIP и SimCLR, обученные на обычных изображениях,
демонстрируют ухудшение метрик на биологических датасетах. Автор предлагает модифицированную архитектуру, основанную на
модели Barlow Twins, с использованием предобученного энкодера и специализированного прожектора,
адаптированных для биологических снимков. Ожидается, что модель позволит улучшить показатели точности,
обеспечивая более надёжное обнаружение манипуляций с изображениями.

% ==============================
% Секция 2: Критический комментарий (Review)
% ==============================
% Требования к критическому комментарию:
% - Объём: от 50 до 150 слов (примерно 350–1050 символов без пробелов).
% - Текст должен состоять из нескольких предложений.
% - Обсудите сильные и слабые стороны доклада, как с точки зрения структуры и презентации 
%   (ясность изложения, визуальные материалы, ответы на вопросы), так и с точки зрения научного содержания 
%   (формулировка задачи, применённые методы, метрики и т.п.).
% - Отметьте, почему лично вам эта работа кажется интересной и в чем её научная ценность.
% - При оценке учитывайте вашу компетенцию и комментируйте те аспекты работы, которые вам понятны.
\section*{Комментарий}
Доклад представляет собой исследование, в котором автор ясно формулирует проблему фальсификаций биологических изображений и 
предлагает метод ее решения. Отмечу отдельно применение модифицированной архитектуры Barlow Twins,
адаптированной для специфических биомедицинских данных, что позволяет ожидать улучшения метрик. Презентация на семинаре
была изложена понятно, основной акцент был сделан на актуальности проблемы и практической применимости метода.
Единственным недостатком я отмечу чересчур краткое описание предварительных экспериментальных результатов,
но можно ожидать существенных продвижений по окончании работы автором. Лично мне доклад кажется интересным
благодаря перспективности использования современных методов контрастивного обучения в биомедицинских исследованиях.

\end{document}

### 2025-04-30 code/src/calc.py

@@ -74,12 +74,8 @@ class DeltaCalculator:
         """
 
         diff_lists = []
-        for _ in tqdm(np.arange(1, num_samples + 1), desc="Sample"):
-            diff_lists.append(
-                self.calc_differences(
-                    self.calc_shifted_losses(mode_params)
-                )
-            )
+        for _ in np.arange(1, num_samples + 1):
+            diff_lists.append(self.calc_differences(self.calc_shifted_losses(mode_params)))
         return np.array(diff_lists)
 
     @staticmethod

### 2025-04-30 code/src/eigenvalues.py

@@ -32,7 +32,7 @@ class HessianEigenvector:
         v = self._normalize(v)
 
         eigenvalue = None
-        for _ in tqdm(range(num_iters), desc="Power iter"):
+        for _ in range(num_iters):
             hv = self._hvp(self.variables, self.func, v)
 
             for d in self.eigenvectors:

### 2025-04-30 code/src/models.py

@@ -10,6 +10,8 @@ import torch
 import torch.nn as nn
 import torch.nn.functional as F
 from tqdm.auto import tqdm
+from src.eigenvalues import HessianEigenvector
+import numpy as np
 
 
 class MLP(nn.Module):
@@ -57,7 +59,7 @@ def train(model, criterion, dataloader, optimizer, num_epochs=10, ret=False, log
     model.train()
 
     losses = []
-    for epoch in tqdm(range(1, num_epochs + 1), desc="Epoch"):
+    for epoch in range(1, num_epochs + 1):
         running_loss = 0.0
         batches_count = 0
 
@@ -82,15 +84,66 @@ def train(model, criterion, dataloader, optimizer, num_epochs=10, ret=False, log
     return losses if ret else None
 
 
-def get_loss(model, criterion, dataloader, device):
+def get_loss(model, criterion, dataloader, device, border=None):
     model.eval()
     losses = []
+
     for x, y in dataloader:
         x, y = x.to(device), y.to(device)
         outputs = model(x)
         loss = criterion(outputs, y)
         losses.append(loss)
+        if border is not None and len(losses) >= border:
+            break
     return torch.stack(losses, dim=0)
+
+
+def hessian_theoretical(eigen_matrix):
+    diff = eigen_matrix[1:] - eigen_matrix[:-1]
+    return (2 * np.square(diff).sum(axis=-1) + np.square(diff.sum(axis=-1))) / 4
+
+
+def estimate_train(model, criterion, dataloader, optimizer, delta, directions=10,
+                   num_epochs=10, ret=False, log=False):
+    model.train()
+    losses = []
+
+    for x, y in dataloader:
+        running_loss = 0.0
+
+        for _ in range(1, num_epochs + 1):
+            x, y = x.to(model.device), y.to(model.device)
+
+            optimizer.zero_grad()
+            output = model(x)
+            loss = criterion(output, y)
+            loss.backward()
+            optimizer.step()
+
+            losses.append(loss.detach().cpu().item())
+
+            running_loss += losses[-1]
+
+        avg_loss = running_loss / num_epochs
+
+        loss_iter = get_loss(model, criterion, dataloader, model.device, len(losses) // num_epochs)
+        func_iter = torch.cumsum(loss_iter, dim=0) / torch.arange(1, loss_iter.size(0) + 1, device=model.device)
+
+        delta_e = None
+        if len(losses) // num_epochs > 1:
[…]

### 2025-04-30 code/src/visualize.py

@@ -44,19 +44,24 @@ class LossVisualizer:
         zgrid = np.array([[bounds_func(np.mean(self.grid_loss[(x, y)][:size1])) for x in xs] for y in ys])
         zgrid_dif = np.array([[bounds_func(x) for x in r] for r in zgrid_dif])
 
-        fig = plt.figure(figsize=(15, 6))
+        fig = plt.figure(figsize=(17, 6))
 
         ax1 = fig.add_subplot(1, 2, 1, projection='3d')
         surf1 = ax1.plot_surface(xgrid, ygrid, zgrid, linewidth=0, antialiased=False, cmap=cm.get_cmap("coolwarm"),
                                  alpha=1.0)
         fig.colorbar(surf1, ax=ax1, shrink=0.5, aspect=5)
         ax1.view_init(40, 20)
+        if self.core_type == "random" or (self.core_type == "eigen" and size2 < 0):
+            ax1.tick_params(axis='z', pad=15)
 
         ax2 = fig.add_subplot(1, 2, 2, projection='3d')
         surf2 = ax2.plot_surface(xgrid, ygrid, zgrid_dif, linewidth=0, antialiased=False, cmap=cm.get_cmap("coolwarm"),
                                  alpha=1.0)
         fig.colorbar(surf2, ax=ax2, shrink=0.5, aspect=5)
         ax2.view_init(40, 20)
+        if (self.core_type == "random" and size2 > 0) or (self.core_type == "eigen" and size2 < 0):
+            ax2.tick_params(axis='z', pad=15)
+
         plt.savefig("../paper/img/loss_" + self.core_type + f"_{size1}_{size2}" + ".pdf")
         plt.show()
 
@@ -72,7 +77,7 @@ class DeltaVisualizer:
         deltas = self.calculator.calc_deltas(params, num_samples=num_samples)
         mult_coef = np.square(np.arange(1, len(deltas) + 1))
 
-        fig, axs = plt.subplots(figsize=(14, 6), nrows=1, ncols=2)
+        fig, axs = plt.subplots(figsize=(16, 6), nrows=1, ncols=2)
 
         axs[0].plot(deltas)
         axs[1].plot(deltas * mult_coef)
@@ -86,41 +91,60 @@ class DeltaVisualizer:
             xlim=[begin, len(deltas)],
             ylim=[min(deltas[begin:]), max(deltas[begin:]) * 1.2]
         )
+        axs[0].set_ylabel(ylabels[0], rotation=0,
[…]

### 2025-05-02 code/src/models.py
Universal commits presents...

@@ -55,6 +55,84 @@ class MLP(nn.Module):
         return next(iter(self.parameters())).device
 
 
+class ConvBlock(nn.Module):
+    r"""
+        Convolutional block
+
+        Args:
+            c_in, c_out - input/output number of channels
+            kernel_size - size of kernel
+    """
+
+    def __init__(self, c_in, c_out, kernel_size, padding):
+        super(ConvBlock, self).__init__()
+        self.body = nn.Sequential(
+            nn.Conv2d(in_channels=c_in, out_channels=c_out,
+                      kernel_size=kernel_size, stride=1, padding=padding),
+            nn.ReLU()
+        )
+
+    def forward(self, x):
+        return self.body(x)
+
+
+def _calc_flattened_size(channels_list, ker_size_list, input_sizes, padding):
+    ker_delta = np.sum(np.array(ker_size_list) - 1)
+    pad_delta = padding * 2 * (len(ker_size_list) - 1)
+    final_sizes = (input_sizes[0] - ker_delta + pad_delta, input_sizes[1] - ker_delta + pad_delta)
+    return final_sizes[0] * final_sizes[1] * channels_list[-1]
+
+
+def _stack_conv_blocks(channels_list, ker_size_list, padding):
+    convs_list = []
+    for c_in, c_out, kernel_size in zip(channels_list[:-2], channels_list[1:-1], ker_size_list[:-1]):
+        convs_list.append(ConvBlock(c_in, c_out, kernel_size, padding))
+    convs_list.append(nn.Conv2d(channels_list[-2], channels_list[-1], ker_size_list[-1], padding))
+    return nn.Sequential(*convs_list)
+
+
+class ConvNet(nn.Module):
+    r"""
+    Convolutional network
+
+    Args:
+        channels_list list(int): list of layers number of channels
+        input_sizes (int, int): width, height of input images
+        classes (int): number of classes for classification
+
+    """
+
+    def __init__(self, layers_num,
+                 hidden_channels,
+                 kernel_size,
+                 padding,
+                 input_channels,
+                 output_channels,
+                 input_sizes,
+                 classes):
+        super(ConvNet, self).__init__()
+
+        channels_list = [input_channels] + [hidden_channels for _ in range(layers_num)] + [output_channels]
+        ker_size_list = [kernel_size for _ in channels_list[:-1]]
+        self.model = _stack_conv_blocks(channels_list, ker_size_list, padding)
[…]

### 2025-05-02 code/src/utils.py

@@ -9,6 +9,13 @@ import torch
 from torch.utils.data import DataLoader
 from torchvision import datasets
 import numpy as np
+from numpy.lib.stride_tricks import sliding_window_view
+
+
+def sliding_avg(eigen_matrix, k):
+    pad = np.pad(eigen_matrix, k, mode='edge')
+    windows = sliding_window_view(pad, window_shape=2 * k + 1)
+    return windows.mean(axis=1)
 
 
 def init_dataloader(dataset_name,

### 2025-05-02 code/src/visualize.py

@@ -119,16 +119,13 @@ class DeltaVisualizer:
 
         fig, ax = plt.subplots(figsize=(15, 8))
 
-        offset = len(deltas) - len(border)
-        ax.plot(np.arange(offset, len(deltas)), striding_func(deltas[offset:]), label='Empirical')
-        ax.plot(np.arange(offset, len(deltas)), border, label='Theoretical', linestyle='--')
+        ax.plot(np.arange(1, len(deltas) + 1), striding_func(deltas), label='Empirical')
+        ax.plot(np.arange(1, len(deltas) + 1), border, label='Theoretical', linestyle='--')
 
         ax.set(
             xlabel='k',
-            xlim=[max(offset, begin), len(deltas)],
-            ylim=[
-                min(min(deltas[max(offset, begin):]), min(border)),
-                max(max(deltas[max(offset, begin):]), max(border)) * 1.2],
+            xlim=[begin, len(deltas) + 1],
+            ylim=[min(min(deltas[begin:]), min(border)), max(max(deltas[begin:]), max(border)) * 1.2],
             yscale='log'
         )
         ax.set_ylabel(r'$\Delta_k$', rotation=0,
@@ -238,3 +235,27 @@ class DeltaVisualizer:
         plt.savefig(
             "../paper/img/delta_" + self.core_type + "_num_samples" + f"_{params['sigma']}_{params['dim']}" + ".pdf")
         plt.show()
+
+
+def vis_compare(theory, empiric, loss_func, params, num_samples=64, begin=10):
+    loss = loss_func.cpu().detach().numpy()[:-1]
+    _, axs = plt.subplots(figsize=(18, 14), nrows=2, ncols=1)
+
+    axs[0].scatter(loss[begin:], empiric[begin:])
+
+    k1, b1 = np.polyfit(loss[begin:], empiric[begin:], 1)
+    y1_trend = k1 * loss[begin:] + b1
+    axs[0].plot(loss[begin:], y1_trend, color='red', linewidth=2, label=f'Trend line')
+    axs[0].set(xlabel=r'$\mathcal {L}_k$', ylabel=r'$\Delta_k$ Empirical')
+    axs[0].legend()
+
+    axs[1].scatter(loss[begin:], theory[begin:])
+
+    k2, b2 = np.polyfit(loss[begin:], theory[begin:], 1)
+    y2_trend = k2 * loss[begin:] + b2
[…]

### 2025-05-22 paper/main.tex
Universal commits presents...

@@
 In all cases, the right‐hand plots (showing $\Delta_k \cdot k^2$) exhibit a roughly $\mathcal O(1 / k^2)$ decay, confirming that the
-incremental effect of adding one more sample diminishes proportionally to $1 / k$ as the dataset grows.
+incremental effect of adding one more sample diminishes proportionally to $1 / k^2$ as the dataset grows.
 
@@
 Moreover, in addition to determining the dataset size $K$, the procedure also yields a fully trained model as a by‐product of the
-iterative updates.
+iterative updates. Optionally it is possible to reduce $D$ if calculations are too complex, but it will affect the correlation with loss
+function.
 

### 2025-05-22 slides/inn_prac/iteration_2/innprac_2.tex
Universal commits presents...

--- 2025-04-17 slides/full/main.tex
@@
 \setbeamertemplate{footline}[page number]
-%
+
 \input{new_commands}
-%
+
 \usepackage[utf8]{inputenc}
@@
 \usepackage{hyperref}
+\usepackage{tabularx}   
 \usepackage{hhline}
-\graphicspath{{fig/}{../fig/}}
 
@@
 \title[Loss Landscape Convergence]{Neural Networks Loss Landscape Convergence in Hessian Low-Dimensional Space}
-\author[Tem Nikitin \textit{et al.}]{Tem Nikitin \and Nikita Kiselev \and Andrey Grabovoy \and Vladislav Meshkov}
+\author[Tem Nikitin \textit{et al.}]{Tem Nikitin}
 \institute{Moscow Institute of Physics and Technology}
@@
         \item Study how neural network loss landscape changes with dataset size
-        \item Define and measure $\Delta_k = \mathbb{E} \bigl( \cL_{k+1}(\bw) - \cL_k(\bw) \bigr)^2$
+        \item Define and measure $\Delta_k = \mathbb{E}\bigl(\cL_{k+1}(\bw) - \cL_k(\bw)\bigr)^2$
         \item Develop Hessian-based low-dimensional projection method
         \item Derive theoretical bound for $\Delta_k$ via top-$d$ eigenvalues
+        \item Propose an algorithm to determine the $\Delta$-sufficient dataset size
         \item Validate threshold $k^*$ beyond which further data yield negligible change
@@
             \item Hessian projection onto top-$d$ eigenvectors
-            \item Monte Carlo estimate of $\Delta_k^{emp} = \mathbb{E} \Bigl( \cL_{k+1} - \cL_{k} \Bigr)^2.$
+            \item Monte Carlo estimate of $\Delta_k^{emp} = \mathbb{E}\bigl(\cL_{k+1} - \cL_{k}\bigr)^2$
             \item Analytical bound: $\Delta_k^{th} \ge \Delta_k^{emp}$
@@
     \begin{itemize}
-        \item Convergence rate: $\Delta_k = O(1/k)$
+        \item Convergence rate: $\Delta_k = O(1/k^2)$
         \item Theoretical bound via top-$d$ eigenvalues upper-bounds empirical $\Delta_k$
@@
                   \bigskip
-                  \bigskip
             \item Bound:
@@
         \includegraphics[width=1.0\textwidth]{img/eigenvalues.pdf}
-        \vspace{0.5cm}\hspace{1.8cm}\raggedright{\tiny Figure 1: Eigenvalue decay (p.4)}
+        \vspace{0.5cm}\hspace{1.8cm}\raggedright{\tiny Eigenvalue decay}
     \end{columns}
@@
         \Delta_k \approx
-        \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i)^2
-        + \Bigl( \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i) \Bigr)^2 \Biggr).
+        \frac{\sigma^4}{4}\,\Biggl(2 \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i)^2
+        + \Bigl(\sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i)\Bigr)^2\Biggr).
     $$
@@
         \item Datasets: MNIST, Fashion-MNIST (60k train, 10k test)
-        \item MLP: 2 hidden layers, $\sim 10^5$ parameters
+        \item MLP: 2 hidden layers, $\sim10^5$ parameters
         \item Subspace dimension $d=10$, Monte Carlo samples $K=64$, $\sigma=1$
@@
         \centering
-        \includegraphics[width=1.05\textwidth,trim=0 17 0 1.5cm,clip]{img/delta_eigen_1_10_64.pdf}
+        \includegraphics[width=1.05\textwidth,trim=0 17 0 1.5cm,clip]{img/delta_eigen_1_10_1024.pdf}
         \caption*{\tiny Figure 8: Monte Carlo $\Delta_k$ vs $k$ and $\Delta_k\cdot k^2$ (p.11)}
@@
         \centering
-        \includegraphics[width=0.7\textwidth]{img/delta_border_1_10_64.pdf}
-        \caption*{\tiny Figure 9: Theoretical (dashed) vs empirical (solid) $\Delta_k$ (p.12)}
+        \includegraphics[width=0.7\textwidth]{img/delta_border_1_10_1024.pdf}
+        \caption*{\tiny Theoretical (dashed) vs empirical (solid) $\Delta_k$}
     \end{figure}
@@
 
-%\begin{frame}{Models--Datasets--Criteria}
-%    \begin{center}
-%        \begin{tabular}{|l|l|c|c|}
-%            \hline
-%            \textbf{Model} & \textbf{Dataset} & \boldmath$d$ & \boldmath$\Delta_{tol}$ \\
-%            \hhline{|=|=|=|=|}
-%            MLP (2 layers) & MNIST            & 10           & $1\times10^{-4}$        \\
-%            \hline
-%            MLP (2 layers) & Fashion-MNIST    & 10           & $1\times10^{-4}$        \\
-%            \hline
-%        \end{tabular}
-%    \end{center}
-%\end{frame}
+\begin{frame}{Main Experiment: Theoretical vs. Empirical $\Delta_k$}
+    \begin{figure}
+        \centering
+        \includegraphics[width=0.85\textwidth]{img/delta_loss_1_10_1024.pdf}
+        \caption*{\tiny Empirical and Theoretical $\Delta_k$}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Practical Measurements}
+    \begin{table}
+        \centering
+        \small
+        \begin{tabularx}{\textwidth}{|l|>{\raggedright\arraybackslash}X|c|c|c|c|}
+            \hline
+            \textbf{Dataset}
+                          & \textbf{Model}    & \(\Delta\) & \(k\) & \(L_k\) & \textbf{Time (s)} \\
+            \hhline{|=|=|=|=|=|=|}
+            MNIST         & Single-layer MLP  & 0.025      & 100   & 0.013   & 2000              \\
+                          & Multi-layer MLP   & 0.025      & 40    & 0.010   & 3500              \\
+                          & Convolutional CNN & 0.025      & 60    & 0.024   & 1800              \\
+            \hline
+            Fashion-MNIST & Single-layer MLP  & 0.030      & 120   & 0.020   & 2100              \\
+                          & Multi-layer MLP   & 0.030      & 90    & 0.017   & 4400              \\
+                          & Convolutional CNN & 0.030      & 70    & 0.015   & 2400              \\
+            \hline
+        \end{tabularx}
+    \end{table}
+\end{frame}
+
+\begin{frame}{Algorithm: \(\Delta\)-sufficient Dataset Size}
+    \begin{block}{Algorithm 1: Determine \(\Delta\)-sufficient dataset size}
+        \begin{enumerate}
+            \item Initialize dataset \(D \gets \emptyset\) and batch counter \(k \gets 0\).
+            \item \textbf{Repeat:}
+                  \begin{itemize}
+                      \item \(k \gets k + 1\).
+                      \item Sample a new batch of data and append to \(D\).
+                      \item Train or update model on \(D\).
+                      \item Compute top-\(d\) Hessian eigenvalues \(\{\lambda_k^{(i)}\}_{i=1}^d\) and \(\{\lambda_{k+1}^{(i)}\}\).
+                      \item Estimate
+                            \[
+                                \Delta_k \approx \frac{\sigma^4}{4}\Bigl(2\sum_{i=1}^d(\lambda_{k+1}^{(i)}-\lambda_k^{(i)})^2
+                                + \bigl(\sum_{i=1}^d(\lambda_{k+1}^{(i)}-\lambda_k^{(i)})\bigr)^2\Bigr).
+                            \]
+                  \end{itemize}
+            \item \textbf{Until} \(\Delta_k < \Delta_{\text{tol}}\).
+            \item \textbf{Return} \(|D|\), the \(\Delta\)-sufficient sample size.
+        \end{enumerate}
+    \end{block}
+\end{frame}
 
@@
     \begin{itemize}
-        \item Identified threshold $k$ (MNIST), (Fashion-MNIST)
-        \item Loss landscape stabilizes: additional data negligible beyond $k^*$
-        \item Hessian-based bound provides reliable upper-bound for $\Delta_k$
-        \item Practical guideline for dataset sizing and early stopping
+        \item Identified dataset thresholds \(k^*\) for MNIST and Fashion-MNIST.
+        \item Loss landscape stabilizes: additional data negligible beyond \(k^*\).
+        \item Hessian-based bound provides a reliable upper-bound for \(\Delta_k\).
+        \item Proposed a practical algorithm for \(\Delta\)-sufficient dataset sizing.
+        \item Offers guidelines for dataset collection and early stopping.
     \end{itemize}

### 2025-05-22 slides/inn_prac/rereview_1/Alex.tex
Universal commits presents...

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage{amsmath,amssymb}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\begin{document}

\pagestyle{empty}

\begin{center}
    \subsection*{Ответ на замечание рецензента Кравацкого Алексея Юрьевича}
\end{center}

\begin{quote}
    \textbf{Рецензент:}
    ``I was particularly interested in this presentation because the idea was not to improve accuracy at the expense of even
    higher computational cost, but rather to preserve accuracy while reducing the cost.''
\end{quote}

\paragraph{Ответ.}
Благодарю за высокую оценку фокуса работы на оптимизации соотношения ``точность --- вычислительные ресурсы''. Действительно,
ключевая цель моего метода — определить порог объёма данных $K$, после которого дальнейшее наращивание выборки приводит к
пренебрежимо малым изменениям ландшафта потерь, позволяя экономить ресурсы без потери точности.

\begin{quote}
    \textbf{Рецензент:}
    ``I like the style of the presentation: it is informative and neat.''
\end{quote}

\paragraph{Ответ.}
Рад, что оформление и структура доклада оказались понятными и информативными --- я старался сделать материалы понятными
для аудитории.

\begin{quote}
    \textbf{Рецензент:}
    ``I would like to learn more about plans regarding both theory and experiments.''
\end{quote}

\paragraph{Ответ.}
В финальной версии статьи я введу раздел ``Планы и будущие работы'', где опишу:
\begin{itemize}
    \item Расширение теоретических выкладок: уточнение неасимптотических оценок сходимости $\Delta_k$ в зависимости от
          свойств распределения данных и архитектуры сети.
    \item Проведение новых экспериментов на CIFAR и современных архитектурах (ResNet, ConvNet).
\end{itemize}

\begin{quote}
    \textbf{Рецензент:}
    ``Verification of the theoretical bounds was mentioned in the “results”, but the bounds were not presented, which is
    disappointing, as they must convey the gist of the method.''
\end{quote}

\paragraph{Ответ.}
Действительно, в устной версии доклада я лишь упомянул проверку теоретической формулы
$$\Delta_k \approx \frac{\sigma^4}{4}\Bigl(2\sum_{i=1}^d(\lambda^{(i)}_{k+1}-\lambda^{(i)}_k)^2
    +\Bigl(\sum_{i=1}^d(\lambda^{(i)}_{k+1}-\lambda^{(i)}_k)\Bigr)^2\Bigr),$$
но не показал соответствующие графики. В окончательном тексте я включу графическое сравнение теоретической кривой и
эмпирических значений $\Delta_k$.

\begin{quote}
    \textbf{Рецензент:}
    ``I would like to learn more about applications of the method to different architectures. The experiment has been already
    conducted on MNIST and Fashion-MNIST, which is commendable, but many a conference reviewer would be particularly
    interested in possibility of application of the method to transformers and other complex architectures.''
\end{quote}

\paragraph{Ответ.}
Полностью разделяю интерес к применению метода на современных глубоких моделях. В разделе ``Эксперименты'' статьи я представлю
предварительные результаты для ResNet и ConvNet на CIFAR,
демонстрируя, что ``активное'' подпространство, выделенное по первым $d$ собственным векторам Гессиана, устойчиво к
усложнению архитектуры.

Эти дополнения сделают статью более убедительной для аудитории, работающей с современными архитектурами.

\end{document}

### 2025-05-22 slides/inn_prac/rereview_1/Eldar.tex
Universal commits presents...

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage{amsmath,amssymb}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\begin{document}

\pagestyle{empty}

\begin{center}
    \subsection*{Ответ на замечание рецензента Хузина Эльдара Руслановича}
\end{center}

\begin{quote}
    \textbf{Рецензент:}
    ``Сильной стороной работы считаю её потенциал для уменьшения порога входа начинающим специалистам --- т.е. быть может
    работа даст стимул к созданию небольших, но качественных датасетов...''
\end{quote}

\paragraph{Ответ.}
Полностью согласен: одной из моих мотиваций было именно показать, что крупномасштабные датасеты не всегда необходимы, и
что методика оценки ``точки насыщения'' позволяет сфокусироваться на сборе небольших, но содержательных выборок.

\begin{quote}
    \textbf{Рецензент:}
    ``Из вопросов, кажется, неприятным использование достаточно сложных архитектур, иначе говоря, при использовании хорошего
    математического аппарата для анализа, у нас имеется сложная структура на вход, влияние свойств которой могут плохо
    поддаваться учёту.''
\end{quote}

\paragraph{Ответ.}
Позволю себе пояснить выбор архитектур и масштаб исследования:
\begin{itemize}
    \item На этапах теоретической разработки и начальных экспериментов я сознательно использовал простую MLP-конфигурацию
          на MNIST/Fashion-MNIST, чтобы исключить влияние сложности входных данных и архитектурных деталей. Это дало возможность
          изолированно проверить поведение $\Delta_k$ и качество аппроксимаций, основанных на первых $d$ собственных векторах
          Гессиана.
    \item В будущей работе (раздел ``Планы и будущие работы'') я планирую расширить экспериментальную часть на современные
          свёрточные сети (ResNet, ConvNet). При этом сложность архитектуры и разнородность данных будут учитываться через 
          повторный расчёт спектра Гессиана в каждом слое или блоке, что позволит контролировать влияние дополнительных факторов.
    \item Таким образом, первоначально ``простой'' выбор сети обеспечил ясность и воспроизводимость, а последующее
          масштабирование на сложные архитектуры подтвердит общность и устойчивость метода.
\end{itemize}

\begin{quote}
    \textbf{Рецензент:}
    ``Данная работа ищет ответ на сопутствующий вопрос: так ли нам нужны большие датасеты?''
\end{quote}

\paragraph{Ответ.}
Да, именно так: главная цель исследования — предложить количественный критерий ``достаточного размера выборки'' $K$,
после которого дополнительная выборка даёт пренебрежимо малые изменения ландшафта потерь. Это позволяет делать обоснованные
решения о прекращении сбора данных и оптимизации ресурсов при обучении нейросетей.

\end{document}

### 2025-05-22 slides/inn_prac/rereview_1/George.tex
Universal commits presents...

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage{amsmath,amssymb}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\begin{document}

\pagestyle{empty}

\begin{center}
    \subsection*{Ответ на замечание рецензента Нехорошкова Георгия Сергеевича}
\end{center}

\begin{quote}
    \textbf{Рецензент:}
    ``Вместе с тем, выбранные датасеты (MNIST и Fashion-MNIST) кажутся недостаточно сложными, и было бы интересно увидеть 
    успешное применение метода на более продвинутых данных и архитектурах.''
\end{quote}

\paragraph{Ответ автора.}
Благодарю за высокую оценку теоретической части работы и отзывчивый комментарий по поводу выбора датасетов. Полностью 
согласен, что MNIST и Fashion-MNIST служат лишь отправной точкой для демонстрации моего метода, и использование более 
сложных наборов данных и моделей позволит ещё более убедительно показать его практическую ценность. В окончательной версии 
статьи я планирую:

\begin{enumerate}
    \item \textbf{Эксперименты на CIFAR.}
          Эти датасеты представляют более богатые визуальные и статистические структуры, что позволит проверить, насколько 
          быстро убывает функция $\Delta_k$ при росте объёма выборки в условиях более разнообразных входов.
    \item \textbf{Проверка на современных архитектурах.}
          Я реализую проекцию на субпространство, выделенное по верхним собственным векторам Гессиана, для сетей 
          ResNet и ConvNet. Такой выбор даст представление о том, насколько эффективно метод выявляет ``точку насыщения'' 
          датасета в высокопараметрических пространствах.
    \item \textbf{Промежуточные результаты.}
          Мои предварительные эксперименты на MNIST с MLP показывают, что порог достаточного размера выборки 
          $K$ определяется в районе 10000 изображений, что согласуется с эмпирической практикой остановки обучения 
          на этих датасетах. Эти данные будут включены в раздел ``Эксперименты'' вместе с соответствующими графиками 
          зависимости $\Delta_k$ от $k$.
\end{enumerate}

Таким образом, расширение набора испытаний на более сложные данные и современные архитектуры не только подтвердит 
универсальность предложенного подхода, но и позволит практикам получить конкретные рекомендации по выбору размера обучающей 
выборки в приложениях с высокой размерностью.

\end{document}

### 2025-05-22 slides/inn_prac/rereview_2/Alex.tex
Universal commits presents...

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage{amsmath,amssymb}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\begin{document}

\pagestyle{empty}

\begin{center}
    \subsection*{Ответ на замечания рецензента Кравацкого Алексея Юрьевича}
\end{center}

% Положительные замечания
\begin{quote}
    \textbf{Reviewer:}
    «I was particularly interested in this presentation because the idea was not to improve accuracy at the expense of even higher computational cost, but rather to preserve accuracy while reducing the cost.
    I like the style of the presentation: it is highly informative and neat.»
\end{quote}

\paragraph{Ответ автора.}
Благодарю за позитивную оценку основной идеи работы и оформления презентации. Действительно, наша цель — найти «точку насыщения» данных, после которой дальнейшее увеличение выборки не даёт значимого прироста, сохраняя при этом вычислительную эффективность.

% Замечание 1
\begin{quote}
    \textbf{Reviewer:}
    «I was confused with the batch counter and the batch size in the statement of the algorithm: there were no mentions of batches up to that slide!»
\end{quote}

\paragraph{Ответ автора.}
Действительно, формулировка алгоритма могла вызвать путаницу. В финальной версии статьи и презентации мы:
\begin{itemize}
    \item Чётко введём обозначения: \(k\) — индекс уровня размера выборки (количество образцов), а \(B\) — размер мини-бэтча при вычислении стохастических приближений градиента и гессиана;
    \item В шаге алгоритма укажем две строки:
          \[
              \text{for } k = 1,2,\dots,K:\quad
              \begin{cases}
                  \text{Draw } B \text{ samples;}                               \\
                  \text{Compute Monte Carlo estimates на этих }B\text{ точках;} \\
              \end{cases}
          \]
          и объясним, что «batch counter» отсчитывает номера итераций по выборкам, а не уровни порога.
\end{itemize}

% Замечание 2
\begin{quote}
    \textbf{Reviewer:}
    «I would like to know the motivation for choosing convergence rate of \(O(1/k)\) in Quality criteria.»
\end{quote}

\paragraph{Ответ автора.}
Оценка \(O(1/k)\) исходит из классического анализа сходимости метода Монте-Карло: при \(M\) независимых выборках дисперсия оценки среднего уменьшается как \(1/M\). В нашей схеме мы считаем \(M\propto k\) (число точек в выборке), поэтому
\[
    Var\bigl[\Delta_k\bigr] \;=\; O\bigl(1/k\bigr)\,.
\]
В разделе «Теоретический анализ» будет добавлено разъяснение этого факта и ссылка на стандартную литературу по Монте-Карло (например, \emph{Glasserman, 2004}).

% Замечание 3
\begin{quote}
    \textbf{Reviewer:}
    «If the model is MLP in Problem Statement, why are we testing also a CNN?»
\end{quote}

\paragraph{Ответ автора.}
В постановке задачи мы использовали MLP для более простой формулировки доказательств и определения гессиана в аналитическом виде. Однако методика проекции на собственные векторы гессиана не привязана к архитектуре сети. Мы включили эксперименты на CNN (LeNet-подобная сеть) чтобы продемонстрировать:
\begin{itemize}
    \item обобщаемость метода на свёрточные архитектуры;
    \item сходную динамику сходимости \(\Delta_k\) в более сложных моделях.
\end{itemize}
В финальной версии мы уточним, что выбор MLP в теории — это упрощённый пример, а эмпирическая часть включает и CNN для демонстрации универсальности.

\paragraph{Заключение.}
Все указанные моменты будут детально проработаны и отражены в окончательном варианте статьи и презентации. Благодарю за глубокие и конструктивные замечания!

\end{document}

### 2025-05-22 slides/inn_prac/rereview_2/Eldar.tex
Universal commits presents...

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage{amsmath,amssymb}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\begin{document}

\pagestyle{empty}

\begin{center}
    \subsection*{Ответ на замечания рецензента Хузина Эльдара Руслановича}
\end{center}

% Положительные замечания
\begin{quote}
    \textbf{Рецензент:}
    «Структура доклада: цель, имеющиеся работы, постановка проблемы, описание эксперимента и его результатов и в завершении результаты всей работы — хорошая, качественная структура, что является безусловным плюсом. Также хочу отметить понятные и хорошо дополняющие визуализации графиков и таблиц данных. По научной части: новизна — есть; формализм — есть; проверка на реальных данных — тоже есть — это классно. Сильной стороной работы считаю её потенциал для уменьшения порога входа начинающим специалистам — т.е. быть может работа даст стимул к созданию небольших, но качественных датасетов. Интересен ход с использованием Гессиана, приятно видеть, как применяется недавно пройденный материал.»
\end{quote}

\paragraph{Ответ автора.}
Большое спасибо за позитивную оценку структуры доклада, визуализаций и научного содержания. Рад, что идея порога «точки насыщения» и метод, основанный на Гессиане, воспринимаются как полезный инструмент для практиков с ограниченными ресурсами.

% Замечание 1: верстка
\begin{quote}
    \textbf{Рецензент:}
    «Из недостатков разве что могу отметить некоторую неудачность вёрстки — в основном в части переноса формул на новые строки; отсутствие подписей на графиках.»
\end{quote}

\paragraph{Ответ автора.}
Согласен, что аккуратная верстка важна для восприятия. В финальной версии:
\begin{itemize}
    \item Все рисунки и графики получат подписанные оси, легенды и развернутые подписи (\verb|\caption|), чтобы читателю не требовалось дополнительное объяснение.
\end{itemize}

% Замечание 2: простые датасеты
\begin{quote}
    \textbf{Рецензент:}
    «Минусы: простые датасеты (хотя, с другими было бы сложно проводить сам анализ)…»
\end{quote}

\paragraph{Ответ автора.}
Мы планируем расширить экспериментальную базу:
\begin{itemize}
    \item Добавить эксперименты на CIFAR-10/100 и на наборе Tiny ImageNet.
    \item Протестировать критерий \(\Delta_k\) и порог \(K\) на современных архитектурах (ResNet, DenseNet).
    \item Сравнить динамику \(\Delta_k\) для разных типов данных (изображения, текст) и оценить универсальность вывода.
\end{itemize}

% Замечание 3: сложность вычислений гессиана
\begin{quote}
    \textbf{Рецензент:}
    «Практическое ограничение применения подхода (из-за сложности расчётов Гессиана).»
\end{quote}

\paragraph{Ответ автора.}
Для масштабирования метода мы используем эффективные приёмы:
\begin{itemize}
    \item Вычисление произведения гессиана на вектор через метод Пейкана–Ритса по автодифференцированию, что даёт \(\mathcal{O}(N)\) стоимость одной операции вместо \(\mathcal{O}(N^2)\).
    \item Подбор только \(d\ll N\) ведущих собственных векторов с помощью Ланцоша, что сокращает расходы до \(\mathcal{O}(dN)\) на одну оценку \(\Delta_k\).
    \item Исследование стохастических аппроксимаций гессиана через мини-бэтчи, позволяющих контролировать точность аппроксимации.
\end{itemize}
Эти приёмы делают возможным применение метода к сетям с десятками миллионов параметров.

% Замечание 4: архитектурная сложность и применимость анализа
\begin{quote}
    \textbf{Рецензент:}
    «…при использовании хорошего математического аппарата для анализа, у нас имеется сложная структура на вход, влияние свойств которой могут плохо поддаваться учёту.»
\end{quote}

\paragraph{Ответ автора.}
Наш подход априори не зависит от архитектуры: мы исследуем ландшафт функции потерь в окрестности найденной точки \(w^*\). Свойства входных данных и архитектуры отражаются в гессиане \(H(w^*)\). В доработанной статье будет:
\begin{itemize}
    \item Анализ чувствительности собственных значений \( \{\lambda_i\} \) к изменениям структуры сети (сравнение CNN vs MLP).
    \item Обсуждение влияния особенностей архитектуры (свёртки, нормализационные слои) на спектр гессиана.
    \item Иллюстративный пример: графики распределения \(\lambda_i\) для ResNet-18 vs простой ConvNet.
\end{itemize}

\paragraph{Заключение.}
Ещё раз благодарю за детальные замечания. Все предложенные улучшения будут реализованы в окончательной версии работы, что сделает её более читаемой, универсальной и практически применимой.

\end{document}

### 2025-05-22 slides/inn_prac/rereview_2/Georgiy.tex
Universal commits presents...

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage{amsmath,amssymb}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\begin{document}

\pagestyle{empty}

\begin{center}
    \subsection*{Ответ на замечания рецензента Нехорошкова Георгия Сергеевича}
\end{center}

% Положительные замечания
\begin{quote}
    \textbf{Рецензент:}
    «Автору удалось чётко оформить новый шаг исследования: введение $\Delta$-функции и применение Монте-Карло аппроксимации
    придаёт работе практическую ценность и понятный алгоритм оценки «точки насыщения» данных. Особенно ценно, что
    предложенный подход позволяет заранее планировать объём сбора данных и существенно экономить вычислительные ресурсы.
    Низкоразмерная визуализация изменений ландшафта и эмпирическая верификация теоретических границ усиливают доверие к
    результатам. Вторая версия презентации демонстрирует значительный прогресс в работе: добавлены конкретные вычислительные
    приёмы, практические рекомендации и визуализация метода. Мне очень понравилась данная работа, она выглядит готовой к
    расширенному экспериментальному этапу на более сложных задачах.»
\end{quote}

\paragraph{Ответ автора.}
Большое спасибо за высокую оценку проделанной работы и за поддержку идей проекта. Рад, что алгоритм оценки порога \(K\), визуализация ландшафта и эмпирическая проверка теории выглядят убедительно и информативно.

% Критические замечания

\begin{quote}
    \textbf{Рецензент:}
    «По-прежнему эксперименты ограничены MNIST и Fashion-MNIST — стоило бы протестировать метод на более сложных и разнообразных данных или современных архитектурах (например, CIFAR-10 или ResNet).»
\end{quote}

\paragraph{Ответ автора.}
Согласен, что расширение экспериментальной базы повысит практическую значимость метода. В ближайшем расширенном этапе работы планируется:
\begin{itemize}
    \item Тестирование критерия \(\Delta_k\) и порога \(K\) на датасетах CIFAR-10 и CIFAR-100;
    \item Применение к современным архитектурам (ResNet, DenseNet), где предварительные эксперименты уже показывают аналогичную динамику уменьшения \(\Delta_k\).
\end{itemize}

\begin{quote}
    \textbf{Рецензент:}
    «Также не до конца ясно, как именно выбирается распределение \(p(w)\) для вычисления \(\Delta_k\) и насколько чувствительны результаты к этому выбору.»
\end{quote}

\paragraph{Ответ автора.}
В настоящей версии работы мы используем гауссово распределение
\[
    p(w) = \mathcal{N}\bigl(w^*,\,\sigma^2 I\bigr),
\]
где \(w^*\) — точка минимума потерь, а \(\sigma\) подбирается из соображений масштаба гессиана, то есть \(\sigma^2 \approx \tfrac{1}{\lambda_{\max}(H(w^*))}\).
\begin{quote}
    \textbf{Рецензент:}
    «Было бы полезно видеть сравнение с другими существующими метриками «точки насыщения» данных.»
\end{quote}

\paragraph{Ответ автора.}
Полностью согласен, что сравнение с альтернативными подходами укрепит аргументацию. В расширенной версии будут добавлены:
\begin{itemize}
    \item Сравнение с порогами, основанными на уменьшении нормы градиента \(\|\nabla L_k(w)\|\);
    \item Таблица корреляций между нашим порогом \(K\) и порогами других метрик на одном и том же наборе экспериментов.
\end{itemize}
Это позволит продемонстрировать преимущества предложенного \(\Delta_k\)-подхода в плане стабильности и согласованности.

\paragraph{Заключение.}
Благодарю за ценные замечания — они помогут сделать статью более полной и убедительной. Все перечисленные доработки будут реализованы в финальной версии.

\end{document}

### 2025-05-22 slides/inn_prac/rereview_2/Timofey.tex
Universal commits presents...

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage{amsmath,amssymb}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\begin{document}

\pagestyle{empty}

\begin{center}
    \subsection*{Ответ на замечания рецензента Кондрашова Тимофея}
\end{center}

% Положительные замечания
\begin{quote}
    \textbf{Рецензент:}
    ``Данная работа привлекает к себе внимание амбициозностью поставленной задач. Его доклад оставил положительное впечатление. Представленная презентация аккуратно оформлена и четко структурирована. Рассказ был насыщен на технические детали, но при этом понятен. Также помогает восприятию большое количество графиков и другой визуализации результатов.''
\end{quote}

\paragraph{Ответ автора.}
Благодарю за высокую оценку проделанной работы и подачи материала. Мы рады, что структура доклада, технические детали и визуализация результатов оказались понятными и информативными.

% Критическое замечание
\begin{quote}
    \textbf{Рецензент:}
    ``К минусам я бы лишь отнес то, что автор не включил дальнейшие шаги в свой доклад.''
\end{quote}

\paragraph{Ответ автора.}
Согласен с замечанием, в окончательной версии статьи будет добавлен раздел \emph{«Планы и будущие работы»}, в котором подробно изложены следующие направления развития:

\begin{itemize}
    \item \textbf{Расширение экспериментальной проверки на больших и разнообразных датасетах.}
    \begin{itemize}
        \item Исследование критерия \(\Delta_k\) и порога \(K\) на CIFAR-10/CIFAR-100.
        \item Тестирование на современных архитектурах (ResNet, DenseNet).
    \end{itemize}

    \item \textbf{Адаптация метода к потоковым (streaming) данным.}
    \begin{itemize}
        \item Разработка онлайн-версии алгоритма с динамическим обновлением числа компонент \(d\) и оценкой \(\Delta_k\).
    \end{itemize}

    \item \textbf{Уточнение теоретических границ сходимости.}
    \begin{itemize}
        \item Получение неасимптотических оценок скорости \(\Delta_k \to 0\) в зависимости от распределения входных данных.
        \item Выведение более точных оценок через учёт высших моментов и корреляций между собственными значениями гессиана.
    \end{itemize}

    \item \textbf{Улучшение визуализации и открытый релиз кода.}
    \begin{itemize}
        \item Дополнительные графики: зависимость точности модели от значения порога \(\Delta\).
        \item Публикация репозитория с реализацией Monte Carlo в низкоразмерном подпространстве и скриптов для reproducing экспериментов.
    \end{itemize}
\end{itemize}

Таким образом, раздел \emph{«Планы и будущие работы»} сделает статью более завершённой и позволит читателям ознакомиться с перспективами развития метода. Еще раз благодарю за полезный совет!

\end{document}

### 2025-05-22 slides/inn_prac/rereview_2/Valeriy.tex
Universal commits presents...

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage{amsmath,amssymb}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\begin{document}

\pagestyle{empty}

\begin{center}
    \subsection*{Ответ на замечания рецензента Леванова Валерия Дмитриевича}
\end{center}

% Положительные замечания
\begin{quote}
    \textbf{Рецензент:}
    «Автор чётко обозначает проделанную работу, план действий и ожидаемые результаты. Работа интересна с точки зрения практического применения, так как оптимизация использования вычислительных ресурсов — очень важная тема. Большим плюсом является чёткая структура доклада, но мне кажется, что тут много лишних слайдов. Также не очень понятно, какие дальнейшие действия будут предприняты в работе.»
\end{quote}

\paragraph{Ответ автора.}
Благодарю за содержательный отзыв и высокую оценку практической значимости исследования. Рад, что структура доклада и изложение плана работ показались ясными и полезными.

\begin{itemize}
    \item \textbf{По поводу структуры и объёма слайдов.}
          % Уточнение про лишние слайды
          Я учту замечание о избыточности некоторых слайдов. В окончательном варианте презентации (а также в тексте статьи) будут:
          \begin{itemize}
              \item Сокращены вспомогательные иллюстрации: объединены близкие по смыслу графики $\Delta_k$ для разных итераций в единые панели;
              \item Упрощены избыточные блоки с описанием базовых свойств гессиана, перенесённые в приложение;
              \item Чётко выделен основный поток изложения без второстепенных технических деталей.
          \end{itemize}

    \item \textbf{О дальнейших шагах.}
          % План будущей работы
          В финальной версии статьи будет отдельный раздел «Планы и будущие работы», где подробно описаны:
          \begin{itemize}
              \item Расширение экспериментов на датасетах CIFAR-10/100 и современных архитектурах (ResNet, DenseNet);
              \item Разработка онлайн-версии метода с динамическим отслеживанием числа собственных векторов $d$ и оценкой изменения ландшафта $\Delta_k$;
              \item Уточнение теоретических границ сходимости через неасимптотические оценки и учёт высших моментов распределения данных;
              \item Улучшение визуализации: зависимость точности модели от выбранного порога $\Delta$ и открытый релиз кода.
          \end{itemize}
\end{itemize}

Эти доработки позволят сделать работу более компактной, ясной и завершённой, а также дадут читателям ясное представление о дальнейших перспективах исследования. Еще раз благодарю за ценные замечания!

\end{document}

### 2025-05-22 slides/inn_prac/review_1/Timofey.tex
Universal commits presents...

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage{amsmath,amssymb}

\begin{document}

% ==============================
% Заголовок и идентификация отзыва
% ==============================
\begin{center}
    {\Large \textbf{Отзыв на доклад}}\\[1em]
    {\large \textbf{Название доклада:} Локализация по графу 3D-сцены}\\[0.5em]
    {\large \textbf{Автор доклада:} Кондрашов Тимофей}\\[0.5em]
    {\large \textbf{Автор отзыва:} Никитин Артем Анатольевич}\\[2em]
\end{center}

% ==============================
% Секция 1: Аннотация
% ==============================
% Требования к аннотации:
% - Объём: от 70 до 150 слов (примерно 450–1050 символов без пробелов).
% - Текст должен состоять из нескольких предложений.
% Необходимые элементы (при наличии в докладе):
%   * Формулировка задачи.
%   * Научная мотивация.
%   * Обзор предыдущих результатов (если есть).
%   * Перспективы и ожидаемые результаты.
%   * Используемые методы/планы и наличие промежуточных результатов (если есть).
\section*{Аннотация}
В докладе рассматривается задача локализации робота по данным с RGBD-сенсоров в помещениях, что является одним из ключевых 
направлений при создании автономных систем. Ключевая идея — использовать графы сцен, где вершины соответствуют объектам, 
а рёбра отражают пространственные либо семантические связи. Такой способ представления данных меньше 
зависит от изменений освещения, точки обзора и динамического окружения. Автор отмечает, что существующих исследований, 
посвящённых графам 3D-сцен в задаче локализации, крайне мало, и предлагает изучать подход на базе методики SceneGraphLoc 
и датасета 3RScan. Ожидается, что итоговый результат будет включать нейронную сеть, способную работать только с RGBD-данными 
и эффективно решать задачу локализации робота на основе графов сцен.

% ==============================
% Секция 2: Критический комментарий
% ==============================
% Требования к критическому комментарию:
% - Объём: от 50 до 150 слов (примерно 350–1050 символов без пробелов).
% - Текст должен состоять из нескольких предложений.
% Укажите сильные и слабые стороны доклада, его структуру и научное содержание, 
% а также, чем тема интересна лично вам.
\section*{Комментарий}
Доклад весьма актуален, так как задача локализации робота остается одной из важнейших в робототехнике. Автор 
акцентируетт внимание именно графам 3D-сцен, позволяющим хранить информацию о топологии и семантике объектов, что 
потенциально снижает зависимость от внешних факторов. Презентация получилась структурированной, хотя хотелось 
бы услышать более детальные результаты предварительных тестов. Тем не менее, сравнение с аналогичными методами, упомянутыми 
в докладе (например, SceneGraphLoc), даёт представление о том, как будет организована эксперименты по оценке нейронной сети. 
Лично меня тема привлекает своей связью с компьютерным зрением и возможностями применения в широком спектре автономных систем.

\end{document}

### 2025-05-22 slides/inn_prac/review_2/Alan.tex
Universal commits presents...

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage{amsmath,amssymb}

\begin{document}

% ==============================
% Заголовок и идентификация отзыва
% ==============================
\begin{center}
    {\Large \textbf{Отзыв на доклад}}\\[1em]
    {\large \textbf{Название доклада:} Оценки чисел Рамсея для произвольных последовательностей графов}\\[0.5em]
    {\large \textbf{Автор доклада:} Беремкулов Алан}\\[0.5em]
    {\large \textbf{Автор отзыва:} Никитин Артем Анатольевич}\\[2em]
\end{center}

% ==============================
% Секция 1: Аннотация
% ==============================
\section*{Аннотация}
В докладе рассмотрены обобщения классических чисел Рамсея на произвольные последовательности графов. Проблема заключается
в узости традиционного определения и отсутствии точных оценок, например для \(R(5,5)\). Автор вводит новые параметры
\(R_{\min}(\{G_n\},k)\) и \(R_{\max}(\{G_n\},k)\), которые при соответствующих выборе графов сводятся к классическим
числам Рамсея. Представлены два основных результата: точные оценки \(R_{\max}\) для графов Джонсона (Теорема 1) и
асимптотические границы для произвольных семейств (Теорема 2). Также приведены начальные оценки \(R_{\min}\) и
обсуждаются дальнейшие направления по поиску новых семейств графов и уточнению оценок.

% ==============================
% Секция 2: Критический комментарий
% ==============================
\section*{Критический комментарий}
Доклад отличается чёткой постановкой задачи и строгой математической структурой. Сильной стороной является введение
обобщённых определений \(R_{\min}\) и \(R_{\max}\), расширяющих классическую теорию Рамсея. Презентация логична и
последовательно развивает идеи, однако хотелось бы увидеть конкретные численные примеры или иллюстрации оценок, что помогло 
бы оценить практическую значимость результатов. Было бы полезно продемонстрировать применение теорем на синтетических или 
известных графах. Лично мне тема близка благодаря её связям с дискретной математикой и потенциалу для дальнейших теоретических 
исследований.

\end{document}

### 2025-05-22 slides/inn_prac/review_2/Eldar.tex
Universal commits presents...

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage{amsmath,amssymb}

\begin{document}

% ==============================
% Заголовок и идентификация отзыва
% ==============================
\begin{center}
    {\Large \textbf{Отзыв на доклад}}\\[1em]
    {\large \textbf{Название доклада:} Comparative Analysis of Data-Driven Approaches for Hydrological Forecasting}\\[0.5em]
    {\large \textbf{Авторы доклада:} Eldar Khuzin, Ivan Novikov, Dmitrii Abramov}\\[0.5em]
    {\large \textbf{Автор отзыва:} Никитин Артем Анатольевич}\\[2em]
\end{center}

% ==============================
% Секция 1: Аннотация
% ==============================
% Объём: 70–150 слов, несколько предложений, включает:
% - формулировку задачи и мотивацию;
% - обзор предыдущих подходов;
% - метод и план исследований;
% - ожидаемые результаты.
\section*{Аннотация}
В статье рассматривается проблема эквифинальности в data-driven гидрологическом моделировании, когда разные наборы признаков 
или архитектуры моделей демонстрируют сходную эффективность прогнозов. Актуальность работы определяется необходимостью 
выбора оптимальной модели с учётом точности, интерпретируемости и вычислительных затрат. Авторы анализируют влияние 
статических физико-географических характеристик водосбора на точность прогнозов, сравнивая классические методы, рекуррентные 
нейронные сети и трансформеры на данных по бассейнам рек Евразии. С помощью анализа важности признаков и поведения моделей 
выявляются доминирующие предикторы и определяется баланс между сложностью архитектуры и устойчивостью моделей при 
эквифинальности.

% ==============================
% Секция 2: Критический комментарий
% ==============================
% Объём: 50–150 слов, несколько предложений, включает:
% - сильные и слабые стороны;
% - личную заинтересованность.
\section*{Критический комментарий}
Работа чётко формулирует ключевую проблему эквифинальности и предлагает комплексный сравнительный анализ подходов, что важно 
для практического применения в гидрологии. Особенно ценю исследование влияния статических физико-географических признаков, 
повышающее интерпретируемость результатов. Сильной стороной является сочетание разных моделей и метрик при оценке 
вычислительной эффективности. Вместе с тем эксперименты ограничены бассейнами Евразии, что может сужать обобщаемость 
выводов. Тема представляется мне интересной в контексте оптимизации ресурсов и повышения надёжности моделей гидрологического 
прогнозирования.

\end{document}

### 2025-05-22 slides/inn_prac/review_2/Georgiy.tex
Universal commits presents...

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage{amsmath,amssymb}

\begin{document}

% ==============================
% Заголовок и идентификация отзыва
% ==============================
\begin{center}
    {\Large \textbf{Отзыв на доклад}}\\[1em]
    {\large \textbf{Название доклада:} Detecting Manual Alterations in Biological Image Data Using Contrastive Learning and Pairwise Image Comparison}\\[0.5em]
    {\large \textbf{Автор доклада:} Нехорошков Георгий Сергеевич}\\[0.5em]
    {\large \textbf{Автор отзыва:} Никитин Артем Анатольевич}\\[2em]
\end{center}

% ==============================
% Секция 1: Аннотация
% ==============================
\section*{Аннотация}
Доклад посвящён задаче обнаружения мануальных модификаций биологических изображений, возникающих при фальсификации 
и несанкционированном заимствовании научных данных. Актуальность обусловлена необходимостью защиты оригинальности снимков, 
так как современные SSL-модели (Barlow Twins, SimCLR, CLIP) демонстрируют слабую точность в биомедицинских датасетах. 
Автор предлагает адаптированную архитектуру на базе ResNet-50 с проекторами и линейным классификатором, обучаемыми 
последовательно и параллельно на искусственно модифицированных парах. Используется оптимизатор AdamW с понижающимся 
шагом и специально собранный датасет из 700 изображений. Ожидается существенное улучшение метрик accuracy, F1-score, 
precision, recall и AUC-ROC по сравнению с исходным Barlow Twins.

% ==============================
% Секция 2: Критический комментарий
% ==============================
\section*{Критический комментарий}
Доклад убедительно формулирует проблему устойчивого детектирования подделок биологических снимков и предлагает 
практическое SSL-решение. Сильной стороной является ясное сравнение двух режимов обучения — последовательного и 
параллельного — а также подробное описание архитектуры с ResNet-50 и прожектором. Отдельно стоит отметить тщательный 
подбор датасета и настройку оптимизатора AdamW. Хотелось бы больше информации о распределении типов атак и о чувствительности 
к каждой из них. Тема представляется интересной благодаря прямой применимости верификации результатов биомедицинских 
исследований.

\end{document}

### 2025-05-22 slides/inn_prac/review_2/Timofey.tex
Universal commits presents...

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage{amsmath,amssymb}

\begin{document}

% ==============================
% Заголовок и идентификация отзыва
% ==============================
\begin{center}
    {\Large \textbf{Отзыв на доклад}}\\[1em]
    {\large \textbf{Название доклада:} Локализация по графу сцены (Scene graph place recognition)}\\[0.5em]
    {\large \textbf{Автор доклада:} Кондрашов Тимофей}\\[0.5em]
    {\large \textbf{Автор отзыва:} Никитин Артем Анатольевич}\\[2em]
\end{center}

% ==============================
% Секция 1: Аннотация
% ==============================
\section*{Аннотация}
В докладе представлена задача локализации робота в помещении с помощью графов сцены, где вершины—объекты, а
рёбра отражают пространственные и семантические связи. Актуальность обусловлена необходимостью повышения точности и
надёжности локализации при использовании RGBD-данных и семантической сегментации. В обзоре рассматриваются методы
SceneGraphLoc, демонстрирующие R@1 до 97\% на пяти комнатах, а также соседние подходы (LiDAR GAT, Multiview Scene Graph).
Для практической реализации автор интегрирует библиотеку OpenPlaceRecognition с датасетом 3RScan и изучает альтернативу
DINOv2 в SegVLAD—«мешок запросов» (Bag of Queries). Ожидается повышение метрик поиска места и устойчивость к
изменению окружения.

% ==============================
% Секция 2: Критический комментарий
% ==============================
\section*{Критический комментарий}
Доклад логично структурирован: чёткое определение задачи, обзор ключевых публикаций и анализ базового метода SceneGraphLoc.
Особенно ценным является интеграция с OpenPlaceRecognition и практическая работа с 3RScan. Эксперименты по замене DINOv2
на Bag of Queries и введение весов сегментов выглядят многообещающе. Недостаток—пока нет количественных результатов новых
вариантов, хотелось бы увидеть сравнение до и после изменений. Тема представляется интересной в контексте развития
графовых методов для place recognition; лично я особенно ценю практическую направленность и возможность применения в
реальных автономных системах.

\end{document}

### 2025-05-22 slides/inn_prac/review_2/Valeriy.tex
Universal commits presents...

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T2A]{fontenc}
\usepackage[russian]{babel}
\usepackage{amsmath,amssymb}

\begin{document}

% ==============================
% Заголовок и идентификация отзыва
% ==============================
\begin{center}
    {\Large \textbf{Отзыв на доклад}}\\[1em]
    {\large \textbf{Название доклада:} Использование методов подсчёта неопределённости для борьбы с атаками на детекторы машинно-сгенерированного текста}\\[0.5em]
    {\large \textbf{Автор доклада:} Леванов В.Д.}\\[0.5em]
    {\large \textbf{Автор отзыва:} Никитин Артем Анатольевич}\\[2em]
\end{center}

% ==============================
% Секция 1: Аннотация
% ==============================
\section*{Аннотация}
В докладе рассматривается задача надёжного обнаружения машинно-сгенерированного текста в присутствии атак на генеративную
модель и её вывод. Предлагается использовать методы оценки неопределённости (Uncertainty Estimation) в настройках White-box
и Black-box, ранее успешно применявшиеся в NLP-задачах, но мало исследованные для класcификации Human- vs Machine-Generated.
В качестве экспериментальных данных используются логиты контекста Llama-3.1-8B-Instruct на корпусе M4GT-Bench, где метрики
Perplexity и Mean Token Entropy служат признаками для кластеризации и бинарного классификатора. Ожидается, что обученные на 
RAID и M4GT модели продемонстрируют устойчивую точность при различных атаках.

% ==============================
% Секция 2: Критический комментарий
% ==============================
\section*{Критический комментарий}
Доклад ясно формулирует проблему уязвимости AI-детекторов текстов и обоснованно предлагает Uncertainty Estimation как путь 
к повышению робастности. Сильной стороной является практическое использование реализаций на Llama-3.1-8B-Instruct и корпусе 
M4GT-Bench с демонстрацией кластеризации по Perplexity и Mean Token Entropy. Тем не менее, хотелось бы увидеть конкретные 
численные значения точности и ROC-кривые для разных методов оценки неопределённости. В целом работа выглядит актуальной 
для борьбы с дезинформацией и списыванием, и я лично интересуюсь её продолжением на датасете RAID и расширением набора 
методов Uncertainty Estimation.

\end{document}

### 2025-05-22 slides/inn_prac/thesis/dd.tex
Universal commits presents...

\documentclass[a4paper,12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\geometry{margin=2.5cm}

\begin{document}

\noindent\textbf{UDC 519.688.4}\\[2ex]
\begin{center}
  {\LARGE \textbf{Neural Networks Loss Landscape Convergence in Hessian Low-Dimensional Space}}\\[2ex]
  Tem Nikitin$^{1}$, Nikita Kiselev$^{1}$, Andrey Grabovoy$^{1}$, Vladislav Meshkov$^{1}$\\[1ex]
  $^{1}$Moscow Institute of Physics and Technology, Moscow, Russia
\end{center}

\section*{Thesis}

Deep neural networks often improve when we feed them more data, because extra samples make the loss surface smoother and help the model generalize. But after a certain point, new data barely changes the local shape of the loss, and collecting more can waste time and resources. We need a way to find the minimal amount of data that still shapes the model’s behaviour well.

We offer a simple bound (Theorem~1) for the average squared change in loss:
$$
\Delta_k = \mathbb{E}[L_{k+1}(w) - L_k(w)]^2 \approx \frac{\sigma^4}{4} \Bigl(2\sum_{d=1}^D(\lambda_{k+1}^{(d)} - \lambda_k^{(d)})^2 + \bigl(\sum_{d=1}^D(\lambda_{k+1}^{(d)} - \lambda_k^{(d)})\bigr)^2\Bigr),
$$
where $\lambda_k^{(d)}$ are the top $D$ eigenvalues of the Hessian at batch size $k$, and $\sigma^2$ measures random perturbations. This formula works in a low-dimensional space spanned by the main curvature directions.

Our method runs as follows:
\begin{itemize}
  \item Use power iterations to find the top $D=10$ Hessian eigenvectors.
  \item After each new mini-batch, project updates onto this subspace and compute both the actual $\Delta_k$ by sampling and the bound above.
  \item Stop adding data once the bound drops below a chosen tolerance $\Delta$.
\end{itemize}

We tested on MNIST and Fashion-MNIST with simple MLP and CNN models. Figure~\ref{fig:delta} shows that the theoretical bound stays above the measured $\Delta_k$ curves and shrinks roughly like $O(1/k)$. Table~1 lists the minimal batch sizes $K$ needed for tolerances of 0.025 and 0.030. Values range from 40 to 120, and more complex models converge faster because they capture sharper directions.

Each step costs $O(I\,D\,C_{Hv})$ operations, where $I$ is power-iteration steps and $C_{Hv}$ is the cost of a Hessian-vector product. We also discuss simple tricks, like reducing $D$ or stopping iterations early, to speed this up.

In future work, we plan to apply this to deeper, non-convex networks, data streams, and active sampling. Embedding this check in online learning could let systems decide in real time whether to ask for more data.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\textwidth]{figure9.png}
  \caption{Empirical and theoretical $\Delta_k$ for MNIST experiments.}
  \label{fig:delta}
\end{figure}

\begin{table}[h]
  \centering
  \caption{Δ-sufficient sample size results}
  \begin{tabular}{lccc}
    \hline
    Dataset       & Model & $\Delta$ & $K$ \\
    \hline
    MNIST         & MLP   & 0.025    & 100 \\
    MNIST         & CNN   & 0.025    & 60  \\
    Fashion-MNIST & CNN   & 0.030    & 70  \\
    \hline
  \end{tabular}
\end{table}

\paragraph{Funding and Acknowledgements}
This work was supported by the Moscow Institute of Physics and Technology. We thank the HPC Center for computing resources and helpful discussions on Hessian analysis.

\begin{thebibliography}{9}
  \bibitem{soy2020} D. Soydaner, \textit{A comparison of optimization algorithms for deep learning}, Int. J. Pattern Recognit. Artif. Intell., 34(13):2052013, 2020.
  \bibitem{wu2017} L. Wu, Z. Zhu, W. E., \textit{Towards understanding generalization of deep learning: Perspective of loss landscapes}, arXiv:1706.10239, 2017.
  \bibitem{cookbook} K. B. Petersen, M. S. Pedersen, \textit{The Matrix Cookbook}, 2012.
  \bibitem{deng2012} L. Deng, \textit{The MNIST database of handwritten digit images for machine learning research}, IEEE Signal Process. Mag., 29(6), 2012.
  \bibitem{xiao2017} H. Xiao, K. Rasul, R. Vollgraf, \textit{Fashion-MNIST: a novel image dataset for benchmarking machine learning algorithms}, arXiv:1708.07747, 2017.
\end{thebibliography}

\end{document}

### 2025-05-22 slides/inn_prac/thesis/main.tex
Universal commits presents...

--- 2025-05-22 slides/inn_prac/thesis/dd.tex
@@
 \usepackage{graphicx}
-\usepackage{amsmath}
+\usepackage{amsmath,amssymb}
 \geometry{margin=2.5cm}
@@
 
-\noindent\textbf{UDC 519.688.4}\\[2ex]
+\noindent\textbf{UDC 004.85}\\[2ex]
 \begin{center}
   {\LARGE \textbf{Neural Networks Loss Landscape Convergence in Hessian Low-Dimensional Space}}\\[2ex]
-  Tem Nikitin$^{1}$, Nikita Kiselev$^{1}$, Andrey Grabovoy$^{1}$, Vladislav Meshkov$^{1}$\\[1ex]
-  $^{1}$Moscow Institute of Physics and Technology, Moscow, Russia
+  Tem Nikitin, Nikita Kiselev\\[1ex]
+  Moscow Institute of Physics and Technology, Moscow, Russia
 \end{center}
+Understanding how a neural network’s loss landscape changes as we add more training data is important for efficient training.
+Although larger datasets reshape this high-dimensional surface, the point when extra data stop making a big difference is unclear.
+We show that near a local minimum the loss landscape stabilizes once the dataset exceeds a certain size.
+To study this, we project the full parameter space onto a lower-dimensional subspace formed by the Hessian’s top eigenvectors, highlighting the most important curvature directions.
+Within this subspace, we use Monte Carlo sampling to estimate how the loss changes more precisely.
+Experiments on standard image classification tasks demonstrate that our low-dimensional analysis pinpoints when the landscape stops evolving, offering practical guidance for balancing training cost with performance improvements.
 
@@
 
-We tested on MNIST and Fashion-MNIST with simple MLP and CNN models.
-Figure~\ref{fig:delta} shows that the theoretical bound stays above the measured $\Delta_k$ curves and shrinks roughly like $O(1/k)$.
+We tested on MNIST and Fashion-MNIST with simple MLP and CNN models. Figure~\ref{fig:delta} shows that the 
+theoretical bound stays above the measured $\Delta_k$ curves and shrinks roughly like $O(1/k^2)$.
 Table~1 lists the minimal batch sizes $K$ needed for tolerances of 0.025 and 0.030.
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \centering
-  \includegraphics[width=0.8\textwidth]{figure9.png}
+  \includegraphics[width=0.8\textwidth]{img/delta_border_1_10_1024.pdf}
   \caption{Empirical and theoretical $\Delta_k$ for MNIST experiments.}
@@
 
-\begin{table}[h]
+\begin{table}[ht]
   \centering
-  \caption{Δ-sufficient sample size results}
+  \caption{$\Delta$-sufficient sample size results}
   \begin{tabular}{lccc}
@@
     \hline
-    MNIST         & MLP   & 0.025    & 100 \\
-    MNIST         & CNN   & 0.025    & 60  \\
+    MNIST & Single-layer MLP & 0.025 & 100 \\
+    MNIST & Convolutional CNN & 0.025 & 60 \\
     Fashion-MNIST & CNN & 0.030 & 70 \\

### 2025-05-22 slides/full/main.tex
Universal commits presents...

@@
 \usepackage{hyperref}
+\usepackage{tabularx}   
 \usepackage{hhline}
-\graphicspath{{fig/}{../fig/}}
+
+\setbeamertemplate{itemize item}{}
+\setbeamertemplate{itemize subitem}{}
+\setbeamertemplate{itemize subsubitem}{}
 
@@
 \title[Loss Landscape Convergence]{Neural Networks Loss Landscape Convergence in Hessian Low-Dimensional Space}
-\author[Tem Nikitin \textit{et al.}]{Tem Nikitin \and Nikita Kiselev \and Andrey Grabovoy \and Vladislav Meshkov}
+\author[Tem Nikitin \textit{et al.}]{Tem Nikitin, Nikita Kiselev, Vladislav Meshkov, Andrey Grabovoy}
 \institute{Moscow Institute of Physics and Technology}
@@
 
-\begin{frame}{Goals}
-    \begin{itemize}
+\begin{frame}{Neural Networks Loss Landscape Convergence in Hessian Low-Dimensional Space}
+    \textbf{Goals and Tasks}
+    \begin{enumerate}
         \item Study how neural network loss landscape changes with dataset size
@@
         \item Derive theoretical bound for $\Delta_k$ via top-$d$ eigenvalues
+        \item Propose an algorithm to determine the $\Delta$-sufficient dataset size
         \item Validate threshold $k^*$ beyond which further data yield negligible change
-    \end{itemize}
+    \end{enumerate}
 \end{frame}
 
-\begin{frame}{One-slide}
+\begin{frame}{Loss landscape stabilizes after sufficient sample size.}
     \begin{columns}[c]
@@
         \column{0.6\textwidth}
-        \includegraphics[width=1.2\textwidth]{img/loss_eigen_-1_individ.pdf}
+        \includegraphics[width=1.34\textwidth]{img/loss_eigen_-1_individ.pdf}
         \column{0.5\textwidth}
@@
     \end{columns}
-    \bigskip
-    \textbf{Main message:} Loss landscape stabilizes after sufficient sample size.
 \end{frame}
 
-\begin{frame}{Literature}
-    \begin{itemize}
-        \item Wu \textit{et al.} (2017): loss landscapes vs dataset size [2]
-        \item Sagun \textit{et al.} (2018): Hessian low effective rank [7]
-        \item Li \textit{et al.} (2018): visualizing loss surfaces [11]
-        \item Ghorbani \textit{et al.} (2019): eigenvalue density analysis [8]
-        \item Bousquet \& Elisseeff (2002): stability and generalization bounds [12]
-    \end{itemize}
-\end{frame}
-
-\begin{frame}{Problem Statement: Hypothesis and Model}
+\begin{frame}{Problem Statement}
     \begin{block}{Hypothesis}
@@
     \end{block}
-\end{frame}
 
-\begin{frame}{Problem Statement: Quality Criteria}
+    \begin{block}{Criteria}
         \begin{itemize}
-        \item Convergence rate: $\Delta_k = O(1/k)$
+            \item Convergence rate: $\Delta_k = O(1/k^2)$
             \item Theoretical bound via top-$d$ eigenvalues upper-bounds empirical $\Delta_k$
@@
         \end{itemize}
+    \end{block}
 \end{frame}
 
-\begin{frame}{Problem Solution: Theoretical Analysis}
+\begin{frame}{Theoretical Analysis}
     \begin{columns}[c]
@@
                   \bigskip
-            \item Bound:
         \end{itemize}
-        \column{0.6\textwidth}
+        \column{0.8\textwidth}
         \includegraphics[width=1.0\textwidth]{img/eigenvalues.pdf}
-        \vspace{0.5cm}\hspace{1.8cm}\raggedright{\tiny Figure 1: Eigenvalue decay (p.4)}
+        \vspace{0.5cm}\hspace{3.1cm}\raggedright{Eigenvalue decay}
     \end{columns}
-    $$
-        \Delta_k \approx
+    Bound:
+    $ \Delta_k \approx
         \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i)^2
-        + \Bigl( \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i) \Bigr)^2 \Biggr).
-    $$
+        + \Bigl( \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i) \Bigr)^2 \Biggr).$
 \end{frame}
 
-\begin{frame}{Goals of Computational Experiment}
+\begin{frame}{Computational Experiment}
     \begin{itemize}
@@
         \centering
-        \includegraphics[width=1.05\textwidth,trim=0 17 0 1.5cm,clip]{img/delta_eigen_1_10_64.pdf}
-        \caption*{\tiny Figure 8: Monte Carlo $\Delta_k$ vs $k$ and $\Delta_k\cdot k^2$ (p.11)}
+        \includegraphics[width=1.05\textwidth,trim=0 17 0 1.5cm,clip]{img/delta_eigen_1_10_1024.pdf}
+        \caption*{Monte Carlo $\Delta_k$ vs $k$ and $\Delta_k\cdot k^2$}
     \end{figure}
@@
 
-\begin{frame}{Error Analysis}
+\begin{frame}{Combining results}
     \begin{itemize}
@@
         \centering
-        \includegraphics[width=0.7\textwidth]{img/delta_border_1_10_64.pdf}
-        \caption*{\tiny Figure 9: Theoretical (dashed) vs empirical (solid) $\Delta_k$ (p.12)}
+        \includegraphics[width=0.9\textwidth]{img/delta_border_1_10_1024.pdf}
+        \caption*{Theoretical (dashed) vs empirical (solid) $\Delta_k$}
     \end{figure}
@@
 
-%\begin{frame}{Models--Datasets--Criteria}
-%    \begin{center}
-%        \begin{tabular}{|l|l|c|c|}
-%            \hline
-%            \textbf{Model} & \textbf{Dataset} & \boldmath$d$ & \boldmath$\Delta_{tol}$ \\
-%            \hhline{|=|=|=|=|}
-%            MLP (2 layers) & MNIST            & 10           & $1\times10^{-4}$        \\
-%            \hline
-%            MLP (2 layers) & Fashion-MNIST    & 10           & $1\times10^{-4}$        \\
-%            \hline
-%        \end{tabular}
-%    \end{center}
-%\end{frame}
+\begin{frame}{Theoretical vs. Empirical $\Delta_k$}
+    \begin{figure}
+        \centering
+        \includegraphics[width=1\textwidth]{img/delta_loss_1_10_1024.pdf}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Practical Measurements}
+    \begin{table}
+        \centering
+        \small
+        \begin{tabularx}{\textwidth}{|l|>{\raggedright\arraybackslash}X|c|c|c|c|}
+            \hline
+            \textbf{Dataset}
+                          & \textbf{Model}    & \(\Delta\) & \(k\) & \(L_k\) & \textbf{Time (s)} \\
+            \hhline{|=|=|=|=|=|=|}
+            MNIST         & Single-layer MLP  & 0.025      & 100   & 0.013   & 2000              \\
+                          & Multi-layer MLP   & 0.025      & 40    & 0.010   & 3500              \\
+                          & Convolutional CNN & 0.025      & 60    & 0.024   & 1800              \\
+            \hline
+            Fashion-MNIST & Single-layer MLP  & 0.030      & 120   & 0.020   & 2100              \\
+                          & Multi-layer MLP   & 0.030      & 90    & 0.017   & 4400              \\
+                          & Convolutional CNN & 0.030      & 70    & 0.015   & 2400              \\
+            \hline
+        \end{tabularx}
+    \end{table}
+\end{frame}
+
+\begin{frame}{Algorithm: \(\Delta\)-sufficient Dataset Size}
+    \begin{block}{Algorithm 1: Determine \(\Delta\)-sufficient dataset size}
+        \begin{enumerate}
+            \item Initialize dataset \(D \gets \emptyset\) and batch counter \(k \gets 0\).
+            \item \textbf{Repeat:}
+                  \begin{itemize}
+                      \item \(k \gets k + 1\).
+                      \item Sample a new batch of data and append to \(D\).
+                      \item Train or update model on \(D\).
+                      \item Compute top-\(d\) Hessian eigenvalues \(\{\lambda_k^{(i)}\}_{i=1}^d\) and \(\{\lambda_{k+1}^{(i)}\}\).
+                      \item Estimate
+                            \[
+                                \Delta_k \approx \frac{\sigma^4}{4}\Bigl(2\sum_{i=1}^d(\lambda_{k+1}^{(i)}-\lambda_k^{(i)})^2
+                                + \bigl(\sum_{i=1}^d(\lambda_{k+1}^{(i)}-\lambda_k^{(i)})\bigr)^2\Bigr).
+                            \]
+                  \end{itemize}
+            \item \textbf{Until} \(\Delta_k < \Delta_{\text{tol}}\).
+            \item \textbf{Return} \(|D|\), the \(\Delta\)-sufficient sample size.
+        \end{enumerate}
+    \end{block}
+\end{frame}
 
 \begin{frame}{Results and Conclusions}
-    \begin{itemize}
-        \item Identified threshold $k$ (MNIST), (Fashion-MNIST)
-        \item Loss landscape stabilizes: additional data negligible beyond $k^*$
-        \item Hessian-based bound provides reliable upper-bound for $\Delta_k$
-        \item Practical guideline for dataset sizing and early stopping
-    \end{itemize}
+    \begin{enumerate}
+        \item Identified dataset thresholds \(k^*\) for MNIST and Fashion-MNIST.
+        \item Loss landscape stabilizes: additional data negligible beyond \(k^*\).
+        \item Hessian-based bound provides a reliable upper-bound for \(\Delta_k\).
+        \item Proposed a practical algorithm for \(\Delta\)-sufficient dataset sizing.
+        \item Offers guidelines for dataset collection and early stopping.
+    \end{enumerate}
+\end{frame}
+
+\begin{frame}{Literature}
+    \begin{enumerate}
+        \item Wu \textit{et al.} (2017): loss landscapes vs dataset size
+        \item Sagun \textit{et al.} (2018): Hessian low effective rank
+        \item Li \textit{et al.} (2018): visualizing loss surfaces
+        \item Ghorbani \textit{et al.} (2019): eigenvalue density analysis
+        \item Bousquet \& Elisseeff (2002): stability and generalization bounds
+    \end{enumerate}
 \end{frame}

### 2025-05-22 slides/full/main.tex
Universal commits presents...

@@
         \item Study how neural network loss landscape changes with dataset size
-        \item Define and measure $\Delta_k = \mathbb{E} \bigl( \cL_{k+1}(\bw) - \cL_k(\bw) \bigr)^2$
-        \item Develop Hessian-based low-dimensional projection method
+        \item Define and measure metrics $\Delta_k$
         \item Derive theoretical bound for $\Delta_k$ via top-$d$ eigenvalues
         \item Propose an algorithm to determine the $\Delta$-sufficient dataset size
-        \item Validate threshold $k^*$ beyond which further data yield negligible change
     \end{enumerate}
@@
 
-\begin{frame}{Results and Conclusions}
+\begin{frame}{Summary}
+    \begin{block}{Results}
         \begin{enumerate}
-        \item Identified dataset thresholds \(k^*\) for MNIST and Fashion-MNIST.
-        \item Loss landscape stabilizes: additional data negligible beyond \(k^*\).
+            \item Loss landscape stabilizes.
             \item Hessian-based bound provides a reliable upper-bound for \(\Delta_k\).
@@
         \end{enumerate}
+    \end{block}
+
+    \begin{block}{Future work}
+    \begin{enumerate}
+        \item Test the method on additional architectures.
+        \item Evaluate the approach on other datasets.
+        \item Incorporate automatic selection of the optimal number of Hessian eigenvalues.
+    \end{enumerate}
+    \end{block}
 \end{frame}
